### 2020-03-04 analysis/intro.Rmd

---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(datasets)
```

```{r}
library(tidyverse)
library(magrittr)

data(iris)
head(iris)
```

```{r}
# I do not like uppercase:
iris %<>% rename_all(tolower)
head(iris)

# I prefer tibble format:
iris %<>% as.tibble 
iris

```

We want to compute the mean value for each variable for each species. We will
use the functions of the `dplyr` package.

```{r}
iris_species <- iris %>%
  gather(variable, value, sepal.length:petal.width) %>%
  group_by(species, variable) %>%
  summarise(avg = mean(value)) %>%
  spread(variable, avg)
```

## Tables

You can generate tables in markdown with `kable()` but you can also generate it
in LaTeX.

```{r}
knitr::kable(iris_species)

```

```{r}
library(xtable)
xtable(iris_species)

```

Let's tweak a bit the table by removing row numbering, reduce digit number,
etc... See `r ?xtable` and `r ?print.xtable`

```{r}
iris_species_xtable <- xtable(iris_species, 
  caption = "Average flower phenology (in cm) of different species of the
  \textit{Iris} genus.",
  label = "tab:iris_species",
  align = c("c","l", "|", "r", "r", "r", "r"),
  digits = 1
  )
print(iris_species_xtable, include.rownames = FALSE)

```

## Figures

```{r}
iris_petall <- select(iris, species, petal.length)
p <- ggplot(iris_petall, aes(y = petal.length, x = species)) + 
  geom_violin()
p
```

This plot is not ready to be put in a paper (grey area, bad axes, font not
ajusted). The package `cowplot` load a theme with better default. See the
cowplot
[vignette](https://cran.r-project.org/web/packages/cowplot/vignettes/introduction.html).

```{r}
library(cowplot)
p

```

You can also arrange plots in a grid:

```{r}
p2 <- ggplot(mpg, aes(x = cty, y = hwy, colour = factor(cyl))) + 
  geom_point(size=2.5)
p2

```
```{r, fig.dim = c(7, 5)}
plot_grid(p, p2, labels = "AUTO")
```