### 2020-12-01 temporal_dynamics.Rmd
functions to generate plot of the report

---
title: "Temporal dynamics"
author: "Alain Danet"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    math: katex
---

```{r}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE 
)

library(drake)
library(tidyverse)
library(magrittr)
library(cowplot)
library(viridis)
source("R/misc.R") # Define your custom code as a bunch of functions.
source_dir(get_mypath("R")) # Define your custom code as a bunch of functions.

#theme_set(theme_cowplot())
```

```{r}
loadd(rigal_classification, full_data2, temporal_dynamics_plot)
```

```{r}
myload(biomass_ts_sax, envir = globalenv(),dir = get_mypath("data"))

biomass_classif_rigal <- rigal_classification %>%
  filter(variable %in% c("bm_std", "log_bm_std")) %>%
  unnest(classif) %>%
  mutate(station = as.character(station)) %>%
  select(station, variable,shape_class)

p_biomass_classif_rigal <- biomass_classif_rigal %>%
  ggplot(aes(x = shape_class)) +
  geom_bar() +
  facet_wrap(~variable) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Temporal dynamics

```{r}
p_cap <- paste0(
  "Temporal dynamics of biomass and temporal dynamics of
  metrics for 10 randomly selected communities."
)
```

```{r p-tmp-dyn, cache = TRUE, eval = FALSE}

p <- vector(mode = "list", length = 10)
for (i in seq_len(length(p))) {
  to_plot <- temporal_dynamics_plot %>% 
    filter(station == sample(temporal_dynamics_plot$station, 1))
  p[[i]] <- cowplot::plot_grid(plotlist = to_plot$plot, labels = to_plot$station)
}
p
```

## Temporal dynamics decrease/increase constant

### Log biomass

```{r, cache = TRUE, fig.height = 45}
st_decrease_increase_rigal <- biomass_classif_rigal %>%
  filter(variable == "log_bm_std") %>%
  filter(shape_class %in% c("increase_constant", "decrease_constant"))
data_decrease_increase <- full_data2 %>%
  filter(station %in% st_decrease_increase_rigal$station)

log_bm_min_max <- map_dbl(c(min, max),
  ~.x(data_decrease_increase$log_bm_std, na.rm = TRUE))
nb_year_min_max <- map_dbl(c(min, max),
  ~.x(data_decrease_increase$nb_year, na.rm = TRUE))
temporal_dynamics_plot$plot %<>%
  map(., ~ .x + lims(y = log_bm_min_max, x = nb_year_min_max))

temporal_dynamics_plot %>%
  filter(variable == "log_bm_std", station %in% st_decrease_increase_rigal$station) %>%
  plot_grid(plotlist = .$plot, ncol = 3, labels = st_decrease_increase_rigal$station)
```

### Raw Biomass

```{r, cache = TRUE, fig.height = 45}
st_decrease_increase_rigal_biomass <- rigal_classification %>%
  filter(variable == "bm_std") %>%
  unnest(classif) %>%
  filter(shape_class %in% c("increase_constant", "decrease_constant"))

bm_min_max <- map_dbl(
  c(min, max),
  ~.x(data_decrease_increase$bm_std, na.rm = TRUE))
nb_year_min_max <- map_dbl(
  c(min, max),
  ~.x(data_decrease_increase$nb_year, na.rm = TRUE))
temporal_dynamics_plot[temporal_dynamics_plot$variable == "bm_std",]$plot %<>%
  map(., ~ .x + lims(y = bm_min_max, x = nb_year_min_max))

temporal_dynamics_plot %>%
  filter(variable == "bm_std", station %in% st_decrease_increase_rigal_biomass$station) %>%
  plot_grid(plotlist = .$plot, ncol = 3, labels = st_decrease_increase_rigal_biomass$station)
```

### Richness

```{r, cache = TRUE, fig.height = 45}
var_tmp <- "rich_std"
st_decrease_increase_rigal <- rigal_classification %>%
  filter(variable == var_tmp) %>%
  unnest(classif) %>%
  filter(shape_class %in% c("increase_constant", "decrease_constant"))

min_max <- map_dbl(
  c(min, max),
  ~.x(data_decrease_increase[[var_tmp]], na.rm = TRUE))
nb_year_min_max <- map_dbl(
  c(min, max),
  ~.x(data_decrease_increase$nb_year, na.rm = TRUE))
temporal_dynamics_plot[temporal_dynamics_plot$variable == var_tmp,]$plot %<>%
  map(., ~ .x + lims(y = min_max, x = nb_year_min_max))

temporal_dynamics_plot %>%
  filter(variable == var_tmp, station %in% st_decrease_increase_rigal$station) %>%
  plot_grid(plotlist = .$plot, ncol = 3, labels = st_decrease_increase_rigal$station)
```

### Number of nodes

```{r, cache = TRUE, fig.height = 45}
var_tmp <- "nbnode_std"
st_decrease_increase_rigal <- rigal_classification %>%
  filter(variable == var_tmp) %>%
  unnest(classif) %>%
  filter(shape_class %in% c("increase_constant", "decrease_constant"))

min_max <- map_dbl(
  c(min, max),
  ~.x(data_decrease_increase[[var_tmp]], na.rm = TRUE))
nb_year_min_max <- map_dbl(
  c(min, max),
  ~.x(data_decrease_increase$nb_year, na.rm = TRUE))
temporal_dynamics_plot[temporal_dynamics_plot$variable == var_tmp,]$plot %<>%
  map(., ~ .x + lims(y = min_max, x = nb_year_min_max))

temporal_dynamics_plot %>%
  filter(variable == var_tmp, station %in% st_decrease_increase_rigal$station) %>%
  plot_grid(plotlist = .$plot, ncol = 3, labels = st_decrease_increase_rigal$station)
```