### 2020-12-23 supplementary.Rmd
Compute l, ld, gini, fish-fish network, temporal network, species trends

---
title: "Supplementary methods and results"
author: "Alain Danet"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    math: katex
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE 
)
```

```{r}
library(drake)
library(tidyverse)
library(magrittr)
library(cowplot)
library(viridis)
library(kableExtra)
library(sf)
source("R/misc.R") # Define your custom code as a bunch of functions.
source("R/variable_shortcut.R")
source_dir(get_mypath("R")) # Define your custom code as a bunch of functions.

#theme_set(theme_cowplot())
```

```{r}
loadd(temporal_dynamics_coef)
loadd(temporal_dynamics_plot)
loadd(full_data2)
loadd(biomass_group)
```

# Methods

## Station grouping

### Rigal station increase / decrease

Les stations ont moins de 40 kilos au départ ont l'air comparable en terme de
dynamique.

```{r, eval=FALSE}

p_rigal_decrease_increase_slope_st <- 
  st_decrease_increase_rigal_biomass %>%
  left_join(biomass_group, by = "station") %>%
  ggplot(aes(y = linear_slope, x = bm_slope, color = shape_class)) +
  geom_point() +
  facet_wrap(~variable)
p_rigal_decrease_increase_slope_st
```

### Tendancies in rigal increase / decrease

```{r}
bm_vs_net_trends %>%
  ggplot(aes(x = reg_weight)) +
  geom_histogram() +
  facet_wrap(~variable)

bm_net_group_median %>%
  ggplot(aes(x = bm_slope, y = linear_slope, size = reg_weight, weight = reg_weight, color = log(bm))) +
  geom_point() +
  scale_color_viridis() +
  geom_vline(xintercept = 0, color = "red") +
  geom_smooth() +
  facet_wrap(~variable, scales = "free_y") +
  labs(x = "Biomass trend (g.m-1.y-1)", y = "Network trend (U.y-1)")

log_bm_net_group_f3y %>%
  ggplot(aes(x = bm_slope, y = linear_slope, size = reg_weight, weight = reg_weight, color = log(bm))) +
  geom_point() +
  scale_color_viridis() +
  geom_vline(xintercept = 0, color = "red") +
  geom_smooth() +
  facet_wrap(~variable, scales = "free_y") +
  labs(x = "Log Biomass trend (g.m-1.y-1)", y = "Network trend (U.y-1)")
```

### With group

#### Median

```{r}
##Summary of group 
biomass_group %>%
  filter(group_type == "median") %>%
  ggplot(aes(x = bm)) +
  geom_histogram() +
  facet_wrap(bm_var~summary_type, scales = "free_x")
```
```{r, eval = FALSE}
group_results <- 
  rigal_biomass_coeff_dynamics %>%
  left_join(biomass_group, by = "station") %>% 
  filter(group_type == "median", bm_var == "bm_std")

p1 <- group_results %>%
  filter(summary_type == "first_3_year") %>%
  ggplot(aes(x = linear_slope, y = estimate, color = group,
      size = reg_weight, weight = reg_weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~coeff, scales = "free_y")

p3 <- group_results  %>%
  filter(summary_type == "median") %>%
  ggplot(aes(x = linear_slope, y = estimate, color = group, size = reg_weight, weight = reg_weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~coeff, scales = "free_y")

p1
p3
```

#### Little and big rivers

```{r}
loadd(stream_group)
#grandes/petites rivières
stream_group %>%
  ggplot(aes(x = width_river_mean, fill = group)) +
  geom_histogram() +
  facet_grid(~group_type)

stream_group %>%
  filter(group_type == "median") %>%
  ggplot(aes(x = width_river_mean, fill = protocol_type)) +
  geom_histogram()

stream_group %>%
  filter(group_type == "median") %>%
  left_join(
    biomass_group %>%
    filter(group_type == "median", summary_type == "median") %>%
    select(-group_type, -summary_type, -group),
  by = "station"
  ) %>%
  filter(!is.na(protocol_type)) %>%
  ggplot(aes(x = bm, fill = protocol_type)) +
  geom_histogram() +
  facet_wrap(~bm_var, scales = "free_x")
```

# Patterns

```{r}
bm_net_group_median %>%
  ggplot(aes(x = bm_slope, y = linear_slope, color = log(bm), size = reg_weight, weight = reg_weight)) +
  geom_point() +
  scale_color_viridis() +
  #geom_smooth(method = "lm") +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Group as mean biomass")

bm_net_group_median %>%
  ggplot(aes(x = bm_slope, y = linear_slope, color = log(bm))) +
  geom_point() +
  scale_color_viridis() +
  geom_errorbar(aes(
      xmin = bm_slope - bm_slope_strd_error,
      xmax = bm_slope + bm_slope_strd_error),
    alpha = .2
    ) +
  geom_errorbar(aes(
      ymin = linear_slope - linear_slope_strd_error,
      ymax = linear_slope + linear_slope_strd_error),
    alpha = .2
    ) +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Group as mean biomass")

bm_net_group_median %>%
  filter(!is.na(protocol_type)) %>%
  ggplot(aes(x = bm_slope, y = linear_slope, color = log(bm), size = reg_weight, weight = reg_weight)) +
  geom_point() +
  scale_color_viridis() +
  #geom_smooth(method = "lm") +
  facet_grid(variable~protocol_type, scales = "free") +
  labs(title = "Group as mean biomass")
```

```{r, fig.height = 12, fig.width = 12}
log_bm_net_group_f3y %>%
  filter(!is.na(protocol_type)) %>%
  ggplot(aes(x = bm_slope, y = linear_slope, color = log(bm))) +
  geom_point() +
  scale_color_viridis() +
  geom_errorbar(aes(xmin = bm_slope - bm_slope_strd_error,
      xmax = bm_slope + bm_slope_strd_error), alpha = .3) +
  geom_errorbar(aes(ymin = linear_slope - linear_slope_strd_error,
      ymax = linear_slope + linear_slope_strd_error), alpha = .3) +
  facet_grid(variable~protocol_type, scales = "free") +
  labs_log_bm_net_trends()
```

```{r}
knitr::opts_chunk$set(
  eval = FALSE 
)
```

## Other Statistical model

#### Quadratic one

```{r}
pred_log_bm_quad <- summary_log_bm_f3y %>%
  filter(model == "mod_bm_quad") %>%
  mutate(
    pred = map(model_obj, ~ggpredict(.x,
	terms = c("bm_slope", "bm [quart2]"))),
    pred_df = map(pred, as_tibble),
    plot = map(pred, ~plot(.x, add.data = FALSE)),
    plot_data = map(pred, ~plot(.x, add.data = TRUE)),
    #myplot = map2(pred, data, ~plot_final_model(ggpred = .x, rawdata = .y))
  )

test <- map2(model_log_bm_f3y$data, pred_log_bm_quad$pred_df, plot_pred_data,
  std_error = TRUE)
plot_grid(plotlist = test, labels = model_log_bm_f3y$variable)
```

```{r}

vif_log_bm_f3y %>%
  ggplot(aes(y= vif, x = term)) +
  geom_hline(yintercept = 10) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(model_type~variable, scales = "free")
```

### By protocol

```{r}
loadd(vif_log_bm_by_protocol)

protocol_vec <- vif_log_bm_by_protocol$protocol_type %>% unique
p_vif_protocol <- map(protocol_vec, function (x) {
  vif_log_bm_by_protocol %>%
    filter(protocol_type == x) %>%
    ggplot(aes(y= vif, x = term)) +
    geom_hline(yintercept = 10) +
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = paste0("Protocol type: ", x)) +
    facet_grid(model_type~variable, scales = "free_y")
  })
p_vif_protocol[[1]]
p_vif_protocol[[2]]
```

```{r}
loadd(model_log_bm_f3y_by_protocol, summary_log_bm_f3y_by_protocol)
```

```{r, eval = FALSE}
complete_log_quad2 <- get_table_from_summary(.data = summary_log_bm_f3y_by_protocol, model = "mod_bm_quad2", protocol = "complete")

complete_log_quad2[["reg_table"]] %>% 
  kable(.,
    booktabs = TRUE, 
    caption = 
      paste0(
      "Regression table for the model: dN =",
      "bm + dBm + dBm^2 + bm*dBm + bm*dBm^2."
    ) 
  )

complete_log_quad2[["anova_table"]] %>% 
  kable(., booktabs = TRUE,
    caption = 
      paste0(
      "ANOVA table for the model: dN =",
      "bm + dBm + dBm^2 + bm*dBm + bm*dBm^2."
    )
  )
```

```{r, eval = FALSE}
filter(summary_log_bm_f3y_by_protocol, model == "mod_log_bm_inc")[["resume"]]
filter(summary_log_bm_f3y_by_protocol, model == "mod_log_bm_inc")[["anova"]]

filter(summary_log_bm_f3y_by_protocol, model == "mod_bm_quad")[["resume"]]
filter(summary_log_bm_f3y_by_protocol, model == "mod_log_bm_inc")[["anova"]]
```

#### Quadratic two

```{r}
pred_log_bm_quad2 <- summary_log_bm_f3y_by_protocol %>%
  filter(model == "mod_bm_quad2", protocol_type == "complete") %>%
  mutate(
    pred = map(model_obj, ~ggpredict(.x,
	terms = c("bm_slope", "bm [quart2]"))),
    pred_df = map(pred, as_tibble),
    plot = map(pred, ~plot(.x, add.data = FALSE)),
    plot_data = map(pred, ~plot(.x, add.data = TRUE)),
    #myplot = map2(pred, data, ~plot_final_model(ggpred = .x, rawdata = .y))
  )
model_log_bm_f3y_by_protocol %<>% filter(protocol_type == "complete")

pred_log_bm_quad2_plot <- map2(model_log_bm_f3y_by_protocol$data, pred_log_bm_quad2$pred_df,
  plot_pred_data)
plot_grid(plotlist = pred_log_bm_quad2_plot, labels = model_log_bm_f3y$variable)
```

#### Quadratic one

```{r}
pred_log_bm_quad <- summary_log_bm_f3y_by_protocol %>%
  filter(model == "mod_bm_quad", protocol_type == "complete") %>%
  mutate(
    pred = map(model_obj, ~ggpredict(.x,
	terms = c("bm_slope", "bm [quart2]"))),
    pred_df = map(pred, as_tibble),
    plot = map(pred, ~plot(.x, add.data = FALSE)),
    plot_data = map(pred, ~plot(.x, add.data = TRUE)),
    #myplot = map2(pred, data, ~plot_final_model(ggpred = .x, rawdata = .y))
  )
model_log_bm_f3y_by_protocol %<>% filter(protocol_type == "complete")

pred_log_bm_quad_plot <- map2(model_log_bm_f3y$data, pred_log_bm_quad$pred_df, plot_pred_data,
  std_error = TRUE)
plot_grid(plotlist = pred_log_bm_quad_plot, labels = model_log_bm_f3y$variable)
```

```{r, eval = FALSE}
pred <- summary_log_bm_f3y_by_protocol %>%
  filter(model == "mod_log_bm_inc") %>%
  mutate(
    pred = map(model_obj, ~ggpredict(.x,
	terms = c("bm_slope", "log_bm [quart2]", "inc_f"))),
    pred_df = map(pred, as_tibble),
    plot = map(pred, ~plot(.x, add.data = FALSE)),
    plot_data = map(pred, ~plot(.x, add.data = TRUE)),
  )
plot_grid(plotlist = pred$plot, labels = pred$variable)
```

```{r, eval = FALSE}
pred <- summary_log_bm_f3y_by_protocol %>%
  filter(model == "mod_log_bm_inc") %>%
  mutate(
    pred = map(model_obj, ~ggpredict(.x,
	terms = c("bm_slope", "log_bm [quart2]", "inc_f"))),
    pred_df = map(pred, as_tibble),
    plot = map(pred, ~plot(.x, add.data = FALSE)),
    plot_data = map(pred, ~plot(.x, add.data = TRUE)),
  )
plot_grid(plotlist = pred$plot, labels = pred$variable)
```

### Can we predict the model without all simple effects?

```{r, eval = FALSE}
# Perform a valid model 
pred <- model_int %>%
  select(-data, -mod4) %>%
  pivot_longer(cols = c(mod1:mod5), names_to = "model_type", values_to =
    "mod_obj") %>%
  mutate(
    pred_ref = map2(mod_obj, model_type, function (x, y) {
       if (y %in% c("mod3", "mod5"))
	 terms_spec <- c("bm_slope", "inc_f", "bm")
       else
	 terms_spec <- c("bm_slope", "inc_f", "bm_group") 
      ggpredict(x, terms = terms_spec)
      }
    ),
    pred_cte = map2(mod_obj, model_type, function (x, y) {
       if (y %in% c("mod3", "mod5"))
	 terms_spec <- c("linear_slope", "inc_f", "bm")
       else
	 terms_spec <- c("linear_slope", "inc_f", "bm_group") 
      ggeffect(x, terms = terms_spec)
      }
    )
  )

pred %<>%
  mutate(
  plot_pred_ref = map(pred_ref, plot, add.data = TRUE),
  plot_pred_cte = map(pred_cte, plot, add.data = TRUE)
  )
pred_richness <- pred %>%
  filter(variable == "richness", model_type == "mod3")

pred$plot_pred_ref[[1]]
pred$plot_pred_cte[[1]]
pred$plot_pred[[2]]
pred$plot_pred[[3]]
pred$plot_pred_ref[[4]]
```

```{r, eval = FALSE}
library(ggeffects)
mydf <- ggpredict(model_int$mod3[[1]], terms = c("linear_slope", "inc_f",
    "bm"))
plot(mydf, add.data = TRUE)

```

```{r, eval = FALSE}
model_log %<>%
  mutate(
    pred5 = map(mod3, ~ggpredict(.x,
	terms = c("linear_slope [-8, -6, -4,-3, -2, -1, -.5, -.01, .01, .5, 1, 2, 3, 4, 6]", "inc_f", "bm"))),
    pred5_df = map(pred5, as_tibble),
    plot = map(pred5, ~plot(.x, add.data = TRUE)),
    myplot = map2(pred5, data, ~plot_final_model(ggpred = .x, rawdata = .y))
  )

#plot_final_model(model_int$pred5[[1]], model_int$data[[1]])
plot_grid(plotlist = model_int$plot, labels = model_int$variable)
model_int$plot[[1]]
summary(model_int$mod3[[1]])
test <- model_int$data[[1]]
model_int$pred5_df[[1]]
test %>%
  ggplot(aes(x = bm_slope, y = linear_slope, weight = reg_weight, size = reg_weight)) +
  geom_smooth(method = "lm") +
  geom_point()

pred_test <-
  model_int$pred5_df[[1]] %>%
  as_tibble %>%
  rename(increasing = group, group = facet) %>%
  filter(increasing == FALSE & x < 0 | increasing == TRUE & x > 0)

p <- ggplot(pred_test, aes(x = x , y = predicted, color = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha =
    .1) +
  geom_line() +
  geom_point(data = rename(test,
      predicted = linear_slope, x = bm_slope),
    aes(x = x, y = predicted),
    inherit.aes = FALSE)
```

```{r, eval = FALSE}
model_int$pred5_df[[1]] %>%
  ggplot()

plot_final_model(model_int$pred5[[1]], rawdata = model_int$data[[1]], facet =
  TRUE)
```

# Correlation

```{r}
cor_bm_net <- bm_net_group_median %>%
  select(bm_slope, variable, linear_slope, bm, width_river_mean) %>%
  pivot_wider(names_from = variable, values_from = linear_slope) 
corrplot::corrplot(cor(na.omit(cor_bm_net)), type = "upper")

library(ade4)
library(factoextra)
res.pca <- dudi.pca(na.omit(cor_bm_net),
                    scannf = FALSE,   # Cacher le scree plot
                    nf = 3            # Nombre d'axes gardés
                    )
screeplot(res.pca, main = "Screeplot - Eigenvalues")
p1 <- fviz_pca_var(res.pca,
  col.var = "contrib", 
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
  repel = TRUE     
)
p2 <- fviz_pca_var(res.pca,
  axes = c(1, 3),
  col.var = "contrib", 
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
  repel = TRUE     
)

p3 <- fviz_pca_var(res.pca,
  axes = c(2, 3),
  col.var = "contrib", 
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
  repel = TRUE     
)
plot_grid(p1, p2, p3)
```

## Richness

```{r}
knitr::opts_chunk$set(
  eval = FALSE 
)
```

```{r}
st_big_bm_tendancies <- group_results %>%
  filter(abs(linear_slope) > 5000) %>%
  .$station %>%
  unique
plot_big_bm_tendancies <- temporal_dynamics_plot %>%
  filter(variable == "biomass", station %in% st_big_bm_tendancies) %>%
  .$plot
plot_grid(
  plotlist = map(plot_big_bm_tendancies, 
    ~.x + lims(x = c(0, 10), y = c(0, 155000)))
)
```

```{r}
rigal_shape_var <- rigal_classification %>%
  unnest(classif) %>%
  select(station, variable, shape_class) %>%
  mutate(station = as.character(station))

rigal_shape_var %>%
  filter(station %in% st_decrease_increase_rigal_biomass$station) %>%
  ggplot(aes(x = shape_class)) +
  geom_bar() +
  facet_wrap(~ variable) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## SAX

```{r, eval = FALSE}
library(jmotif)
loadd(full_data2)
sax_all_variable <- full_data2 %>%
  select(station, nb_year, biomass:log_bm) %>%
  pivot_longer(cols = biomass:log_bm, names_to = "variable", values_to = "value") %>%
  group_by(station, variable) %>%
  arrange(nb_year) %>%
  nest() %>%
  mutate(
    zts = map(data, function(x) znorm(x$value, threshold = 0.01)),
    paa = map(zts, ~paa(., 3)),
    sax = map_chr(paa, ~series_to_string(., 3))
  ) %>%
  ungroup()

# CBA:
sax_all_variable %>%
  filter(station %in% cba_rigal_sax$station, !variable %in% c("biomass", "nbnode")) %>%
  ggplot(aes(x = sax)) +
  geom_bar() +
  facet_wrap(facets = ~variable)

sax_all_variable %>%
  filter(
    station %in% cba_rigal_sax$station,
    !variable %in% c("biomass", "nbnode")
    ) %>%
  ggplot(aes(x = sax)) +
  geom_bar() +
  facet_wrap(facets = ~variable)
```

```{r}
loadd(hyp_table)
hyp_table <- make_hyp_table(hyp_table = effect_quad2_piece, mod = model_log_bm)
hyp_table %>%
  kable(., booktabs = TRUE,
    caption = 
      paste0(
      "Comparison of the two Statistical model for the median biomass model"
    )
  ) %>%
  collapse_rows(columns = 1, valign = "top")
```

### Compare Quad2 and piecewiseregression

```{r}
loadd(model_log_bm, summary_model_log_bm,
  summary_log_bm, vif_model_log_bm)
```

```{r}
loadd(effect_quad2_piece)
effect_quad2_piece %>%
  kable(., booktabs = TRUE,
    caption = paste0(
      "Hypothesis and matching coefficients."))
```

```{r}
loadd(hyp_table_f3y)
hyp_table_f3y <- make_hyp_table(hyp_table = effect_quad2_piece, mod = model_log_bm_f3y)

hyp_table_f3y %>%
  kable(., booktabs = TRUE,
    caption = 
      paste0(
      "Comparison of the two Statistical model for first three year"
    )
  ) %>%
  collapse_rows(columns = 1, valign = "top")
```

## Raw bm

```{r, eval = TRUE}
loadd(model_bm, model_bm_vif)

model_bm_vif %>%
  ggplot(aes(y = vif, x = term)) +
  geom_hline(yintercept = 10) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(model_type~variable, scales = "free")
```

```{r, eval = FALSE}
library(yhat)
test <- model_int$mod3[[1]]
CCData <- commonalityCoefficients(model_int$data[[1]],
  "linear_slope",
  list("bm_slope", "inc_f", "bm"), "F")
CCData

test_data <- model_int$data[[1]]
lm_test <- lm(
  formula = linear_slope ~ bm_slope*inc_f*bm,
  weights = reg_weight
)
test <- regr(lm_test)
```

### Effect of protocol type on the model validity

```{r}
loadd(model_bm_by_protocol, model_bm_by_protocol_vif)

protocol_vec <- model_bm_by_protocol_vif$protocol_type %>% unique
p_vif_protocol <- map(protocol_vec, function (x) {
  model_bm_by_protocol_vif %>%
    filter(protocol_type == x) %>%
    ggplot(aes(y= vif, x = term)) +
    geom_hline(yintercept = 10) +
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = paste0("Protocol type: ", x)) +
    facet_grid(model_type~variable, scales = "free_y")
  })
p_vif_protocol[[1]]
p_vif_protocol[[2]]
```

```{r, eval = FALSE}
filter(model_bm_by_protocol_vif, model_type == "mod_bm_quad")
```

```{r}
loadd(bm_net_group)
bm_net_group %>%
  filter(variable == "connectance") %>%
  ggplot(aes(x = bm_slope, y = bm)) +
  geom_point()
```

### Current best model

$$
Y = \alpha + \beta_1D_b + \beta_2D_bI_b + \beta_3D_bI_bB
$$

```{r, eval = TRUE}
loadd(model_bm_by_protocol_summary)
model_bm_by_protocol_summary %<>%
  filter(protocol_type == "complete")
```

```{r, eval = FALSE}
filter(model_bm_by_protocol_summary, model == "mod_inc_bm")[["resume"]]
filter(model_bm_by_protocol_summary, model == "mod_inc_bm")[["anova"]]

filter(model_bm_by_protocol_summary, model == "mod_bm_inc")[["resume"]]
filter(model_bm_by_protocol_summary, model == "mod_bm_inc")[["anova"]]

filter(model_bm_by_protocol_summary, model == "mod_bm_quad")[["resume"]]
filter(model_bm_by_protocol_summary, model == "mod_bm_quad")[["anova"]]

filter(model_bm_by_protocol_summary, model == "mod_bm_quad2")[["resume"]]
filter(model_bm_by_protocol_summary, model == "mod_bm_quad2")[["anova"]]
```

```{r, eval = TRUE}
pred <- model_bm_by_protocol_summary %>%
  filter(model == "mod_bm_inc") %>%
  mutate(
    pred = map(model_obj, ~ggpredict(.x,
	terms = c("bm_slope [-8, -6, -4,-3, -2, -1, -.5, -.01, .01, .5, 1, 2, 3, 4, 6]", "bm [quart2]", "inc_f"))),
    pred_df = map(pred, as_tibble),
    plot = map(pred, ~plot(.x, add.data = FALSE)),
    plot_data = map(pred, ~plot(.x, add.data = TRUE)),
    #myplot = map2(pred, data, ~plot_final_model(ggpred = .x, rawdata = .y))
  )
plot_grid(plotlist = pred$plot, labels = pred$variable)
```

```{r}
pred <- model_bm_by_protocol_summary %>%
  filter(model == "mod_bm_quad2") %>%
  mutate(
    pred = map(model_obj, ~ggpredict(.x,
	terms = c("bm_slope [-8, -6, -4,-3, -2, -1, -.5, -.01, .01, .5, 1, 2, 3, 4, 6]", "bm [quart2]"))),
    pred_df = map(pred, as_tibble),
    plot = map(pred, ~plot(.x, add.data = FALSE)),
    plot_data = map(pred, ~plot(.x, add.data = TRUE)),
    #myplot = map2(pred, data, ~plot_final_model(ggpred = .x, rawdata = .y))
  )
plot_grid(plotlist = pred$plot_data, labels = pred$variable)
```

#### All protocol

```{r}
loadd(model_bm_summary)
```

```{r, eval = FALSE}
filter(model_bm_summary, model == "mod_inc_bm")[["resume"]]
filter(model_bm_summary, model == "mod_inc_bm")[["anova"]]

filter(model_bm_summary, model == "mod_bm_inc")[["resume"]]
filter(model_bm_summary, model == "mod_bm_inc")[["anova"]]

filter(model_bm_summary, model == "mod_bm_quad")[["resume"]]
filter(model_bm_summary, model == "mod_bm_quad")[["anova"]]

filter(model_bm_summary, model == "mod_bm_quad2")[["resume"]]
filter(model_bm_summary, model == "mod_bm_quad2")[["anova"]]
```

```{r}
pred_bm_quad2 <- model_bm_summary %>%
  filter(model == "mod_bm_quad2") %>%
  mutate(
    pred = map(model_obj, ~ggpredict(.x,
	terms = c("bm_slope", "bm [quart2]"))),
    pred_df = map(pred, as_tibble),
    plot = map(pred, ~plot(.x, add.data = FALSE)),
    plot_data = map(pred, ~plot(.x, add.data = TRUE)),
    #myplot = map2(pred, data, ~plot_final_model(ggpred = .x, rawdata = .y))
  )
plot_grid(plotlist = pred_bm_quad2$plot, labels = pred_bm_quad2$variable)
pred_bm_quad2_plot <- map2(model_bm$data, pred_bm_quad2$pred_df, plot_pred_data)

plot_grid(plotlist = pred_bm_quad2_plot, labels = pred_bm_quad2$variable)

pred_bm_quad <- model_bm_summary %>%
  filter(model == "mod_bm_quad") %>%
  mutate(
    pred = map(model_obj, ~ggpredict(.x,
	terms = c("bm_slope", "bm [quart2]"))),
    pred_df = map(pred, as_tibble),
    plot = map(pred, ~plot(.x, add.data = FALSE)),
    plot_data = map(pred, ~plot(.x, add.data = TRUE)),
    #myplot = map2(pred, data, ~plot_final_model(ggpred = .x, rawdata = .y))
  )

pred_bm_quad_plot <- map2(model_bm$data, pred_bm_quad$pred_df, plot_pred_data)
plot_grid(plotlist = pred_bm_quad_plot, labels = pred_bm_quad2$variable)
```

#### Quadratic two

```{r}
#model_bm$data[[1]] %>%
  #ggplot(aes(x = bm_slope, y = linear_slope, fill = log(bm)), shape = 23, size = 4) +
  #scale_fill_viridis() +
  #geom_point() +
  #geom_line(
    #data = pred_bm_quad2$pred_df[[1]] %>%
      #rename(bm_f = group, bm_slope = x, linear_slope = predicted),
    #aes(y = linear_slope, x = bm_slope, color = bm_f),
    #inherit.aes = FALSE)

test <- map2(model_bm$data, pred_bm_quad2$pred_df, function (.data, pred) {

  .data %>%
    ggplot(aes(x = bm_slope, y = linear_slope, fill = log(bm)), shape = 23, size = 4) +
    scale_fill_viridis() +
    geom_point() +
    geom_line(
      data = pred %>%
	rename(bm_f = group, bm_slope = x, linear_slope = predicted),
      aes(y = linear_slope, x = bm_slope, color = bm_f),
      inherit.aes = FALSE)

    })
plot_grid(plotlist = test)
```

#### Quadratic one

```{r}
test <- map2(model_bm$data, pred_bm_quad$pred_df, function (.data, pred) {

  .data %>%
    ggplot(aes(x = bm_slope, y = linear_slope, fill = log(bm)), shape = 23, size = 4) +
    scale_fill_viridis() +
    geom_point() +
    geom_line(
      data = pred %>%
	rename(bm_f = group, bm_slope = x, linear_slope = predicted),
      aes(y = linear_slope, x = bm_slope, color = bm_f),
      inherit.aes = FALSE)

    })
plot_grid(plotlist = test, labels = model_bm$variable)
```