### 2021-01-27 report.Rmd
New model prediction plot + outline introduction

---
title: "Report"
author: "Alain Danet"
self_contained: yes
date: "`r Sys.Date()`"
tables: yes
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    math: katex
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE 
)
```

```{r}
library(drake)
library(tidyverse)
library(magrittr)
library(cowplot)
library(viridis)
library(kableExtra)
library(sf)
source("R/misc.R") # Define your custom code as a bunch of functions.
source("R/variable_shortcut.R")
source_dir(get_mypath("R")) # Define your custom code as a bunch of functions.

mytheme <- theme_cowplot() +
background_grid()

theme_set(mytheme)
theme_get()
```

# Research questions

## Abstract

Determine how community structure will respond to ongoing change remains an ongoing
challenge. Studies reported strong temporal trends in community biomass, notably
strong decline in insects and vertebrate populations.

However, those studies says nothing about how empirical findings about how
different aspects of community structure such as species richness react
to a change in community biomass.

Ecological theory might provide insight on how community structure varies
according to biomass. Especially BEF predicts a positive link between species
richness and community biomass especially in grassland.

However, communities used in BEF experiments might not be relevant to get
insight about what happens in natural settings. Firstly there are statics, i.e.
not temporal. Secondly, community composition are set to randomly and thus do
not reflect real perturbations which do not affect communities randomly. 
(to be fair one could argue that random communities can mimic an average outcome
over a set of possible perturbations, as perturbation 1 would cause extinction of species
C, perturbation 2 for species A, etc...).

Overall, we miss empirical findings about the link between biomass and
community structure.

We investigated the link between the temporal dynamics of community biomass at one side and 
species richness, connectance and average trophic level at another side using a
1995-2018 standardized monitoring of fish communities. We questionned the shape
of the link between both temporal trends.  More specifically, we questionned if
(2) the link between foodweb structure and biomass dynamics vary between growing
and shrinking communities, and if the size of the community can modulate this
response.

We found significant positive relationships the dynamics of community biomass
and the dynamics of species richness, average trophic level and connectance but
a negative one for weighted connectance. We also found this relationship was
different depending on if the community biomass was growing or shrinking through
time for all variables, suggesting that assembly processes such as colonization
processes might.

# Introduction

1. Option1 Global change -> change of biomass 
   Option2 Ecological communities are dynamics + example

2. Ecological communities often describe by their richness, biomass

3. BEF

```{r b-bivar-cap}
bivar_cap <- paste0(
  "Links between temporal dynamics of biomass and temporal dynamics of
  metrics for the communities which have a median biomass across time inferior
  to the median of the station (little) and the other (big)"
)
```

# Results

- Species richness, Avg trophic level, connectance:
  - increases when biomass increases while decreases when biomass decreases
  - Avg trophic level: same for species richness  
  - connectance:  decreases when richness increases while increases when
    richness decreases (reverse sign compared to biomass)

- Pielou (eveness) :
  - Indiv: 
    - do not varies with biomass temporal trends
    - increases with increasing temporal trends of species richness
  - Biomass: 
    - increase when biomass decreases and decreases when biomass increases
    - decreases when species richness decreases and increase when species
      richness increases

- Connectance: positive relationship with avg trophic level temporal trends

# Discussion

## Summary of result

- There was a significant positive link between the dynamics of community biomass and the
  dynamics of community structure for the variables: species richness, average trophic level, connectance except for pielou of
  individuals and pielou of biomass which had respectively a no and a negative
  relationship.

- There was a significant positive link between the dynamics of species richness and the
  dynamics of community structure for the variables: average trophic level, connectance, pielou of
  individuals and pielou of biomass; except for connectance which had a negative
  relationship.

- There are a significant link between the dynamics of community biomass and the
  dynamics of community structure for all the variables expect pielou of
  biomass: species richness, average trophic level, connectance and pielou of
  individuals

- We found that the link between network temporal dynamics and biomass temporal
  dynamics **does not change** according if the community loses or gain biomass, for all
  variables.

## Answer to research questions

- Lost and gain of biomass or species richness resulted in same network structure dynamics,
  suggesting that the two processes are symmetric.
- The low effect of community size suggests that the architecture of the
  community in terms of species richness, number of piscivorous species, plays a
  role in its dynamics

## Interpretation

- Richness:
  - When biomass increases, species richness increases as well
    as when biomass decreases conjointly to species richness.

- Avg trophic level: 
  - Higest trophic level concentrates most of the biomass, so gain and loss are
    statistically more linked to them
  - Low and gain of biomass are linked to big individuals

    - The observation that the number of link and link density follow the same
      pattern confirm this interpretation.

- Connectance:
  - In little communities, few piscivorous species and so loss of biomass do not
    change the foodwebs that much because fish-resource links are quite redondant.
    - there is a positive correlation between nb of pisc node and biomass.
    - In bigger communities, the loss of the highest piscivorous nodes/species
      (see average trophic level), which have a lot of interactions, leads to a
      decrease of connectance over time.
  - Increase in biomass was weakly related to an increase in node number or species
    richness, so it is not surprising that connectance to not vary.

- Connectance (fish-fish interactions):
  - Globally positive relationship
  - when biomass increases: communities gains higher trophic level that might
    explain this relationship (but species richness increase as well)
  - When biomass decreases: communities

- Weighted connectance:
  - When biomass decreases in little communities, loss of highest trophic level
    and so less big predators and weighted connectance become more even
    - In big communities, the highest number can make the absence of such an effect 
  - When Biomass increases:
    - but species richness and number of nodes increases and gini coefficient
      does not move: so it is ninly due to the number of species
    - Gini decrease (more even link distribution) and number of species
      decrease also, so weighted_connectance should increase (it is what we see
      for low biomass communities). No idea for the difference between little
      and big communites.

The key to understand the difference between little and big communities is the
number of piscivorous nodes.

```{r}
loadd(predict_plot2)
pred_plot_signif_com <- predict_plot2 %>%
  filter(
    y %in% c(get_com_str_var(), "log_rich_std", "piel_nind", "piel_bm"),
    !y %in% c("bm_std", "rich_std"),
    x == "log_bm_std"
  )
legend_signif <- get_legend(
  pred_plot_signif_com$pred_plot[[1]] +
    theme(legend.position = "bottom")
)
pred_plot_signif_com %<>% 
  mutate(
  pred_plot = pmap(list(pred_plot, y, x),
    function(p, y, x) {
      # Replace label
      label <- str_replace_all(c(x,y), var_replacement(slope = TRUE)) 
      p +
	labs(x = label[1], y = label[2]) +
	theme(legend.position = "none")
    } )
  )

list_p <- pred_plot_signif_com$pred_plot
list_p[[length(list_p) + 1]] <- legend_signif

pred_plot1 <- plot_grid(
  plotlist = list_p,
  labels = "AUTO",
  nrow = 2
)
#pred_plot1 <- plot_grid(top, legend_signif, nrow = 2, rel_heights = c(1, .1))
```

```{r}
pred_plot_rich <- predict_plot2 %>%
  filter( y %in% c(get_com_str_var(), "log_bm_std", "piel_nind", "piel_bm"), x == "log_rich_std") %>%
  filter(!y %in% c("rich_std"))
legend_signif <- get_legend(
  pred_plot_rich$pred_plot[[1]] +
    theme(legend.position = "bottom")
)
pred_plot_rich %<>% 
  mutate(
  pred_plot = pmap(list(pred_plot, y, x),
    function(p, y, x) {
      # Replace label
      label <- str_replace_all(c(x,y), var_replacement(slope = TRUE)) 
      p +
	labs(x = label[1], y = label[2]) +
	theme(legend.position = "none")
    } )
  )

list_p <- pred_plot_rich$pred_plot
list_p[[length(list_p) + 1]] <- legend_signif

pred_plot2 <- plot_grid(
  plotlist = list_p,
  nrow = 2,
  labels = "AUTO" 
)
#pred_plot2 <- plot_grid(top, legend_signif, nrow = 2, rel_heights = c(1, .1))
```

```{r}

pred_plot_other <- predict_plot2 %>%
  filter( y %in% c("ct_ff", "w_trph_lvl_avg"), x %in% c("ct_ff", "w_trph_lvl_avg"))
legend_signif <- get_legend(
  pred_plot_other$pred_plot[[1]] +
    theme(legend.position = "bottom")
)
pred_plot_other %<>% 
  mutate(
  pred_plot = pmap(list(pred_plot, y, x),
    function(p, y, x) {
      # Replace label
      label <- str_replace_all(c(x,y), var_replacement(slope = TRUE)) 
      p +
	labs(x = label[1], y = label[2]) +
	theme(legend.position = "none")
    } )
  )

list_p <- pred_plot_other$pred_plot
list_p[[length(list_p) + 1]] <- legend_signif

pred_plot3 <- plot_grid(
  plotlist = list_p,
  labels = "AUTO", nrow = 2
)
#pred_plot3 <- plot_grid(top, legend_signif, nrow = 2, rel_heights = c(1, .1))
```

```{r, fig.dim = c(10, 7)}
pred_plot1
pred_plot2
pred_plot3
```

```{r, eval = FALSE, fig.dim = c(12, 12)}
plot_grid(
  , plot_grid(, , ncol = 2, rel_widths = c(.75, 1)),
  nrow = 2
)
```

```{r, eval = FALSE, fig.dim = c(10, 5)}
loadd(betapart_bin, rigal_classification, st_mono_trends_combined_list)

mask <- st_mono_trends_combined_list$variable == "log_bm_std"
xdata <- rigal_classification %>%
  filter(variable == "log_bm_std") %>%
  unnest(classif) %>%
  filter(station %in% st_mono_trends_combined_list[mask, ]$station[[1]])

beta_xdata <- xdata %>%
  select(station, linear_slope) %>%
  left_join(betapart_bin, by = "station")

beta_xdata %>%
  pivot_longer(cols = c(beta.SIM:beta.SOR), names_to = "type", values_to = "beta") %>%
  ggplot(aes(x = linear_slope, y = beta)) +
  geom_point() +
  facet_grid(cols = vars(type)) +
  labs(x = "Slope of Log biomass (% biomass by year)", y = "Beta-diversity")

```

```{r}
#loadd(temporal_dynamics_plot)
```

```{r}
loadd(model_log_bm_f3y_table_medium)
model_log_bm_f3y_table_medium[["anova_table"]] %>%
  filter(
    y %in% c(get_com_str_var(), "ct_ff", "log_rich_std"),
    !y %in% "rich_std") %>%
  kable(., booktabs = TRUE,
    caption = 
      paste0(
      "ANOVA table for the model: dN =",
      "dB + B + dB*B + Inc*B + dB*Inc*B. The community structure versus the
      biomass"
    )
  ) %>%
  kable_styling(latex_options =c("scale_down", "stripped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

```{r}
loadd(model_log_rich_f3y_table_medium)
model_log_rich_f3y_table_medium[["anova_table"]] %>%
  filter(
    y %in% c(get_com_str_var(), "log_bm_std"),
    !y %in% "rich_std") %>%
  kable(., booktabs = TRUE,
    caption = 
      paste0(
      "ANOVA table for the model: dN =",
      "dB + B + dB*B + Inc*B + dB*Inc*B. The community structure versus the
      species richness"
    )
  ) %>%
  kable_styling(latex_options =c("scale_down", "stripped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

### Further analysis

```{r}
loadd(summary_var_f3y)

summary_var_f3y %>%
  na.omit() %>%
  select(bm_std, avg_IS, l, ld, m, A, gini, log_bm_std, connectance, ct_ff, rich_std, w_trph_lvl_avg,
    weighted_connectance, nbnode_std, nb_pisc_node_std, prop_pisc_node, piel_bm,
    piel_nind) %>%
  cor() %>%
  corrplot::corrplot(corr = .)
```

## Other variables

```{r pred-other-com, fig.dim = c(12, 10)}
pred_plot_signif <- predict_plot2 %>%
  filter(!y %in% c(get_com_str_var(), "ct_ff"), x == "log_bm_std")
legend_signif2 <- get_legend(
  pred_plot_signif_com$pred_plot[[1]] + theme(legend.position = "bottom")
)
pred_plot_signif$pred_plot %<>% map(., ~.x + theme(legend.position = "none"))
top2 <- plot_grid(
  plotlist = pred_plot_signif$pred_plot,
  labels = pred_plot_signif$y
)
plot_grid(top2, legend_signif2, nrow = 2, rel_heights = c(1, .1))
```

```{r}
model_log_bm_f3y_table_medium[["anova_table"]] %>% 
  filter(!y %in% get_com_str_var(all = FALSE)) %>%
  kable(., booktabs = TRUE,
    caption = 
      paste0(
      "ANOVA table for the model: dN =",
      "dB + B + dB*B + Inc*B + dB*Inc*B."
    )
  ) %>%
  kable_styling(latex_options =c("scale_down", "stripped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

```{r}
# Rich and rich slope
loadd(summary_var_f3y, st_mono_trends_combined_list, rigal_classification)

mask <- st_mono_trends_combined_list$variable == "richness"
xdata <- rigal_classification %>%
  filter(variable == "richness") %>%
  unnest(classif) %>%
  filter(station %in% st_mono_trends_combined_list[mask, ]$station[[1]])

xdata %>%
  select(station, linear_slope) %>%
  left_join(summary_var_f3y %>% select(station, richness), by = "station") %>%
  ggplot(aes(x = linear_slope, y = richness)) +
  geom_point() +
  labs(x = "Slope of richness (year)", y = "richness")
```

```{r setup,  eval = TRUE}
knitr::opts_chunk$set(eval=FALSE)
```

## Temporal series of networks

```{r}
st_id_plot <- log_bm_net_group_f3y %>%
  filter(variable %in% get_com_str_var()) %>%
  ggplot(aes(x = bm_slope, y = linear_slope, color = log(bm))) +
  scale_color_viridis() +
  geom_text(aes(label = station)) +
  facet_wrap(~variable, scales = "free_y")
st_id_plot
st_network_to_plot <- c(513, 1453, 16654, 6020)
```

- Station ploted: `r #st_network_to_plot`

```{r}
library(tidygraph)
library(ggraph)
#source(mypath("R", "plot_methods.R"))
myload(network_analysis,
  envir = environment(),
  dir = get_mypath("data", "classes"))
myload(op_analysis,
  metaweb_analysis,
  envir = environment(),
  dir = get_mypath("data"))
meta <- metaweb_analysis; rm(metaweb_analysis)
st_sp_node <- network_analysis %>%
  select(opcod, composition) %>%
  left_join(dplyr::select(op_analysis, opcod, station, date), by = "opcod") %>%
  unnest(cols = c(composition)) %>%
  group_by(station) %>%
  summarise(
    nbnode = length(unique(sp_class)),
    nbsp = length(unique(str_extract(sp_class, "[A-Z]+")))
  ) %>%
  arrange(nbsp)

net <- network_analysis %>%
  dplyr::select(opcod, network) %>%
  unnest(cols = c(network)) %>%
  left_join(., dplyr::select(op_analysis, opcod, station, date)) %>%
  mutate(date = lubridate::year(date))

myload(species_color_temporal_plot,
  envir = environment(),
  dir = get_mypath("data"))
```

```{r}
#st_id_plot
temporal_network <- map(st_network_to_plot, function (st) {
  net_ex <- filter(net, station == st)
  p <- my_crap_temporal_network(
    net = net_ex,
    metaweb = meta$metaweb,
    network_data = network_analysis,
    nrow_sp_legend = 2,
    return_data = TRUE,
    my_y_lim = c(0, 4.1),
    bm_var = "bm_std",
    color_node = species_color_temporal_plot
  )
  plot_grid(plotlist = p$plots)
  })
temporal_network[[1]]
temporal_network[[2]]
temporal_network[[3]]
temporal_network[[4]]
```

```{r}
loadd(rigal_species)
species_trends <- rigal_species %>%
  unnest(classif) %>%
  select(station, species = variable, linear_slope, linear_slope_strd_error, linear_slope_pvalue)

species_trends2 <- species_trends %>%
  left_join(filter(log_bm_net_group_f3y, variable == "connectance") %>%
    select(station, bm_slope),
    by = "station")
species_trends2 %<>% na.omit

species_trends2 %>%
  filter(linear_slope_pvalue < 0.05) %>%
  ggplot(aes(y = linear_slope, x = bm_slope, color = species)) +
  geom_point() +
  geom_smooth(method = "lm")

species_trend_cor <-
  species_trends2 %>%
  filter(linear_slope_pvalue < 0.05) %>%
  group_by(species) %>%
  summarise(
    spear = cor(linear_slope, bm_slope, method = c("spearman"), use = "pairwise.complete.obs"),
    n = n()
  )

species_trend_cor %>%
  filter(n >= 5) %>%
  ggplot(aes(y = spear, x = species)) +
  geom_bar(stat = "identity") +
  geom_label(aes(y = 0, label = paste("N=", n)), vjust = 1) +
  theme(axis.text.x = element_text(angle = 45))
```

```{r}
# Obtain avg troph lvl by species
dead_material <- c("det", "biof")
tlvl <- NetIndices::TrophInd(meta$metaweb, Dead = dead_material)$TL
names(tlvl) <- colnames(meta$metaweb) 
tlvl <- enframe(tlvl, name = "sp_class", value = "tlvl")

species_tlvl <- 
  tlvl %>%
  filter(str_detect(sp_class, "[A-Z]{3}.")) %>%
  mutate(species = str_sub(sp_class, 1, 3)) %>%
  group_by(species) %>%
  summarise(tlvl = mean(tlvl, na.rm = TRUE))
```

```{r}
species_trend_cor %>%
  left_join(species_tlvl, by = "species") %>%
  filter(n >= 5) %>%
  ggplot(aes(x = tlvl, y = spear)) +
  geom_label(aes(label = species)) +
  geom_smooth(method = "lm")
```

## Composition and biomass trajectory

```{r}
library(vegan)
myload(com_mat, envir = environment(), dir = get_mypath("data"))

bm_slope_df <- log_bm_net_group_f3y %>% 
  filter(variable == "connectance") %>%
  select(station, bm_slope, bm) %>%
  arrange(station)

loadd(com_analysis_data, op_data)
com_mat <- compute_com_mat(
  .op = ungroup(op_data) %>% filter(station %in% bm_slope_df$station),
  com = ungroup(com_analysis_data))

com <- com_mat %>% 
  arrange(station) 
#check:
all(com$station == bm_slope_df$station)
com <- com[, colSums(com != 0) > 0] %>%
  select(-station)

```

```{r, cache = TRUE}
nmds <- metaMDS(com, k = 2, trymax = 100)
```

```{r, fig.width = 10, fig.height = 8}
diverging_pal <- c("#67001f", "#b2182b", "#d6604d", "#f4a582", "#fddbc7",
  "#f7f7f7", "#d1e5f0", "#92c5de", "#4393c3", "#2166ac", "#053061")

par(mfrow = c(1, 2))
stressplot(nmds)
title("Stress plot")
bluefunc <- colorRampPalette(c("lightyellow", "darkred"))
ordiplot(nmds,type="n")
orditorp(nmds,display="species",col="gray",air=0.01)
orditorp(nmds, display="sites",
  col= diverging_pal[findInterval(bm_slope_df$bm_slope, seq(-0.2,.2, length.out = length(diverging_pal)))],
   cex = findInterval(bm_slope_df$bm, seq(1, 4, 1))/2,
  air=0.01, pch = "+", labels = "+")
title("Color = ; Size of points = biomass")
```

## Species richness dynamics and network dynamics

```{r}
loadd(model_log_rich_f3y_table_medium)
```

```{r}
model_log_rich_f3y_table_medium[["anova_table"]] %>% 
  filter(variable %in% get_com_str_var(all = FALSE)) %>%
  kable(., booktabs = TRUE,
    caption = 
      paste0(
      "ANOVA table for the model: dN =",
      "dB + B + dB*B + Inc*B + dB*Inc*B."
    )
  ) %>%
  kable_styling(latex_options =c("scale_down", "stripped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

```{r}
loadd(pred_plot_signif_log_rich)
pred_plot_signif_log_rich_com <- pred_plot_signif_log_rich %>%
  filter(variable %in% get_com_str_var())
legend_signif <- get_legend(
  pred_plot_signif_log_rich_com$myplot[[1]] +
    theme(legend.position = "bottom")
)
pred_plot_signif_log_rich_com$myplot %<>% map(., ~.x + theme(legend.position = "none"))
top <- plot_grid(
  plotlist = pred_plot_signif_log_rich_com$myplot,
  labels = pred_plot_signif_log_rich_com$variable
)
plot_grid(top, legend_signif, nrow = 2, rel_heights = c(1, .1))
```

## Link biomass dynamic and environment

```{r}
loadd(habitat_pressure, region_polygon, station_analysis)
loadd(richness_group)
loadd(log_bm_net_group_f3y)  
test <- log_bm_net_group_f3y %>%
  filter(variable == "connectance") %>%
  left_join(habitat_pressure, by = "station")
test %>%
  pivot_longer(
    cols = c(bm, width_river_mean, temperature_med, alt , RC1, RC2),
    names_to = "evt_var", values_to = "value") %>%
  ggplot(aes(y = bm_slope, x = value)) +
  geom_smooth(method = "lm") +
  geom_point() +
  facet_wrap(~evt_var, scales = "free_x")
```

```{r}
p <- ggplot() +
  geom_sf(data = st_geometry(region_polygon)) 
st_bm_slope <- station_analysis %>%
  rename(station = id) %>%
  left_join(select(test, station, bm_slope)) %>%
  filter(!is.na(bm_slope)) %>%
  mutate(trend = ifelse(bm_slope < 0, "negative", "positive"))
p_col_bm_slope <- p + 
  geom_sf(data = st_bm_slope, aes(color = bm_slope)) +
  scale_color_viridis()
p_trend_bm_trend <- p +
  geom_sf(data = st_bm_slope, aes(color = trend))
plot_grid(p_col_bm_slope, p_trend_bm_trend, nrow = 1)
```

```{r}
loadd(habitat_press)
hab_press <- habitat_press %>%
  select(station, long, lat, slope, alt, d_source, strahler, width_river_mean,
    avg_depth_station_mean, DBO_med, flow_med)
dyn_press <- log_bm_net_group_f3y %>%
  filter(variable == "connectance") %>%
  select(station, bm_slope) %>%
  left_join(hab_press, by = "station") %>%
  na.omit
corrplot::corrplot(cor(dyn_press[, -1], method = "spearman"))
```

## Richness and dynamics

```{r}
log_bm_net_group_f3y %>%
  filter(variable == "rich_std", bm_slope < 0) %>%
  left_join(habitat_pressure, by = "station") %>%
  pivot_longer(cols = c(bm, width_river_mean, temperature_med, alt , RC1, RC2),
    names_to = "evt_var", values_to = "value") %>%
ggplot(aes(y = linear_slope, x = value)) +
geom_smooth(method = "lm") +
geom_point() +
facet_wrap(~evt_var, scales = "free_x")
```

```{r}
rich_std_f3y <- richness_group %>% 
  filter(rich_var == "rich_std", summary_type == "first_3_year", group_type ==
    "median") %>%
  select(station, rich)

log_bm_net_group_f3y %>%
  filter(variable == "rich_std") %>%
  left_join(rich_std_f3y, by = "station") %>%
  ggplot(aes(y = linear_slope, x = rich)) +
  geom_smooth(method = "lm") +
  labs( x = "Species richness (by m2)",
    y = "Species richness trend") +
  geom_point()
```

##

## The two models

```{r, fig.height = 12, fig.width = 12}
loadd(p_pred_quad2_f3y)
legend_quad2 <- get_legend(p_pred_quad2_f3y[[1]])
p_pred_quad2_f3y %<>% map(., ~.x + theme(legend.position = "none"))
top <- plot_grid(
  plotlist = p_pred_quad2_f3y,
  labels = names(p_pred_quad2_f3y)
)
quad2_plot <- plot_grid(top, legend_quad2, rel_widths = c(1, .1))
```

```{r}
loadd(p_pred_medium_f3y)
p_pred_medium_f3y %<>% map(., ~.x + theme(legend.position = "none"))
medium_plot <- plot_grid(plotlist = p_pred_medium_f3y,
  labels = names(p_pred_medium_f3y))
```

```{r, fig.height = 12}
plot_grid(quad2_plot, medium_plot, nrow = 2, labels = c("poly", "piecereg"))
```

Finally, both models does not disagree that much, for the tendancies according
to the biomass.

```{r}
loadd(log_bm_net_group_f3y, model_log_bm_f3y,
  summary_log_bm_f3y, vif_log_bm_f3y)
loadd(vif_log_bm)
```

```{r}
loadd(p_bm_vs_bm_slope)
p_bm_vs_bm_slope
```

```{r}
loadd(p_hist_med_bm)
p_hist_med_bm
```

# Methods

- Methodological choice:
  - Dynamics of log of the biomass: put big and little station on the same scale
    (% of biomass gained or lost per year)

2. $dN = dB + B + dB \times B + Inc \times B + dB \times Inc \times B$

with:
- $dN$: temporal trend of network structure
- $dB$: temporal trend of total (log) biomass
- $B$: total median biomass or total biomass of the first three year
- $Inc$: Is $dB > 0$?

We kept the second model.