### 2021-01-30 paper/wordstack.Rmd
write results and wordstack/outline

---
title: "Wordstack and outline of the manuscript"
author: "Alain Danet"
self_contained: yes
date: "`r Sys.Date()`"
tables: yes
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    math: katex
---

# Wordstack

- Foodwebs
- interactions
- perturbation/perturbation sequences
- network structure
- temporal dynamics
- BEF
- space for time
- extinction sequences

- omnivory
- equilibrium / dynamics

# Outline

### Introduction

1. community structure are dynamics and affect ecosystem functions 
    1. species interactions ?
    2. biomass or richness

2. Community structure dimension are linked 
    1. Biomass and species richness 
	1. experiment: richness and productivity richness and stability
	2. Empirical:
    2. Foodweb/Network structure and biomass/richness 
	1. connectance richness
	2. persistance / connectance

3. Colonisation / Extinction sequences in natural settings
    1. Space for time substitution
    2. Communities are not affected randomly as environment species niche  rel
    3. sequence: is losing something = adding something ?

4. Research questions 
    1. Link between the temporal dynamics of different aspects of community structure ?
	1. Is BEF valid? 
	2. is space for time substitution?
    2. effect of assembly/disassembly

5. blabla Foodweb/fish/super huge top dataset

## Material and methods

1. Data selection
2. Community biomass and richness 
3. Network reconstruction
4. Statistical analysis

## Results

1. No effects of community assembly/disassembly

2. Community structure and biomass
    1. positive links everywhere except piel ind (neutral) and piel biomass
       (negative)
    2. No effect of differences between gain and losses

3. Community structure and species richness
   1. Positive links everywhere except connectance

## Discussion

1. Temporal trends valid theory: BEF and network assembly
    1. Biomass and species richness
    2. Connectance and species richness:
    3. Tlvl and species richness
2. Top trophic levels drives the response to biomass and species richness variation
    1. Top predators represent both few individuals and most of the biomass 
    2. Predators are key to the network structure
3. Biomass and richness gains and losses are symmetric processes in terms of
   community structure
    1. Because big individuals are drivers
    2. We could have imagine other ways, that losses and gain of biomass/species are
       overall due to little individuals. Is it a statistical or an ecological
       effect? because big individuals are less numerous? (can be check with
       nind in covariable?)
4. Are community dynamics fluctuate around equilibrium point