### 2021-07-16 draft.Rmd
first try of a method figure

# For Loubna

```{r}
library(drake)
library(tidyverse)

mypath <- rprojroot::find_package_root_file

source(mypath("R", "misc.R"))
source_dir(mypath("R"))
theme_set(theme_cowplot())
```

```{r}
get_bm_dyn_year <- function (com = NULL, station = NULL) {

  st <- station
  test <- com %>%
    filter(station == st)

  test %<>%
    mutate(sp_tibble = map(sp_vector, ~tibble(species = names(.x), biomass_sp = .x))) %>%
    unnest(sp_tibble) %>%
    mutate(bm_std_sp = biomass_sp / surface) %>%
    dplyr::select(station, year, species, bm_std_sp)

  test %<>% 
    pivot_wider(names_from = species, values_from = "bm_std_sp")%>%
    mutate_at(vars(matches("[A-Z]{3}", ignore.case = FALSE)), ~ifelse(is.na(.), 0, .)) %>%
    pivot_longer(cols = matches("[A-Z]{3}", ignore.case = FALSE), names_to = "species", values_to = "bm_std_sp")

  return(test)
}
get_network_summary <- function (com = NULL, station = NULL, stab = TRUE, com_avg = TRUE) {

  st <- station
  tmp <- com %>%
    filter(station == st) %>%
    mutate_if(is.double, ~round(., 2))

  label_stab <- c(
    "Stability" = tmp$bm_std_stab,
    "Synchrony" = tmp$sync, 
    "CVsp" = tmp$cv_sp,
    "Median total biomass" = tmp$prod_std #"' * (g/m^2) * '"
    )
  label_com <- c(
    "Median species richness" = tmp$richness_med,
    "Median connectance" = tmp$ct,
    "Median avg trophic level" = tmp$t_lvl
  )
  label <- NULL

  if (stab) {
    label <- c(label, label_stab)

  }

  if (com_avg) {

    label <- c(label, label_com)

  }

  return(label)

}

plot_dyn_web  <- function (
  com = NULL, sem_df = NULL, station = NULL,
  stacked = FALSE, .log = FALSE, color_species = NULL, label = NULL,
  label_parsed = FALSE, label_size = 4.5, y_label = NULL, my_ylim = NULL) {

  # get bm dynamic 
  com <- get_bm_dyn_year(com = com, station = station)

  # get total biomass 
    total_bm <- com %>%
	  group_by(year) %>%
	  summarise(bm_std_sp = sum(bm_std_sp)) %>%
	  mutate(species = "Total")

  p <- com %>%
    ggplot(aes(x = year, y = bm_std_sp, color = species))

  if (stacked) {
    p <- p + 
      geom_area(aes(fill = species))
  } else {
    p <- p + 
      geom_line() +
      geom_line(data = total_bm, color = "black")
  }

  if ( !is.null(my_ylim) ) 
  {
    p <- p +
      ylim(my_ylim)
  }

  # Make it professional:
  p <- p +
    labs(y = expression(Biomass~density~(g/m^2)), x = "Year")

  if (!is.null(sem_df)) {

    label <- get_network_summary(com = sem_df, station = station) 
  }

  if (!is.null(label)) {

    if (is.null(y_label)) {
      y_label <- max(total_bm$bm_std_sp) + max(total_bm$bm_std_sp) * 5/100
    }

    p <- p +
    annotate("text", x = median(total_bm$year),
      y = y_label,
      label = label, parse = label_parsed, size = label_size)
  }

  if (.log) {
    p <- p + scale_y_log10()
  }

  if (!is.null(color_species)) {
    p <- p + 
      scale_fill_manual(values = color_species) +
      scale_color_manual(values = color_species)
  }

  return(p)

}

get_label_from_named_vector <- function ( label = NULL, collapse = "\n") {
  paste0(names(label), " = ", label, collapse = collapse)
} 
```

```{r}
get_adj_net_station <- function (net = NULL, station = NULL) {

  st <- station

  net <- net %>%
    filter(station %in% st) %>%
    dplyr::select(station, year, network) %>%
    unnest(cols = c(network))

  ## Compute metrics:
  net %>%
    dplyr::select(year, from, to) %>%
    arrange(year) %>%
    group_by(year) %>%
    nest(network = c(from, to)) %>%
    mutate(
      igraph_obj = map(network, igraph::graph_from_data_frame,
	directed = TRUE),
      adj_mat = map(igraph_obj, igraph::as_adjacency_matrix,
	sparse = FALSE),
      troph = map(adj_mat, ~NetIndices::TrophInd(.x)),
      obs_troph_level_vector = map(troph, function (x) {
	TL <- x$TL
	names(TL) <- rownames(x) 
	return(TL)
	})
    )
}

get_net_biomass <- function (net = NULL, station = NULL) {

  st <- station

  # Get bm
  net <- net %>%
    filter(station %in% st) %>%
    dplyr::select(year, composition) %>%
    unnest(cols = c(composition))  %>%
    dplyr::select(year, sp_class, bm_std)

   # Make list
    net %<>%
      arrange(year) %>%
      group_by(year) %>%
      nest(bm = c(sp_class, bm_std)) %>%
      mutate(bm = map(bm, deframe))
  net
}

set_color_nodes <- function (
  node_list = NULL,
  color_species_resources = NULL
  ) {

  node_list_sp <- stringr::str_replace_all(node_list, "_\\d", "")

  colour <- color_species_resources[node_list_sp]
  names(colour) <- node_list
  return(colour)
}

set_node_resources_abun <- function (
  node_list = NULL,
  node_abun = NULL, resource_abun = NULL) {

  abun <- vector("numeric", length(node_list))  
  names(abun) <- node_list

  if (is.null(resource_abun)) {
    resource_abun <- mean(node_abun)
  }

  abun[names(node_abun)] <- node_abun 
  abun[abun == 0] <- resource_abun 
  return(abun)

}

get_network_temporal_plot <- function (
  net_list = NULL,
  station = NULL,
  selected_years = NULL,
  node_color = NULL,
  scale_node_size = 0.01,
  par_options = par(mar = c(0,0,4,0)),
  add_title = FALSE,
  rel_abun = FALSE,
  log_abun = FALSE
  ) {

  net <- get_adj_net_station(net = net_list, station = station)
  bm <- get_net_biomass(net = net_list, station = station)

  if (!is.null(selected_years)) {
    net %<>% filter(year %in% selected_years) 
    bm %<>% filter(year %in% selected_years) 
  }

  if (is.null(node_color)) {
    node_color <- set_color_species(
      node_list = unique(names(unlist(net$obs_troph_level_vector))),
      species_list = NULL, resource_list = NULL, col_resource = NULL)
  }

  list_p <- list()
  index <- seq_along(net$year)

  for (i in index) {
    if(!is.null(par_options)) {
      par_options
    }
    plot.new()
    PlotWeb(
      TL = net$obs_troph_level_vector[[index[i]]],
      webTL = net$adj_mat[[index[i]]],
      colnode = set_color_nodes(
	node_list = names(net$obs_troph_level_vector[[index[i]]]),
	color_species_resources = node_color),
      abund = set_node_resources_abun(
	node_list = names(net$obs_troph_level_vector[[index[i]]]),
	node_abun = bm$bm[[index[i]]],
	resource_abun = 1
	),
      collink="grey70",
      scale_abun = scale_node_size,
      rel_abun = rel_abun,
      log_abun = log_abun
    )
    if (add_title) {
     title(main = paste0(net$year[[index[i]]]))
    }

    p <- recordPlot()

    list_p[[i]] <- p 
  }

  gg <- map(list_p, ggdraw)
  gg
}
```

```{r}
loadd(rigal_classification, com_data, op_data)
op_data %<>%
  mutate(year = as.integer(year))
myload(community_metrics, dir =
  "~/Documents/post-these/mnhn/fishcom_biomass_dynamics/data", envir =
  environment())

community_metrics %<>%
    left_join(dplyr::select(op_data, station, opcod, year))
ti <- rigal_classification %>%
  filter(variable == "bm_std") %>%
  unnest(classif)
first_st <- ti %>%
  filter(shape_class == "decrease_constant") %>%
  arrange(linear_slope)

tes <- map(first_st$station[1:12], ~plot_dyn_web(
    com = community_metrics,
    sem_df = NULL,
    station = .x,
#    .log = TRUE,
    stacked = TRUE
    ))
plot_grid(plotlist = tes)
```
```{r}

last_st <- ti %>%
  filter(shape_class == "increase_constant") %>%
  arrange(linear_slope)

tes <- map(last_st$station[1:12], ~plot_dyn_web(
    com = community_metrics,
    sem_df = NULL,
    station = .x,
#    .log = TRUE,
    stacked = TRUE
    ))
plot_grid(plotlist = tes)
```

```{r}
st <- first_st$station[4]
st[2] <- last_st$station[13]
names(st) <- c("decreasing", "increasing") 
#11432 my choice
```

```{r}
myload(network_analysis, dir =
  "~/Documents/post-these/mnhn/fishcom_biomass_dynamics/data/classes", envir =
  environment())
network_analysis %<>%
  left_join(dplyr::select(op_data, opcod, station, year)) %>%
  filter(!is.na(opcod))

p <- get_network_temporal_plot(
  net_list = network_analysis,
  station = st[1],
  selected_years = NULL,
  node_color = NULL, scale_node_size = 0.06,
  par_options = par(mar = c(0,0,0,0)),
  log_abun = TRUE
)
g <- plot_grid(plotlist = map(p, ggdraw), nrow = 3)

save_plot(
  "figures/decreasing_biomass_network.png", g,
  nrow = 3,
  ncol = 4,
  base_height = 4,
  base_width = 4)

p2 <- get_network_temporal_plot(
  net_list = network_analysis,
  station = st[2],
  selected_years = NULL,
  node_color = NULL, scale_node_size = 0.06,
  par_options = par(mar = c(0,0,0,0)),
  log_abun = TRUE
)
g2 <- plot_grid(plotlist = map(p2, ggdraw), nrow = 4)

save_plot(
  "figures/increasing_biomass_network.png", g2,
  nrow = 4,
  ncol = 4,
  base_height = 4,
  base_width = 4)
```

```{r}
library(lubridate)

tes <- plot_dyn_web(
    com = mutate(community_metrics, year = ymd(year, truncated = 2L)),
    sem_df = NULL,
    station = st[1],
#    .log = TRUE,
    stacked = TRUE
    )

tps_bm_dec  <- tes +
  theme(legend.position = "none")

save_plot(
  "figures/decreasing_biomass.png", tps_bm_dec,
  nrow = 1,
  ncol = 1,
  base_height = 4,
  base_width = 7)

tes2 <- plot_dyn_web(
    com = mutate(community_metrics, year = ymd(year, truncated = 2L)),
    sem_df = NULL,
    station = st[2],
#    .log = TRUE,
    stacked = TRUE
    )

tps_bm_inc  <- tes2 +
  theme(legend.position = "none")

save_plot(
  "figures/increasing_biomass.png", tps_bm_inc,
  nrow = 1,
  ncol = 1,
  base_height = 4,
  base_width = 7)
```

# Plot temporal dynamics

```{r}
loadd(full_data2, rigal_classification)
```
```{r}
rig_st_dec <- rigal_classification %>%
  filter(variable == "ct_ff") %>%
  unnest(classif) %>%
  filter(station == st[1])
```

```{r}
tps_ct_dec <- full_data2 %>%
  mutate(year = ymd(year, truncated = 2L)) %>%
  filter(station == st[1]) %>%
  ggplot(aes(x = year, y = ct_ff)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x,
    size = 1, se = FALSE, color = "black") +
  labs(x = "Year", y = "Connectance")

plot_temporal_variable_lm <- function(
  .df = full_data2,
  var = "ct_ff",
  st = 15818) {
  .df %>%
    filter(station %in% st) %>%
    mutate(
      year = ymd(year, truncated = 2L),
      station = str_replace_all(station, setNames(names(st), st))
      ) %>%
    ggplot(aes_string(x = "year", y = var)) +
    geom_point() +
    geom_smooth(aes(group = station), method = "lm", formula = y ~ x,
      size = 1, se = FALSE, color = "black") +
    labs(x = "Year", y = str_replace_all(var, var_replacement()))

}

tps_dec_inc <- map(c("ct_ff", "w_trph_lvl_avg", "log_rich_std", "piel_bm"),
  ~plot_temporal_variable_lm(st = st, var = .x) +
  facet_grid(cols = vars(station))
)

plot_grid(plotlist = tps_dec_inc)
```

```{r}
loadd(map_st)

test_fig_mat_met <- plot_grid(
  map_st,
  plot_grid( plot_grid(tps_bm_dec, tps_bm_inc)), 
  plot_grid(plotlist = tps_dec_inc),
  nrow = 3
)
save_plot(test_fig_mat_met,
  file = get_mypath("figures", "matmet.png") 
)
```