### 2022-03-11 sem.Rmd
Add SEM

---
title: "SEM"
author: AD 
date: \today
output:
  bookdown::html_document2:
    toc: true
    theme: "readable"
    code_folding: hide
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  message = FALSE,
  warning = FALSE,
  results = TRUE
)
```

```{r}
loadd(slope_com_var_no_covar, st_mono_trends_rich_bm)
library(piecewiseSEM)
library(semEff)
library(kableExtra)
library(lme4)
mypath <- rprojroot::find_package_root_file
source(mypath("R/misc.R"))
source(mypath("R/variable_shortcut.R")) # Define your custom code as a bunch of functions.
source_dir(get_mypath("R")) # Define your custom code as a bunch of functions.

mytheme <- theme_cowplot() +
background_grid()

theme_set(mytheme)
```

```{r}
loadd(sp_st_data)
loadd(slope_com_var_no_covar)
loadd(st_mono_trends_stable_rich_bm)
```

```{r}
var_analysis <- c("ct_ff", "log_bm_std", "log_rich_std", "w_trph_lvl_avg")

data_sp_sem <- sp_st_data %>%
  filter(station %in% st_mono_trends_stable_rich_bm) %>%
  select(c("station", all_of(var_analysis), "basin", "log_RC1", "log_RC2")) %>%
  na.omit()
data_tps_sem <- slope_com_var_no_covar %>%
  filter(station %in% st_mono_trends_stable_rich_bm) %>%
  select(station, starts_with(var_analysis)) %>%
  na.omit()
```
# SEM

## Temporal SEM

```{r}
sem <- list(
  lm(ct_ff ~ log_bm_std + log_rich_std, data = data_tps_sem,
    weight = 1 / log_rich_std_strd_error * log_bm_std_strd_error *
      ct_ff_strd_error),
    lm(w_trph_lvl_avg ~ log_bm_std + log_rich_std, data = data_tps_sem,
      weight = 1 / log_rich_std_strd_error * log_bm_std_strd_error *
        w_trph_lvl_avg_strd_error),
      lm(log_bm_std ~ log_rich_std, data = data_tps_sem,
        weight = 1 / log_rich_std_strd_error * log_bm_std_strd_error
        )
)
plot(as.psem(sem),
  node_attrs = data.frame(
    shape = "rectangle", fixedsize = TRUE, width = 1, color = "black",
    fillcolor = "lightgrey")
)
```

## Spatial SEM

```{r}
sp_sem <- list(
     log_rich_std = lmer(log_rich_std ~ log_RC1 + log_RC2 + (1 | basin), data = data_sp_sem),
     log_bm_std = lmer(log_bm_std ~ log_rich_std + log_RC1 + log_RC2 + (1 | basin), data = data_sp_sem),
     ct_ff = lmer(ct_ff ~ log_rich_std + log_bm_std + log_RC1 + log_RC2 + (1 | basin), data = data_sp_sem),
     w_trph_lvl_avg = lmer(w_trph_lvl_avg ~ log_rich_std + log_bm_std + log_RC1 + log_RC2 +
       (1 | basin), data = data_sp_sem)
)
sp_psem <- as.psem(sp_sem)
summary_sp_sem <- summary(sp_psem)
plot(sp_psem,
  node_attrs = data.frame(
    shape = "rectangle", fixedsize = TRUE, width = .8, color = "black",
    fillcolor = "lightgrey", nodesep = 2))
```

- Bootstrap (10000) takes a very long time: 
```{r, eval=FALSE}
sp_semeff <- semEff(sp_sem, R = 10000, seed = 13, parallel = "no", type = "parametric")
save(sp_semeff, file = get_mypath("data", "sp_semeff.rda"))

semeff <- semEff(sem, R = 10000, seed = 13, parallel = "no")
save(semeff, file = get_mypath("data", "semeff.rda"))

semeff_perc <- semEff(sem, R = 100, seed = 13, parallel = "no",
  ci.type = "perc"
)
save(semeff_perc, file = get_mypath("data", "semeff_perc.rda"))

sp_semeff_perc <- semEff(sp_sem, R = 100, seed = 13, parallel = "no",
  ci.type = "perc", ran.eff = "basin"
)
save(sp_semeff_perc, file = get_mypath("data", "sp_semeff_perc.rda"))

sp_semeff_df <- from_semEff_to_table(x = sp_semeff_perc)
semeff_df <- from_semEff_to_table(x = semeff_perc)
```

# Direct, indirect and total effects

```{r}
load(get_mypath("data", "semeff_perc.rda"))
load(get_mypath("data", "sp_semeff_perc.rda"))
```

```{r}
tps_sem_tbl <- from_semEff_to_table(semeff_perc)
tps_sem_tbl[, c("response", "predictor")] <-
  map(c("response", "predictor"),
  ~ var_replacement()[tps_sem_tbl[[.x]]])

sp_sem_tbl <- from_semEff_to_table(sp_semeff_perc)
sp_sem_tbl[, c("response", "predictor")] <-
  map(c("response", "predictor"),
  ~ var_replacement()[sp_sem_tbl[[.x]]])
```

## Temporal SEM

- Temporal sem:
  - Main difference with previous approach: bootstrapping reveals that
    biomass have no more significant direct effect on connectance where as
    anova, regression, and then SEM told us that it was significant

```{r}
tps_sem_tbl$predictor = cell_spec(tps_sem_tbl$predictor,
  color = case_when(
    tps_sem_tbl$lower_ci > 0 & tps_sem_tbl$upper_ci > 0 ~ "red", 
    tps_sem_tbl$lower_ci < 0 & tps_sem_tbl$upper_ci < 0 ~ "blue",
    TRUE ~ "black"
      ))
kbl(tps_sem_tbl, align = "c", format = "html", escape = FALSE) %>%
  kable_paper(full_width = F) %>%
  collapse_rows(columns = 1:2, valign = "top")
```

## Spatial SEM

```{r}
sp_sem_tbl$predictor = cell_spec(
  sp_sem_tbl$predictor,
  color = case_when(
    sp_sem_tbl$lower_ci > 0 & sp_sem_tbl$upper_ci > 0 ~ "red", 
    sp_sem_tbl$lower_ci < 0 & sp_sem_tbl$upper_ci < 0 ~ "blue",
    TRUE ~ "black"
      ))
kbl(sp_sem_tbl, align = "c", format = "html", escape = FALSE) %>%
  kable_paper(full_width = F) %>%
  collapse_rows(columns = 1:2, valign = "top")
```