### 2022-04-04 sem.Rmd
Update temporal SEM with weight as sum of sd instead of multiplication

---
title: "SEM"
author: AD 
date: \today
output:
  bookdown::html_document2:
    toc: true
    theme: "readable"
    code_folding: hide
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  message = FALSE,
  warning = FALSE,
  results = TRUE
)
library(DiagrammeR)
DiagrammeR::render_graph(graph)
```
```{r}
plot.psem <- function(
  x,
  return=FALSE,
  node_attrs = data.frame(shape = "rectangle", color = "black",
    fillcolor = "white"),
  edge_attrs = data.frame(style = "solid", color="black"),
  ns_dashed = T, alpha=0.05,
  show = "std", digits = 3, 
  add_edge_label_spaces = TRUE, ...
  ){

  #get the coefficients table
  ctab <- coefs(x)
  ctab$Response <- as.character(ctab$Response)
  ctab$Predictor <- as.character(ctab$Predictor)

  #make a nodes DF
  unique_nodes <- unique(c(ctab$Response, ctab$Predictor))
  nodes <- create_node_df(n = length(unique_nodes),
    nodes = unique_nodes,
    type = "lower",
    label = unique_nodes)
  nodes <- cbind(nodes, node_attrs)
  nodes[] <- lapply(nodes, as.character)
  nodes$id <- as.numeric(nodes$id)

  #make an edges DF
  edges <- create_edge_df(
    from = match(ctab$Predictor, unique_nodes),
    to = match(ctab$Response, unique_nodes))

  edges <- data.frame(edges, edge_attrs)
  edges[] <- lapply(edges, as.character)
  edges$id <- as.numeric(edges$id)
  edges$from <- as.numeric(edges$from)
  edges$to <- as.numeric(edges$to)
  if(ns_dashed) edges$style[which(ctab$P.Value>alpha)] <- "dashed"
  if(show == "std") edges$label = round(ctab$`Std.Estimate`, digits)
  if(show == "unstd") edges$label = round(ctab$Estimate, digits)
  if(add_edge_label_spaces) edges$label = paste0(" ", edges$label, " ")

  #turn into a graph
  sem_graph <- create_graph(nodes, edges, directed=TRUE)

  if(return) return(sem_graph)

  render_graph(sem_graph, ...)

}
```

```{r}
library(drake)
loadd(slope_com_var_no_covar, st_mono_trends_rich_bm)
library(piecewiseSEM)
library(semEff)
library(kableExtra)
library(lme4)
mypath <- rprojroot::find_package_root_file
source(mypath("R/misc.R"))
source(mypath("R/variable_shortcut.R")) # Define your custom code as a bunch of functions.
source_dir(get_mypath("R")) # Define your custom code as a bunch of functions.

mytheme <- theme_cowplot() +
background_grid()

theme_set(mytheme)
```

```{r}
loadd(sp_st_data)
loadd(slope_com_var_no_covar)
loadd(st_mono_trends_stable_rich_bm)
```

```{r}
var_analysis <- c("ct_ff", "log_bm_std", "log_rich_std", "w_trph_lvl_avg")

data_sp_sem <- sp_st_data %>%
  filter(station %in% st_mono_trends_stable_rich_bm) %>%
  select(c("station", all_of(var_analysis), "basin", "log_RC1", "log_RC2")) %>%
  na.omit()
data_tps_sem <- slope_com_var_no_covar %>%
  filter(station %in% st_mono_trends_stable_rich_bm) %>%
  select(station, starts_with(var_analysis)) %>%
  na.omit() %>%
  mutate(
    error_mult_ct =
      (log_rich_std_strd_error * log_bm_std_strd_error *
        ct_ff_strd_error),
      error_mult_tlvl =
        (log_rich_std_strd_error * log_bm_std_strd_error *
          w_trph_lvl_avg_strd_error),
        error_mult_bm = 
          (log_rich_std_strd_error * log_bm_std_strd_error),
        error_additive_ct = 
          (log_rich_std_strd_error + log_bm_std_strd_error +
            ct_ff_strd_error),
          error_additive_tlvl =
            (log_rich_std_strd_error + log_bm_std_strd_error +
              w_trph_lvl_avg_strd_error),
            error_additive_bm = (log_rich_std_strd_error + log_bm_std_strd_error)
  )
```
# SEM

## Temporal SEM

```{r}
sem <- list(
  lm(ct_ff ~ log_bm_std + log_rich_std, data = data_tps_sem,
    weight = 1 / log_rich_std_strd_error * log_bm_std_strd_error *
      ct_ff_strd_error),
    lm(w_trph_lvl_avg ~ log_bm_std + log_rich_std, data = data_tps_sem,
      weight = 1 / log_rich_std_strd_error * log_bm_std_strd_error *
        w_trph_lvl_avg_strd_error),
      lm(log_bm_std ~ log_rich_std, data = data_tps_sem,
        weight = 1 / log_rich_std_strd_error * log_bm_std_strd_error
        )
)
plot(as.psem(sem),
  node_attrs = data.frame(
    shape = "rectangle", fixedsize = TRUE, width = 1, color = "black",
    fillcolor = "lightgrey"),
  layout = "tree"
)
render_graph
```

## Spatial SEM

```{r}
sp_sem <- list(
     log_rich_std = lmer(log_rich_std ~ log_RC1 + log_RC2 + (1 | basin), data = data_sp_sem),
     log_bm_std = lmer(log_bm_std ~ log_rich_std + log_RC1 + log_RC2 + (1 | basin), data = data_sp_sem),
     ct_ff = lmer(ct_ff ~ log_rich_std + log_bm_std + log_RC1 + log_RC2 + (1 | basin), data = data_sp_sem),
     w_trph_lvl_avg = lmer(w_trph_lvl_avg ~ log_rich_std + log_bm_std + log_RC1 + log_RC2 +
       (1 | basin), data = data_sp_sem)
)
sp_psem <- as.psem(sp_sem)
summary_sp_sem <- summary(sp_psem)
plot(sp_psem,
  node_attrs = data.frame(
    shape = "rectangle", fixedsize = TRUE, width = .8, color = "black",
    fillcolor = "lightgrey", nodesep = 2))
```

- Bootstrap (10000) takes a very long time: 
```{r, eval=FALSE}
sp_semeff <- semEff(sp_sem, R = 10000, seed = 13, parallel = "no", type = "parametric")
save(sp_semeff, file = get_mypath("data", "sp_semeff.rda"))

semeff <- semEff(sem, R = 10000, seed = 13, parallel = "no")
save(semeff, file = get_mypath("data", "semeff.rda"))

semeff_perc <- semEff(sem, R = 100, seed = 13, parallel = "no",
  ci.type = "perc"
)
save(semeff_perc, file = get_mypath("data", "semeff_perc.rda"))

sp_semeff_perc <- semEff(sp_sem, R = 100, seed = 13, parallel = "no",
  ci.type = "perc", ran.eff = "basin"
)
save(sp_semeff_perc, file = get_mypath("data", "sp_semeff_perc.rda"))

sp_semeff_df <- from_semEff_to_table(x = sp_semeff_perc)
semeff_df <- from_semEff_to_table(x = semeff_perc)
```

# Direct, indirect and total effects

```{r}
load(get_mypath("data", "semeff_perc.rda"))
load(get_mypath("data", "sp_semeff_perc.rda"))
```

```{r}
tps_sem_tbl <- from_semEff_to_table(semeff_perc)
tps_sem_tbl[, c("response", "predictor")] <-
  map(c("response", "predictor"),
  ~ var_replacement()[tps_sem_tbl[[.x]]])

sp_sem_tbl <- from_semEff_to_table(sp_semeff_perc)
sp_sem_tbl[, c("response", "predictor")] <-
  map(c("response", "predictor"),
  ~ var_replacement()[sp_sem_tbl[[.x]]])
```

## Temporal SEM

- Temporal sem:
  - Main difference with previous approach: bootstrapping reveals that
    biomass have no more significant direct effect on connectance where as
    anova, regression, and then SEM told us that it was significant

```{r}
tps_sem_tbl$predictor = cell_spec(tps_sem_tbl$predictor,
  color = case_when(
    tps_sem_tbl$lower_ci > 0 & tps_sem_tbl$upper_ci > 0 ~ "red", 
    tps_sem_tbl$lower_ci < 0 & tps_sem_tbl$upper_ci < 0 ~ "blue",
    TRUE ~ "black"
      ))
kbl(tps_sem_tbl, align = "c", format = "html", escape = FALSE) %>%
  kable_paper(full_width = F) %>%
  collapse_rows(columns = 1:2, valign = "top")
```

## Spatial SEM

```{r}
sp_sem_tbl$predictor = cell_spec(
  sp_sem_tbl$predictor,
  color = case_when(
    sp_sem_tbl$lower_ci > 0 & sp_sem_tbl$upper_ci > 0 ~ "red", 
    sp_sem_tbl$lower_ci < 0 & sp_sem_tbl$upper_ci < 0 ~ "blue",
    TRUE ~ "black"
      ))
kbl(sp_sem_tbl, align = "c", format = "html", escape = FALSE) %>%
  kable_paper(full_width = F) %>%
  collapse_rows(columns = 1:2, valign = "top")
```

# Weight

#

```{r}
weight_df <- data_tps_sem %>%
  select(c(station, starts_with("error"))) %>%
  pivot_longer(-station, names_to = "variable", values_to = "value") %>%
  mutate(variable = str_remove(variable, "error_")) %>%
  separate(variable, c("type", "variable"), sep = "_") %>%
  group_by(type, variable) %>%
  mutate(weight = value / sum(value))
weight_df %>%
  summarise(sum_weight = sum(weight))
weight_df %>%
  ggplot(aes(x = weight)) +
  geom_histogram() +
  facet_grid(rows = vars(variable), cols = vars(type), scale = "free_x")
```

- La distribution des poids est super asymétrique avec les multiplications de
  standard deviation, c'est pour ça que ça merde
- Ci-dessous, j'ai refais les sem avec des poids additifs, les échantillons des
  bootstraps sont nettement plus stables

```{r}
semeff_mult_weight <- semEff(sem, R = 150, seed = 123, parallel = "no",
  ci.type = "perc"
)

sem_no_weight <- list(
  lm(ct_ff ~ log_bm_std + log_rich_std, data = data_tps_sem),
  lm(w_trph_lvl_avg ~ log_bm_std + log_rich_std, data = data_tps_sem),
  lm(log_bm_std ~ log_rich_std, data = data_tps_sem)
)

sem_additive_weight <- list(
  lm(ct_ff ~ log_bm_std + log_rich_std, data = data_tps_sem,
    weight = 1 / (log_rich_std_strd_error + log_bm_std_strd_error +
      ct_ff_strd_error)),
  lm(w_trph_lvl_avg ~ log_bm_std + log_rich_std, data = data_tps_sem,
    weight = 1 / (log_rich_std_strd_error + log_bm_std_strd_error +
      w_trph_lvl_avg_strd_error)),
  lm(log_bm_std ~ log_rich_std, data = data_tps_sem,
    weight = 1 / (log_rich_std_strd_error + log_bm_std_strd_error)
  )
)

semeff_no_weight <- semEff(sem_no_weight, R = 100, seed = 123, parallel = "no",
  ci.type = "perc"
)
semeff_additive_weight <- semEff(sem_additive_weight, R = 100, seed = 123, parallel = "no",
  ci.type = "perc"
)
semeff_avg_weight <- semEff(sem_avg_weight, R = 100, seed = 123, parallel = "no",
  ci.type = "perc"
)
```

```{r}
semeff_add_weight_tab <- from_semEff_to_table(semeff_avg_weight) %>%
  mutate(
    response = var_replacement(several_lines = TRUE)[response],
    predictor = var_replacement(several_lines = TRUE)[predictor]

  )
semeff_no_weight_tab <- from_semEff_to_table(semeff_no_weight) %>%
  mutate(
    response = var_replacement(several_lines = TRUE)[response],
    predictor = var_replacement(several_lines = TRUE)[predictor] 
  )

semeff_mult_weight_tab <- from_semEff_to_table(semeff_mult_weight) %>%
  mutate(
    response = var_replacement(several_lines = TRUE)[response],
    predictor = var_replacement(several_lines = TRUE)[predictor] 
  )
```

```{r}
plot.semeff <- function(
  x,
  return=FALSE,
  node_attrs = data.frame(
    shape = "rectangle", fixedsize = TRUE, width = 1, color = "black",
    fillcolor = "lightgrey"),
  edge_attrs = data.frame(style = "solid", color="black"),
  ns_dashed = T, alpha=0.05,
  show = "ci", digits = 2,
  width_eff = TRUE, scale_fc = 1, 
  attr_theme = "default",
  add_edge_label_spaces = TRUE, ...
  ){

  #get the coefficients table
  ctab <- x %>%
    filter(effect_type == "direct")

  #make a nodes DF
  unique_nodes <- unique(c(ctab$response, ctab$predictor))
  nodes <- create_node_df(n = length(unique_nodes),
    nodes = unique_nodes,
    type = "lower",
    label = unique_nodes)
  nodes <- cbind(nodes, node_attrs)
  nodes[] <- lapply(nodes, as.character)
  nodes$id <- as.numeric(nodes$id)

  #make an edges DF
  edges <- create_edge_df(
    from = match(ctab$predictor, unique_nodes),
    to = match(ctab$response, unique_nodes))

  edges <- data.frame(edges, edge_attrs)
  edges[] <- lapply(edges, as.character)
  edges$id <- as.numeric(edges$id)
  edges$from <- as.numeric(edges$from)
  edges$to <- as.numeric(edges$to)
  if(ns_dashed) edges$style[which(sign(ctab$lower_ci) != sign(ctab$upper_ci))] <- "dashed"
  if(width_eff) edges$penwidth <- abs(ctab$effect) * scale_fc 
  if(show == "std") edges$label = round(ctab$effect, digits)
  if(show == "ci") { edges$label = paste0(
    round(ctab$effect, digits),
    " [", round(ctab$lower_ci, digits),"; ",
    round(ctab$upper_ci, digits), "]"
  )
    }
  if(add_edge_label_spaces) edges$label = paste0(" ", edges$label, " ")

  #turn into a graph
  sem_graph <- create_graph(nodes, edges, directed=TRUE,
    attr_theme = attr_theme)

  if(return) return(sem_graph)

  render_graph(sem_graph, ...)

}
```

```{r}
p_semeff_add_weight_tab <- plot.semeff(
  x = semeff_add_weight_tab,
  show = "ci",
  node_attrs = data.frame(
    shape = "rectangle", fixedsize = TRUE, width = 1.1, color = "black",
    fillcolor = "lightgrey"),
  edge_attrs = data.frame(style = "solid",  color="black"),
  scale_fc = 15,
  return = TRUE, attr_theme = "tb"
)

export_graph(
  graph = p_semeff_add_weight_tab,
  file_name = here::here("figures", "sem_tps_avg_weight.svg"),
  file_type = "svg"
)

p_semeff_no_weight_tab <- plot.semeff(
  x = semeff_no_weight_tab,
  show = "ci",
  node_attrs = data.frame(
    shape = "rectangle", fixedsize = TRUE, width = 1.1, color = "black",
    fillcolor = "lightgrey"),
  edge_attrs = data.frame(style = "solid",  color="black"),
  scale_fc = 15,
  return = TRUE, attr_theme = "tb"
)
render_graph(p_semeff_no_weight_tab)

export_graph(
  graph = p_semeff_no_weight_tab,
  file_name = here::here("figures", "sem_tps_no_weight.svg"),
  file_type = "svg"
)

p_semeff_mult_weight_tab <- plot.semeff(
  x = semeff_mult_weight_tab,
  show = "ci",
  node_attrs = data.frame(
    shape = "rectangle", fixedsize = TRUE, width = 1.1, color = "black",
    fillcolor = "lightgrey"),
  edge_attrs = data.frame(style = "solid",  color="black"),
  scale_fc = 15,
  return = TRUE, attr_theme = "tb"
)

export_graph(
  graph = p_semeff_mult_weight_tab,
  file_name = here::here("figures", "sem_tps_mult_weight.svg"),
  file_type = "svg"
)

```

### Multiplicative weight

```{r}
render_graph(p_semeff_mult_weight_tab)
```

### Additive weight

```{r}
render_graph(p_semeff_add_weight_tab)
```

### No weight

```{r}
render_graph(p_semeff_no_weight_tab)
```