### 2022-11-06 paper/appendix.Rmd
Add tables to appendix

---
title: Appendix
subtitle:
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: false 
fontsize: 10pt
header-includes:
   - \usepackage{natbib}
   - \newcommand{\beginsupplement}{\renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} 
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
   - \renewcommand{\arraystretch}{1}
geometry: margin=2cm
bibliography: references.bib
number_sections: false 
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.align = "center",
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE
)
options(
  #kableExtra.latex.load_packages = FALSE,
  knitr.table.format = "latex"
)
```

```{r, echo=FALSE, message=FALSE, results="hide"}
library(targets)
library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
library(sf)
library(nlme)

mypath <- rprojroot::find_package_root_file
source(mypath("R/variable_shortcut.R")) # Define your custom code as a bunch of
#functions.

source(mypath("R/misc.R"))
source_dir(mypath("R")) # Define your custom code as a bunch of functions.

mytheme <- theme_cowplot() +
  background_grid()

theme_set(mytheme)
```
```{r}
var_analysis <- c("ct_ff", "w_trph_lvl_avg", "log_rich_std", "log_bm_std")
```

\beginsupplement
\newpage

```{r}
tar_load(tab_inla_r2_rand)
kable(tab_inla_r2_rand,
  row.names = FALSE,
  booktabs = T,
  caption = "Marginal and Conditional R-squared for the INLA models estimating
  the temporal trends of the food-web structure. The table also displays the
  standard deviation associated to the random effects and the general error of
  the models. Confidence interval at 95\\% are displayed between square brackets.",
  label = "inlarsq") %>%
  collapse_rows(columns = c(1, 2, 3), valign = "top") %>%
  kable_styling(latex_options = c("scale_down"))
```

```{r}
tps10_cap <- paste0("Distribution of temporal trends per decade of the food-web
  structure by site estimated from the Bayesian models using Best Linear
  Unbiaised Prediction (BLUP). Species richness and biomass temporal trends are
  reported in percentage change per decade. Dashed red line and solid black line
  respectively indicate 0 and median change.")

```
```{r tps10, fig.cap=tps10_cap}
knitr::include_graphics(here::here("figures", "p_hist_site_tps10.pdf"))
```

```{r}
rand_sim_cap <- paste0(
  "Simulations of (a) species and (b) nodes extinctions in the metaweb according
  to three scenarios: random extinction sequence, decreasing trophic level,
  increasing trophic level."

)
```

```{r ext, fig.cap=rand_sim_cap}
knitr::include_graphics(here::here("figures", "tot_ext_sim.pdf"))
```

```{r}
cor_cap <- paste0("Correlation among food-web metrics for a) temporal trends by
  site, b) median values by site (cf Main Methods)", " Pisc: piscivorous, rich.:
species richness, node: node of the food-web corresponding to a trophic species,
prop.: proportion."
)
```

```{r cor, fig.cap=cor_cap}
knitr::include_graphics(here::here("figures", "p_corr.pdf"))
```

```{r}
obs_fit_cap <- "Fitted versus observed values from the INLA models estimating
the temporal trends of the food-web structure. The black line depicts the first
bissector. Pearson's correlation coefficients are displayed."
```

```{r, fig.cap=obs_fit_cap}
knitr::include_graphics(here::here("figures", "p_obs_fit.pdf"))
```

```{r}
tar_load(tabvif)
kable(tabvif,
  row.names = FALSE,
  booktabs = T,
  caption = "Variance Inflation Factor for the more complex linear models
  included in the Structural Equation Models.",
  label = "vif",
  )
```

```{r}
tar_load(r2_sem_tab)
kable(r2_sem_tab,
  row.names = FALSE,
  booktabs = T,
  caption = "Marginal and conditional R squared for the Structural Equation
  Models.",
  label = "r2sem",
  ) %>%
  kableExtra::collapse_rows(columns = 1, valign = "top")
```

```{r}
tar_load(c(pca_table, op_table, com_metric_tab))
tab_quanti <- purrr::map_dfr(
  list(com_metric_tab, op_table, pca_table),
  function(x) {colnames(x)[1] <- "Variable"; x}
)
purrr::map_int(list(com_metric_tab, op_table, pca_table), nrow)

tab_quanti %>%
  kbl(
    caption = "Summary distribution of community metrics, fish monitoring and of
    environmental variables used in the PCA for the spatial SEM.",
    booktabs = T
    ) %>%
  kable_styling() %>%
  pack_rows("Community metrics", 1, 8) %>%
  pack_rows("Summary of fish monitoring", 9, 12) %>%
  pack_rows("Summary of Envionmental variables", 13, 26)
```