### 2024-08-08 paper/manuscript.Rmd
Rewrite intro + reorder appendix

---
title: "Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs"
author: Alain Danet (a), Loubna El Madouri (a, c), Maud Mouchet (a), Elisa Thébault (d) and Colin Fontaine (a)
date: "Research report"
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
  officedown::rdocx_document:
    number_sections: false
    base_format: "bookdown::word_document2"
always_allow_html: true
bibliography: references.bib
link-citations: true
#csl: ecology-letters.csl
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
csl: pnas.csl
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path = "figures/"
)
options(
  knitr.table.format = "latex",
  knitr.graphics.rel_path = FALSE
)
```

```{r, results = FALSE}
library(targets)
library(tidyverse)
library(magrittr)
library(cowplot)
library(viridis)
library(kableExtra)
library(sf)
mypath <- rprojroot::find_package_root_file
source(mypath("R/variable_shortcut.R"))
source(mypath("R/misc.R"))
source_dir(get_mypath("R")) # Define your custom code as a bunch of functions.

mytheme <- theme_cowplot() +
background_grid()

theme_set(mytheme)
```

```{r}
tar_load(full_data2)
nb_st <- unique(full_data2$station) %>% length
nb_op <- nrow(full_data2) %>% round(., -3)

rm(full_data2)
tar_load(glob_tps_trends_decade_no_drivers)
```

- Corresponding author: Alain Danet (a.h.danet@sheffield.ac.uk)

- Author affiliations:

(a) Centre d’Ecologie et des Sciences de la Conservation, UMR 7204 MNHN7
CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de Paris, 43 rue
8 Buffon, 75005 Paris, France

(b) Department of Biosciences, Alfred Denny Building, University of Sheffield,
Western Bank, Sheffield S10 2TN, United Kingdom

(c) Laboratoire d’Écologie Fonctionnelle et Environnement, UMR 5245
CNRS-UT3-INPT, Avenue de l’Agrobiopôle, Toulouse, France

(d) Sorbonne Université, CNRS, IRD, INRAE, Université Paris Est Créteil, 15
Institute of Ecology and Environmental Sciences of Paris (iEES-Paris), Paris,
France

- Classification: Biological Sciences (major) - Ecology (minor)

- Keywords: food-web, biodiversity changes, assembly rules, space-for-time

# Abstract (250 / 250 words max)

While there are widespread evidence of biodiversity decline, little is known how
food-web disassemble in response. Such knowledge is crucial as food-web
structure determine ecosystem functioning and stability to perturbations.
However, assessing how changes in biodiversity affect changes in food-web
structure require the availability of food-web timeseries and to account for the
differential effects of different biodiversity facets on food-web structure.
Here, we used a standardized monitoring of riverine fish communities over the
last decades (1995-2018) in France, from which we inferred the food-web
structure for each sampling events using diet data and allometric scaling rules.
We estimated the temporal trends in community biomass, species richness, and
food-web structure using hierarchical Bayesian models. We assessed the effects
of community biomass and species richness temporal changes on temporal changes
in food-web structure: connectance, average trophic level and redundancy. We
further assess the generality and robustness of our findings by replicating the
analysis over space and using theoretical simulations with a bioenergetic model.
Our result suggest that food-webs that experienced a decline in community
biomass experienced conjointly a loss of species at upper trophic levels and
conversely. While species richness temporal trends had less clear effects on
food-web structure, mainly due through its indirect effects on community biomass
temporal trends. Both community biomass and species richness declines were
associated with a decrease in redundancy, raising concerns about robustness of
such communities to further perturbations. Our study highlight how biodiversity
changes affect food-web structure, raising concerns about consequences on
ecosystem functioning and robustness.

# Significance (109 / 120 words max)

Biodiversity temporal changes can lead to simultaneous changes in food-web
structure, which is turn can have consequences on ecosystem functioning and
stability. We provide empirical evidence that temporal changes in community
biomass and species richness are linked to changes contrasted assembly and
disassembly processes in food-webs in natural settings. We then show that loss
in both community biomass and species richness are related to a loss of
redundancy in the food-webs, which raise concerns about the robustness of such
communities to perturbations. We further added spatial and theoretical evidence
suggesting that our results are quite robust and open to assess how biodiversity
changes might affect ecosystem functioning and robustness.

# Introduction (487 words)

Biodiversity is experiencing massive declines due to continuously increasing
human pressures, which is expected to have striking consequences on ecosystem
functioning and fragility. It is increasing recognised that biodiversity
changes are widely heterogeneous spatially [@mcgill_fifteen_2015;
@blowes_geography_2019], such as species richness declines and increases can be
related to spatial variations in human pressures [@cardinale_is_2018;
@danet_past_2024]. It is increasing recognised that biodiversity changes alter
community structure [@dornelas_assemblage_2014; @hillebrand_biodiversity_2018;
@danet_past_2024], but so far, one key component of biodiversity, species
interactions, has been mostly ignored by studies assessing temporal changes in
biodiversity, despite the importance of biotic interactions for species
coexistence, ecosystem functioning and stability [@chesson_mechanisms_2000;
@thebault_food-web_2003; @oconnor_general_2017; @danet_species_2021]. A better
understanding of biodiversity changes will then need to account how biodiversity
changes alter the structure of ecological networks.

Yet, both species richness and food-web structure have a significant impact on
coexistence [@chesson_interaction_2008], ecosystem functioning
[@thebault_food-web_2003; @wang_biodiversity_2018], stability
[@danet_species_2021; @eschenbrenner_diversity_2023], and ecosystem robustness
to perturbations [@dunne_network_2002; @allesina_functional_2009;
@bodini_using_2009; @allesina_functional_2009]. Many predator species play a
keystone role by protecting preys from competitive exclusion
[@heithaus_predicting_2008; @johnson_rarity_2007; @estes_trophic_2011], thereby
maintaining species richness. Food-web with higher trophic levels also increase
ecosystem productivity [@thebault_food-web_2003; @wang_biodiversity_2018] and
community temporal stability [@danet_species_2021;
@eschenbrenner_diversity_2023]. Finally, the number and distribution of trophic
links in food-webs determine the robustness of communities to perturbations
[@dunne_network_2002; @allesina_functional_2009; @bodini_using_2009;
@allesina_functional_2009]. Accounting for how food-web assemble and disassemble
according to biodiversity changes is then crucial to evaluate the consequences
of biodiversity changes on ecosystem functioning and stability.

However, assessing the effects of biodiversity changes on food-web structure
requires the availability of multiple temporal timeseries of biodiversity and
food-webs structure, to disentangle the direct and indirect effects of
biodiversity changes on food-web structure, and to account for theories about
the multiple constraints on food-web assembly. We propose to monitor three
structural properties of food-webs and two common monitored biodiversity facets
(i.e. community biomass and species) that are grounded in food-web assembly,
ecosystem functioning, and stability research areas: connectance (i.e. degree of
generalism), weighted average trophic level (i.e. vertical location of biomass),
and redundancy (i.e. the proportion of redundant links). Their are pieces of
evidence to connect those food-web structural metrics to changes in ecosystem
functioning and robustness that we review before to present the temporal
timeseries that allows to assess them.

Food-web assembly theory provides predictions on how food-web structure should
change according to changes in community biomass and species richness. The
energetic paradigm and the trophic theory of biogeography
[@lindeman_trophic-dynamic_1942; @gravel_trophic_2011] stand that the presence
of top predators is conditional on the presence of preys in number enough that
it can provide enough energy from the bottom to the top to sustain high trophic
level species. In turn, it can that the presence of generalist predators, which is
positively associated to average trophic level and connectance, will be linked
more to community biomass changes than to species richness. In turn, connectance
and average trophic level are commonly reported to decrease with as species
richness increases [@dunne_network_2006; @danet_species_2021] because predators
are limited in the number of species they can consume and conversely preys are
limited in the number of predator they can sustain. Food-web structure has also
been linked to the robustness to species extinctions [@dunne_network_2002;
@allesina_who_2004; @bodini_using_2009]. Historically having being linked to
connectance, i.e. the number of links, more recent studies have proposed to
separate links that are functional or redundant [@allesina_who_2004;
@allesina_functional_2009; @bodini_using_2009]. The idea is that the loss of
functional links will trigger secondary extinctions because it suppresses
obligatory pathways from resources to consumers while the loss of redundant
links do not remove such obligatory pathways [@allesina_who_2004]. It means that
redundant trophic links enhance ecosystem robustness as they can be safely
removed without triggering secondary extinctions. Thereby we expected that
higher species richness results in higher redundancy in food-webs, meaning that
they create more independent pathways connecting basal producers to higher
trophic level. Finally, biodiversity and ecosystem functioning theory suggest
that in context of significant niche complementarity, community biomass should
be positively related to species richness [@wang_biodiversity_2018;
@schneider_animal_2016], such as it is necessary to account for the indirect
effects of species richness on food-web structure through community biomass.

Here, we report on a study that combine empirical assessment of food-web
structure and biodiversity temporal changes in natural settings, extinction
simulations in empirical food-webs, and theoretical simulations using
bioenergetic model. The availability of food-web timeseries used to be scarce
[@kortsch_disentangling_2021], but the increasing availability of biodiversity
timeseries and the development of methods to reconstruct food-webs
[@bonnaffe_comparison_2021; @gravel_inferring_2013; @danet_species_2021] makes
possible to overcome this limitation. We used a standardized monitoring of fish
riverine communities on the period 1995-2018 over France, documenting the
individual size of fishes in `r nb_st` sites during ten years or more (Table S1). We
reconstructed around `r nb_op` food-webs based on allometric relationships and
food diet data, and then computed connectance, and average trophic level
weighted by trophic species biomass as well as community biomass and species
richness (Fig. \@ref(fig:matmet)). This dataset and method has already been used
in previous studies [@danet_species_2021; @bonnaffe_comparison_2021;
@bonnaffe_interaction_2024]. Using hierarchical Bayesian models, we estimated
temporal trends in food-web structure, community biomass and richness at each
site [@danet_past_2024]. We then used structural equation models (SEM) to
assess the effects of community biomass and species richness temporal changes on
changes in food-web structure, while accounting for the effects of species
richness on community biomass. We then assessed if those effects are holding in
space, i.e. if the space-for-time substitution hypothesis hold
[@blois_space_2013; @damgaard_critique_2019; @adler_contrasting_2007]. We did so
by considering median values of food-web structure and community structure over
the timeseries at each site (Fig. \@ref(fig:matmet)). Finally, we ran
simulations using the bioenergetic model to assess if our results hold beyond
our empirical fish food-webs.

```{r}
fig_mat_cap <- paste0("Using sites of stream fishes in French streams that
  have been monitored at least 10 years, 1) we infered the food-web at each site
and each sampling event based on ontogenic food diet information for each
species and predator-prey body size ratio. 2) We computed the temporal trends
and the median of community biomass, species richness, connectance, average
trophic level and evenness of biomass at each site. 3) We then assessed the
relationships among food-web structure descriptors, (a) in time and (b) across space.")
```

```{r matmet, fig.cap=fig_mat_cap, out.width="100%"}
knitr::include_graphics(get_mypath("figures", "mat_met.png"))
```

# Results

```{r, results=FALSE}
gbm <- get_effect_ci(resp = "log_bm_std", ci_lvl = "level:0.80", p = TRUE)
gr <- get_effect_ci(resp = "log_rich_std", ci_lvl = "level:0.80", p = TRUE)
gt <- get_effect_ci(resp = "w_trph_lvl_avg", ci_lvl = "level:0.80", p = FALSE)
gct <- get_effect_ci(resp = "connectance", ci_lvl = "level:0.80", p = FALSE)
gl <- get_effect_ci(resp = "prop_redundant_links", ci_lvl = "level:0.80", p = FALSE)
```

```{r}
tar_load(c(r2_inla_tab, r2_inla, r2_sem))
r2_inla_report <- r2_inla %>%
  select(response, r2_mvp_marg95hpdci, r2_mvp_cond95hpdci, r2_mvp_cond_mean)
r2cond <- map(
  setNames(r2_inla_report$response, r2_inla_report$response),
  ~round(r2_inla_report[r2_inla_report$response == .x, ]$r2_mvp_cond_mean, 2)
)
r2ic <- map(
  setNames(r2_inla_report$response, r2_inla_report$response),
  ~r2_inla_report[r2_inla_report$response == .x, ]$r2_mvp_cond95hpdci
)
```

```{r}
tar_load(summary_pred_station_tps10)
dfv <- function(
  x = summary_pred_station_tps10,
  r = "connectance",
  v = "min",
  d = 4
) {
  out <- x[x$response == r, ][[v]]
  format(round(out, d), nsmall = d)
}
grbmn <- dfv(r = "log_bm_std", v = "min", d = 0)
grbmx <- dfv(r = "log_bm_std", v = "max", d = 0)
grrhn <- dfv(r = "log_rich_std", v = "min", d = 0)
grrhx <- dfv(r = "log_rich_std", v = "max", d = 0)
grctn <- dfv(r = "connectance", v = "min", d = 2)
grctx <- dfv(r = "connectance", v = "max", d = 2)
grtln <- dfv(r = "w_trph_lvl_avg", v = "min", d = 2)
grtlx <- dfv(r = "w_trph_lvl_avg", v = "max", d = 2)
grln <- dfv(r = "prop_redundant_links", v = "min", d = 2)
grlx <- dfv(r = "prop_redundant_links", v = "max", d = 2)
```

Using a Bayesian hierarchical model to jointly estimate the temporal trends of
food-web structure at all sites, we found weak evidence of an overall decrease
in community biomass (CI80%:`r gbm` by decade; median [Credible Interval], Fig.
S1) but no evidence of an overall temporal increase in species richness
(CI80%: `r gr` by decade) across the `r nb_st` sites. We did not find evidence
of average temporal trends in connectance and average trophic level (CI80%:
`r gct` and `r gt` respectively), but an increase in redundancy (measured by the
proportion of redundant links, CI80%: `r gl`). However, we found a great
heterogeneity in the temporal trends estimated at the site level (community
biomass: from `r grbmn`% to `r grbmx`% per decade, species richness: `r grrhn`
to +`r grrhx`%, average trophic level: `r grtln` to +`r grtlx`, connectance:
`r grctn` to +`r grctx`, redundancy: `r grln` to +`r grlx`, Fig. S1) which was
well captured by the hierarchical model with conditional R2 going from
`r r2cond$log_bm_std` for community biomass to `r r2cond$log_rich_std` for
species richness (Table S2).

```{r}
sem_cap <- "(a) Structural equation models linking the temporal trends of the food-web
structure. The path widths are proportional to the standardized slope
coefficients. The standardized slope coefficients and their associated 95%
confidence intervals obtained by bootstrapping are displayed along the path. The
marginal R-squared of the linear models are displayed. (b to f) Direct bivariate
relationships derived from the structural equation model. The distribution of
the site temporal trends are displayed in Fig. S1. All the effects are reported
in Table S3."
```

```{r sem, fig.cap = sem_cap, out.width = "100%"}
tar_load(p_sem_tps_file)
knitr::include_graphics(p_sem_tps_file)
```

```{r}
tar_load(c(semeff_tot, semeff_tot_ok))
semeff_tot <- semeff_tot %>%
  mutate(across(c(response, predictor), ~str_remove(.x, "mean_")))

rdtps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std")
    )
)
rdsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", model = "spatial"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", model = "spatial")
    )
)
rdm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Direct", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Direct", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
```

```{r}
ridtps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "indirect")
    )
)
ridsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "indirect")
    )
)
ridm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Indirect", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Indirect", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
rttps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "total")
    )
)
rtsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std",
      model = "spatial", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "total")
    )
)
rtm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Total", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Total", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
```

Using structural equation models to disentangle the direct and indirect effects
of temporal trends in fish species richness and community biomass on temporal
trends in food-web structure, we found a strong positive effect of the temporal
trends in species richness on community biomass ($r_\delta =$ `r rdtps$bm$r`,
Fig. \@ref(fig:sem) a and b, $r_\delta$: standardized slope coefficient, [95%
CI], Table S3), meaning that both temporal trends had the same direction in average. In
other word, food-webs that have experienced a decrease in species richness over
time have decreased in community biomass, and conversely, those that increased
in species richness have increased in community biomass.
Temporal trends in community biomass had in turn a positive effect on connectance,
average trophic level, and redundancy (resp. $r_\delta =$ `r rdtps$ct$bm`,
`r rdtps$tlvl$bm`, `r rdtps$red$bm`, Fig. \@ref(fig:sem)b, c, d), meaning that
food-web whose community biomass decreased over time have decreased in
connectance, average trophic level and redundancy, and
conversely, those that have experienced an increase in biomass have increased in
connectance, average trophic level and redundancy. We further found that the temporal
trends of community biomass were positively associated to the temporal trends of
both number and proportion of piscivorous species present (Fig. S2a), meaning
biomass trends are linked to the presence of high trophic level trophic
species.

We further tested if we find the same patterns using spatial variations in
food-web structure rather than temporal changes over time, i.e. if the
space-for-time hypothesis hold. To do so, we used the same dataset and same
structural equation model but with the median food-web structure per site
instead of the temporal trends. We additionally performed theoretical
simulations with a bioenergetic food-web model to further get theoretical
expectations about how food-web structure changes according to community biomass and
species richness variations (see Methods). As in temporal data, species richness
had a positive effect on community biomass both in empirical data and in
theoretical simulations (resp. $r_\delta =$ `r rdsp$bm$r` and `r rdm$bm$r`, Fig.
\@ref(fig:std) and S3). In both empirical data and theoretical simulations, community
biomass had a positive effect average trophic level (resp.
$r_\delta =$ `r rdsp$tlvl$bm` and `r rdm$tlvl$bm`), and redundancy (resp.
$r_\delta =$ `r rdsp$red$bm` and `r rdm$red$bm`), as in temporal analysis.
We found weak evidence of a positive effect of spatial variation in community
biomass on connectance in empirical data ($r_\delta =$ `r rdsp$ct$bm`) but
strong evidence in theoretical simulation ($r_\delta =$ `r rdm$ct$bm`, Fig.
\@ref(fig:std)).

The effects of species richness were more complex, because species richness had
indirect effects on food-web structure through community biomass. Temporal
trends in species richness had a direct negative effect on connectance
($r_\delta =$ `r rdtps$ct$r`, Fig. \@ref(fig:sem)f), a positive
indirect effect through community biomass ($r_\delta =$ `r ridtps$ct$r`), so
that species richness temporal trends had a negative effect in total on temporal in
connectance ($r_\delta =$ `r rttps$ct$r`, Fig. \@ref(fig:std)).
In contrast, temporal trends in species had no direct effects on the temporal
trends in average trophic level ($r_\delta =$ `r rdtps$tlvl$r`, Fig.
\@ref(fig:sem)g), a positive indirect effect through community
biomass ($r_\delta =$ `r ridtps$tlvl$r`), which adds up to a positive total
effect of species richness temporal trends on average trophic level ($r_\delta
=$ `r rttps$tlvl$r`, Fig. \@ref(fig:std)). Species richness temporal trends had
direct and indirect positive effect on the temporal trends in redundancy (resp.
$r_\delta =$ `r rdtps$red$r`, `r ridtps$red$r`), which added up to a strong
positive effect ($r_\delta =$ `r rttps$red$r`). Furthermore, we found that
temporal trends of the proportion of piscivorous species was negatively
associated with the temporal trends of species richness (Fig. S2), meaning that
communities that increase in richness over time have an increase in the
proportion of non-piscivorous species over time and vice versa. It means
food-webs that have experienced declines in species richness over the last
decades have decreased in average trophic level and redundancy but increase in
connectance, and conversely, those that experienced increases in species
richness have increased in average trophic level, redundancy, but decreased in
connectance.

The effects of spatial variations in species richness on food-web structure
mostly agreed with the ones found with temporal trends, expect for average
trophic level. We found a direct negative effect of species richness on
connectance both using data and model simulations
(resp. $r_\delta =$ `r rdsp$ct$r` and `r rdm$ct$r`, Fig. \@ref(fig:std)), on
average trophic level (resp. $r_\delta =$ `r rdsp$tlvl$r` and `r rdm$tlvl$r`),
but a direct positive effect on redundancy (resp. $r_\delta =$ `r rdsp$red$r`
and `r rdm$red$r`). Species richness had indirect positive effect of species
richness through community biomass, which added up to strong total negative
effects on connectance in empirical data and simulations (resp. $r_\delta =$
`r rtsp$ct$r` and `r rtm$ct$r` ) and positive on redundancy (resp. $r_\delta =$
`r rtsp$red$r` and `r rtm$red$r`). We found that species richness had a total
positive effect on average trophic level in bioenergetic model, in agreement
with temporal trends (resp. $r_\delta =$ `r rtm$tlvl$r`), but a total negative
effect in spatial variations (resp. $r_\delta =$ `r rtsp$tlvl$r`).

We performed extinction simulations scenario on the metaweb, the food-web
including all the species and nodes found in the empirical data, to better
understand how species richness change affect food-web structure. When species
are removed in a sequential order beginning by the highest trophic levels,
connectance and average trophic level decreased as we removed species (red
lines, Fig. S4a). At the opposite, removing first low trophic level species first
increase connectance and average trophic level (green lines, Fig. S4a). In turn,
random species extinction sequence had little effect on food-web structure
(black lines, Fig. S4a).

```{r}
std_cap <- paste0(
  "Direct and total effects of community biomass and species richness on
  food-web structure. Standardized coefficients ($r_\\delta$) derived from the
  temporal, spatial, and bioenergetic model Structural Equation Models. Points
  and lines display the estimated effects and their confidence intervals at
  95\\%. Values are displayed in Table S3, R-squared in Table S4, and the
  visualisation of the bivariate relationships for the spatial and bioenergetic
  model SEMs are displayed in Figure S3.")

```

```{r std, fig.cap=std_cap, out.width = "100%"}
tar_load(p_semeff_tot_ok_file)
knitr::include_graphics(p_semeff_tot_ok_file)
```

# Discussion

We found remarkable heterogeneity in community biomass and species richness
changes over the last decades in riverine fish communities, linked to changes in
food-web structure. Community biomass changes spanned from `r grbmn`% to
+`r grbmx`% per decade, species richness from `r grrhn`% to +`r grrhx`% per
decade, in line previous reports of the heterogeneity of local biodiversity
changes in terrestrial, marine and freshwater realms[@dornelas_assemblage_2014;
@blowes_geography_2019; @danet_past_2024]. In turn, connectance temporal changes
spanned from `r grctn` to +`r grctx` per decade, average trophic level, from
`r grtln` to +`r grtlx` per decade and the proportion of redundant trophic links
from `r grln` to +`r grlx` per decades. We questioned the effects of community
biomass and species richness changes over time on temporal changes in food-web
structure, that we also tested across space and with simulations using a
bioenergetic food-web model. We found positive effects of species richness
changes on changes in community biomass that were consistent across temporal and
spatial analysis and with the theoretical model. The effects of community
biomass changes on food-web structure were also consistent across analysis,
suggesting that loss of community biomass are linked to the loss of the trophic
species of high trophic level. However, changes in species richness had
contrasting effects on food-web structure depending on the type of analysis,
temporal or spatial and theoretical simulations. It suggests that food-web
structural changes induced by changes in species richness differs from the one
induced by the changes in community biomass, and that they imply different
processes across time and space. However, all analysis converged to indicate
that community biomass and species richness declines trigger a decrease in
food-web redundancy. In what follows, we discuss how changes in community
biomass and species richness changes over time and space are related to food-web
assembly and disassembly processes.

We found that species richness scaled positively with community biomass, both in
temporal, spatial and theoretical simulations, in accordance with previous
analysis across space in stream communities [@woods_testing_2020;
@danet_species_2021] but also in grasslands, salt marches and forests
[@hector_plant_1999; @tilman_biodiversity_2014; @wu_relationship_2015;
@li_relationship_2018; but see @grace_integrative_2016]. This results is in
agreement with previous findings showing that food-web diversity can increase
biomass by stimulating primary production and resource uptake
[@wang_biodiversity_2018]. Our results suggest that food-web losing species
richness over time also experience a loss of biomass, while those gaining
species experience a increase in community biomass. This results is coherent the
Biodiversity-Ecosystem Functioning theory [@loreau_biodiversity_2001;
@wang_biodiversity_2018], and suggest that the loss of species richness and
biomass might impair ecosystem functioning.

Our results suggest that food-web that have undergone a biomass decline
experienced a loss of top trophic levels species and connectance, those gaining
biomass experienced a gain in top trophic levels and connectance. Importantly,
the same results were found using spatial comparison and theoretical models.
This results can be linked with the energetic and trophic biogeography theories
[@lindeman_trophic-dynamic_1942; @gravel_trophic_2011], which posits that the
presence of top predators is limited by the amount of energy and resource
available [@lindeman_trophic-dynamic_1942, @takimoto_ecosystem_2008;
@post_ecosystem_2000; @post_long_2002]. Thereby, a decrease in community biomass
can limit the amount of resource available to sustain top predators, while
increase in community biomass may result in more resource available for top
predators [@takimoto_ecosystem_2008; @post_ecosystem_2000]. (Add something about
the role of apex predators?)

Species richness temporal changes were positively associated to changes in
average trophic level through community biomass and in connectance mainly
through direct effects, such as food-webs that have suffered a decline in
species richness over last decades have experienced a concomitant decline in
average trophic level and increase in connectance, those have seen an increase
in species richness have experienced an increase in average trophic level and
decline in connectance. However, the effect of species richness temporal trends
on average trophic level were twice lower than the one of community biomass.
Altogether, it suggests that species richness increase and decrease were less
clearly linked to specific changes in top trophic levels than community biomass.
Species richness temporal trends could be more linked to changes in the
redundancy of trophic roles in the food-webs, such as it has little effects of
food-web structure but increases its community biomass. This interpretation
align with the insurance hypothesis [@yachi_biodiversity_1999;
@loreau_biodiversity_2021], such as species richness and associated redundancy
in species roles can increase ecosystem stability by buffering the impacts of
perturbations [@sanders_trophic_2018].

When considering spatial analysis and model simulations, species richness
effects on food-web structure mostly consistent from the temporal analysis. All
analysis converged to indicate that increases in species richness result in
decreases in connectance, a well documented pattern in empirical food-webs
[@dunne_network_2006; @danet_species_2021]. However, spatial analysis found that
species richness decrease average trophic level, while temporal analysis and
theoretical models found that species richness increases result in higher
average trophic level. This result is also in line with other studies which
found a mismatch between spatial and temporal trends [@adler_contrasting_2007;
@johnson_testing_2008]. This discrepancy in space and time is due to the strong
assumption that variations in space reflect differences in assembly stages of
ecological communities and then mirror their temporal assembly/disassembly
trajectories. Temporal changes in species richness might relate to short-term
dynamics while spatial gradients reflect aggregated dynamics over a longer time
scale including a far number of processes including high species turnover along
spatial gradients. Our results then suggest community biomass changes in space
might be a good surrogate to infer assembly and disassembly of food-web in time,
but that it is probably not always the case for species richness.

Our study demonstrates that redundancy of trophic pathways in food-webs are
positively associated to both community biomass and species richness changes, so
as declines in community biomass and species richness triggers a loss in the
proportion of redundant links in food-webs, while increases in community biomass
and species richness over time results in the increase in the proportion of
redundant links in food-webs. Redundant links are the trophic links that can be
removed with triggering secondary extinctions [@allesina_who_2004;
@bodini_using_2009], such an higher proportion of redundant interactions
increase robustness of food-webs to perturbations [@allesina_functional_2009].
It suggests that declines in community biomass and species richness result in a
decline of food-web robustness to further perturbations.

Our study brings evidences that temporal changes in community biomass and
species richness are related to changes in food-web structure in empirical
settings. We completed the temporal analysis with a spatial analysis and
theoretical simulations. Altogether, it suggests that community biomass changes
are associated with changes in the presence of top predators, in agreement with
the hypothesis that food-web maximum trophic level are limited by the amount of
energy available in lower trophic level. In contrast, species richness changes
seems to be linked to changes in upper and lower part of the food-webs.
Importantly, temporal and spatial analysis using empirical fish food-webs are
largely in agreement with theoretical simulations, which highlight the potential
generality of the food-web assembly rules that we report. The study finally
highlights that increases in community biomass and species brings robustness to
food-webs by increasing the proportion of redundant trophic links. Conversely, it
suggests that community declines in species richness and community biomass are
likely to decrease significantly the community robustness to further
perturbations. Thereby, the numerous reported declines in biodiversity might
hide increase in the fragility of ecosystems.

# Material & Methods

We used a monitoring of freshwater fishes conducted all over France and
conducted by the Office Français de la Biodiversité (OFB, previously known as
ONEMA and AFB). The sampling protocols were standardized since 1995 and a
complete description of the sampling process can be found in
@danet_species_2021. The fishes are sampled using
electrofishing all across the streams in shallow streams and over the bank by
boat in deeper streams. All the fishes are counted and the bodysize of fishes are
recorded for a sample of them, enabling the estimate the distribution of
individual bodysize by species for each sampling [@bonnaffe_comparison_2021;
@danet_species_2021].

We selected the sites that have been monitored at least 10 years on the
1995-2018 period, with the same sampling protocol over time. We selected
one sampling a year by site. In each site, we ensured low temporal heterogeneity
in sampling period by selecting sampling that were performed in the same period,
leading to a mean standard deviation of sampled month in each site of three
weeks [@danet_species_2021]. Those steps leads to the selection of
`r nb_st` sites (Fig. \@ref(fig:matmet), Table S1).

## Biomass

We first inferred the body mass of each fish individual considering an allometric
law, $B_i = 0.01 \times l_i^{3.03}$, $B_i$ being the biomass expressed in gram and $l_i$ the
body length of the individual $i$ expressed in millimeters [@froese_cube_2006]. Then the
biomass of each species and community biomass were obtained by summing the body
mass of individuals. Finally, the biomass was reported to the sampled surface
(in $m^2$) to account for sampling effort.

## Food-web inference

We inferred food-web structure in two steps [@bonnaffe_comparison_2021;
@danet_species_2021]: (1) we first computed the metaweb, describing the
potential trophic interactions among all size classes species, (2) then we
extracted each yearly snapshot of food-webs by extracting species size classes
and their interactions according to their presence in a given sampling event.

We inferred the metaweb according to ontogenic diet data and body size
[@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019;
@bonnaffe_comparison_2021; @danet_species_2021]. The method is extensively
described in @bonnaffe_comparison_2021 and @danet_species_2021. The food diet of
stream fishes changes along their life cycle or stage, therefore according to
their body size. Each fish species was divided in nine size classes, named
"trophic species", equally distributed in their observed body size range. Thus,
a trophic species is a group of individuals of the same species and size class
that are expected to share the same preys and predators because trophic
interactions are largely determined by predator-prey size ratio
[@brose_allometric_2006]. We also added seven nodes present in the ontogenic
food diet database: detritus, biofilm, phytoplankton, zooplankton, macrophages,
phytobenthos and zoobenthos.  The resulting metaweb contained 45 fish species
divided into nine size classes plus 7 resource nodes, summing up to 412 nodes.
Consideration of size classes, i.e.  trophic species instead of species,
improves inference of trophic interactions [@jennings_weak_2001].

We determined the feeding interactions of trophic species according to their
species identity and their body size. The diet database described two or three
life stages by fish species, each one being delimited by body size. The
ontogenic fish diet database was filled according to fishbase
[@froese_fishbase_2021] and literature [@bonnaffe_comparison_2021;
@danet_species_2021].

The feeding interactions among fishes were based on the body size ratio
between predators and preys. We defined the predation window of a
piscivorous trophic species as 3% to 45% of its midpoint
[@mittelbach_ontogeny_1998; @claessen_impact_2002; @hart_handbook_2008].  A
trophic link was defined between the piscivorous trophic species and all the
trophic species whose midpoint in body size range was included in the predation
window.  The fish-resource trophic links were set between the trophic species
and the resources nodes according to the food items of their size class.
Finally, the feeding interactions between resources were defined according to the
literature [@allan_stream_2007; @hart_handbook_2008].

## Community and food-web structure

At each site, yearly food-webs were described by their community biomass and
species richness. Yearly food-web structure was described by their connectance,
describing the degree of generalism in a network [@dunne_network_2006;
@delmas_analysing_2019]. It was computed as the ratio between the number of
actual links and the number of potential links ($C = \frac{L}{S(S-1)}$, with $L$
being the number of links and $S$ the number of trophic species). We also
computed the average trophic level, weighted by the biomass of the trophic
species. The trophic level of a trophic species was computed recursively as 1
added to the mean trophic level of the preys that it consumes
[@herendeen_ecological_2008]. We computed the number of redundant and functional
links using the dominator tree approach [@allesina_functional_2009]. A species
$i$ is a dominator of species $j$ if $i$ mediates the connection $j$ to basal
resources such as the extinction of $i$ remove the connection of $j$ to basal
resources. It is then to identify the set of trophic links that are essential to
connect species to basal resources, i.e. functional links, and the
redundant links. This method requires a single basal root in the food-webs. To
do so, we added a basal dummy resource node to the empirical food-webs connected
to biofilm and phytoplankton resource node. For food-webs generated from
theoretical simulations (see below), we added a resource node connected to all
primary producers. We computed the set of dominator trees using the
`dominator_tree()` function in `igraph` R package. We computed the number of
functional links as the number of links in the resulting dominator tree of the
food-web. We checked that the procedure reproduced the examples presented in
@allesina_functional_2009. We then derived the proportion of redundant links in
food-webs of empirical and theoretical simulations.

## Temporal trends of food-web

We modeled the temporal trends of food-web structure at each site ($Y_i$) using
linear hierarchical model. Time was considered as the number of year since the
first sampling ($t = 0$) at each site. We accounted for the spatial structure of
the data by adding random intercepts ($a$) and slopes ($b$) for the
basin ($a_n$, $b_n$) and site ($a_{i|n}$, $b_{i|n}$) identity, site effects
being nested in the basin one.

$$
\begin{aligned}
    Y_{i|n, t} &= \alpha + \beta Time_t + \epsilon_{i|n, t}
  (\#eq:gen)
\end{aligned}
$$

- $\alpha = \alpha_0 + a_n + a_{i|n}$
- $\beta = \mu + b_n + b_{i|n}$
- $a_n, a_{i|n}, b_n, b_{i|n}, \epsilon_{i|n, t} \sim \mathcal{N}(0,\,\sigma^{2})$
- $n$: basin, $i$: site $i$, $t$: time $t$

The model parameters were estimated using INLA (Integrated Nested Laplacian
Approximation) Bayesian inference with default uninformative priors
[@rue_bayesian_2017]. The prior distributions for the fixed intercept and slope were
modeled as a flat normal distribution with a mean of zero ($\mathcal{N}(0,\, 1000)$).
Prior distributions for the Gaussian error $\epsilon_{i|n, t}$ and the random
effects were modelled as a log gamma distribution with
shape and inverse scale parameters ($\mathcal{G}(s, \, \tau) = \mathcal{G}(1, \,
5.10^-5)$), then their respective standard deviation were obtained by
backtransforming the inverse scale parameter ($\sigma = 1 / \sqrt{\tau}$).
Community biomass and species richness were logged transformed.

The performance of the model was assessed visually by plotting the fitted versus
the observed values (Fig. S5). We also visually inspected the Posterior
Integral Transform (PIT) and Conditional Predictive Ordinate (CPO) to assess the
quality of the model fit. We also computed the marginal ($R^2_m$)and conditional
($R^2_c$) R-squared of the models following previous formulations
[@nakagawa_general_2013; @lahuis_explained_2014]. In order to account for the
variability associated with the priors [@gelman_r-squared_2019], we computed the
variance of the posterior distribution  of each predicted values ($Var_{fit} =
\sigma^2(\hat{Y_i})$). We computed $R^2$ values for each predicted values, then
obtaining a $R^2$ distribution [@gelman_r-squared_2019]. We reported mean
marginal and conditional $R^2$ associated the 95% credible interval computed
using the Highest Posterior Density method.

$$
R^2_m = \dfrac{Var_{fit}}{Var_{fit} + Var_{res}} = \dfrac{\sigma^2(\hat{Y_i})}{\sigma^2(\hat{Y_i}) + \sigma^2(Y_i - \hat{Y_i})}
$$

$$
R^2_c = \dfrac{Var_{fit} + (a_n)^2 + (a_{i|n})^2}{Var_{fit}  + (a_n)^2 + (a_{i|n})^2 + Var_{res}}
$$

$y_i$ and $\hat{y_i}$ being respectively the observations and the
predicted values, $Var_{fit}$ and $Var_{res}$ being respectively the variance of
predictive means and the variance of the residuals. $a_n$ and $a_{i|n}$ are
respectively the standard deviation on the random intercept associated to the
basin and the site.

We extracted the temporal trends at each site from the hierarchical model
using mean BLUP values (Best Linear Unbiaised Prediction), then multiplied
by 10 to get temporal trends by decade. For logged transformed biomass and
species richness, we derived percentage change by decade by transforming the
slopes ($\beta\prime = (e^{\beta } - 1) \times 100 \times 10$).

Overall, we found that standard deviation attributed to site effects in both
baseline and temporal trends in community structure were much higher than the
ones found for hydrographic basin (Table S5).

## Spatial

We assessed the validity of the space-for-time substitution hypothesis by
comparing the relationships estimated among food-web structure in time to the
ones inferred in space. We used the same sites as in the temporal analysis,
but we summarised temporal series of each station by the median value over time
of each facet of food-web structure described above (Fig. \@ref(fig:matmet)).

## Statistical analysis

We modelled the relationships among food-web structure, that is
connectance, average trophic level, the proportion of redundant trophic links, and
community biomass ($dB$) and species richness ($dR$) with Structural Equation
Models. We modelled species
richness as related to community biomass and food-web structure. In
turn community biomass was related to food-web structure. We computed one
structural equation model to assess the relationships among the temporal trends
and one for the relationships over the spatial gradient, we added a random effect
of the hydrographic basin on the intercept for the latter one.

In order to control for the uncertainties of the estimated temporal trends, we
weighted each temporal trend by the inverse of the sum of the standard deviation
of the estimated trends, i.e. the estimated slopes for the response and the
explicative variables ($w_j = 1 / \sum{e_j}$, $w_j$ being the weight, and
$s_{e_j}$ the standard deviation on the temporal trends of the variables).

The coefficients of the Structural Equation Models were computed with
PiecewiseSEM R package [@lefcheck_piecewisesem_2016]. The coefficients were
standardized and the confidence intervals were computed with semEff R package
[@murphy_semeff_2021]. Confidence intervals were obtained by bootstrapping the
linear models 1000 times and computing adjusted bootstrap percentile interval
(BCa), which are adjusted for the asymmetric distribution of resampling. The marginal
and conditional $R^2$ associated to the linear models of the SEMs were computed
with the Nagelkerke method [@lefcheck_piecewisesem_2016, Table S4].

All the statistical models displayed low multicollinearity (Variance inflation
factors < 3, Table S6). We visually inspected the residuals of the linear
models.

## Extinction simulation in the metaweb

We simulated extinction sequences in the metaweb according to three scenarios to
get expectations on how the food-web structure should change according to the
addition and deletion of species and nodes. The scenario of extinction sequences
were related to the trophic level of the species/nodes. In the decreasing and
increasing scenario, we removed species/nodes in decreasing, resp. increasing
order, of their trophic level, i.e. respectively beginning by the highest and
lowest trophic level. In the random extinction sequence, we built 50 random
extinction sequences. For each simulation, we removed fish species/nodes that
lost all their preys. We then computed the connectance, the average and the
maximum trophic level after each extinction.

## Theoretical simulations

```{r}
tar_load(sim_param)
nsim <- nrow(sim_param)
nfw <- length(unique(sim_param$fw_id))
s_range <- c(min = min(sim_param$S), max = max(sim_param$S))
c_range <- c(min = min(sim_param$C), max = max(sim_param$C))
k_range <- c(min = min(sim_param$K), max = max(sim_param$K))
```

We used a bioenergetic food-web model to simulate the effects of community
biomass and species richness changes on food-web structure. This food-web model
lifts one of the limitation of our data, that is it includes the biomass of all
the food-webs from primary producers to top predators. This model allows to
simulate complex communities because its parametrization depends mainly on
species bodymasses. We first generated `r nfw` food-webs varying in species
richness (range: `r s_range["min"]`-`r s_range["max"]`)
and connectance (range: `r c_range["min"]`-`r c_range["max"]`) using the niche
model algorithm, which reproduces food-web topology tipically found in empirical
food-webs [@williams_simple_2000]. We set species bodymasses ($M_i$) using a predator-prey
mass ratio of 100 ($Z$) [@brose_allometric_2006; @brose_predator_2019], such as $M_i =
Z^{(TL_i - 1)}$, $TL_i$ being the trophic level of the species $i$. The
metabolic rates of the species ($x_i$) were then derived from allometric
relationships between metabolism and bodymass empirically estimated
[@brose_allometric_2006], such as $x_i = aM_i^b$, with $a = xx$ and $b = -0.25$.

The dynamics of species biomasses are described by an ODE systems of two
equations, for primary producers (eq. \@ref(eq:prod)) and consumers (eq. \@ref(eq:cons)) respectively:

\begin{equation}
    (\#eq:prod)
	\frac{dB_i}{dt} =
    r_i B_i G_i
    - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
\end{equation}

\begin{equation}
    (\#eq:growth)
    G_i = (1 - \dfrac{\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j}{K_i})
\end{equation}

The primary producer biomass increase with a growth rate ($r_i$) of 1 and a
logistic growth (eq. \@ref(eq:growth)), growth being limited by intraspecific
($\alpha_{ii}$, set to 1), interspecific ($\alpha_{ij}$, set to 0.5) competition,
and carrying capacity ($K_i$, varying in the simulations). Primary producers
loses biomass by being eaten by consumers at a rate that depend on consumer
metabolic rates ($x_j$), attack rates ($y_j$, set to 8 @delmas_analysing_2019),
a functional response (F_{ji}), and assimilation efficiency ($e_{ij}$, set to
0.45 for herbivores and 0.85 for carnivores, @delmas_analysing_2019).

\begin{equation}
    (\#eq:cons)
        \frac{dB_i}{dt} =
        \sum_{j \in \{\mathrm{res}\}_i} x_i y_i B_k F_{ij}
        - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
        - x_i B_i
\end{equation}

Consumers gain biomass by eating other species and lose biomass by being
consumed and by metabolic loses ($x_i$, eq. \@ref(eq:cons)). The functional
response (eq. \@ref(eq:holling)) of consumer feeding depends on the preference
for its preys ($\omega_{ij}$, set even across consumer preys) and
half-saturation rate ($B_0$, set to 0.5). The hill coefficient ($h$, set to 1.5)
determines the shape of the functional response (Holling's type II or III).

\begin{equation}
    (\#eq:holling)
    F_{ij} = \frac{\omega_{ij} B_j^h}
    {B_0^h + \sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h}
\end{equation}

We simulated the food-webs dynamics across a gradient of global carrying capacity (K,
range: `r k_range["min"]`-`r k_range["max"]`). The carrying capacity were
standardized according to the number of producers and interspecific competition
($K_i = \frac{1 + \alpha_{ij} (S - 1)}{S}K$) to prevent changes in productivity
due to system size [@ives_stability_2000; @loreau_species_2008]. We ran
simulations for 5000 timesteps. At the end of simulations, we removed dead
species and measured community biomass, species richness and food-web structure.
If the final community contained disconnected producers (i.e. without consumers)
or disconnected consumers (i.e. consumers without preys), we discarded them and
run the simulations again with the same parametrisation. The simulations were
done using the Julia package `EcologicalNetworksDynamics.jl`
[@lajaaiti_ecologicalnetworksdynamicsjl_2024].

# References  {-}

<div id="refs"></div>

\newpage