### 2025-02-03 paper/appendixB.Rmd
Manuscript update before Maud & Colin comments

---
title: "Biodiversity changes are linked to changes in food-web structure and redundancy in riverine ecosystems"
subtitle: "Alain Danet (a,b), Loubna El Madouri (a,c), Andrew P. Beckerman (b),
Maud Mouchet (a), Colin Fontaine (a) and Elisa Thébault (d)"
date: Appendix B - Computation of redundancy of trophic pathways in food webs
output:
  bookdown::pdf_document2:
    fig_caption: true
    keep_tex: true
    toc: false
fontsize: 10pt
header-includes:
   - \newcommand{\beginsupplement}{\renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
   - \renewcommand{\arraystretch}{1}
geometry: margin=2cm
bibliography: references.bib
number_sections: false
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.align = "center",
  fig.fullwidth = TRUE,
  fig.show = "hold",
  fig.pos = "H",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE
)
options(
  #kableExtra.latex.load_packages = FALSE,
  knitr.table.format = "latex",
  knitr.graphics.rel_path = FALSE
)
```

```{r, echo=FALSE, message=FALSE, results="hide"}
library(targets)
library(tidyverse)
library(magrittr)
library(cowplot)
library(kableExtra)
library(igraph)

mypath <- rprojroot::find_package_root_file
source(mypath("R/variable_shortcut.R")) # Define your custom code as a bunch of
#functions.
source(mypath("R/misc.R"))
source_dir(mypath("R")) # Define your custom code as a bunch of functions.

mytheme <- theme_cowplot() +
  background_grid()

theme_set(mytheme)
old_par <- par()
```

\beginsupplement

Because species biomass are involved in trophic interactions, species losses can
have cascading effects and trigger secondary extinctions [@dunne_network_2002].
In the context of biodiversity crisis, assessing the robustness of ecological
networks to species losses has been then a big question in Ecology. Previous
research showed that food-web structure, such as connectance, increased
robustness [@dunne_network_2002]. The idea is that more connected food-web will
be more stable because they contain a large number of pathways convey biomass
through the food-web. However, connectance as some limitations as a robustness
metric, because each trophic link does not contribute equally to the robustness
of an ecological network [@allesina_who_2004]. Trophic links can be categorised as
functional or redundant, depending on their role in transmitting biomass through
the food-web. Functional links are the set of links that is required to connect
every species to the basal resource of the food-web, all of the other being
redundant. Only functional links affect directly robustness, as their removal
will trigger secondary extinctions. The removal of redundant links in turn do
not directly trigger secondary extinctions, but makes the food-webs more fragile
to further extinctions [@allesina_functional_2009].

The identification of functional and redundant links is based on the use of
denominator trees [@bodini_using_2009; @allesina_who_2004;
@allesina_functional_2009]. Dominator trees allows to identify the set of
trophic links that are essential to connect all the consumers to the basal resource, i.e.
the functional links, and the redundant links, i.e. the redundant trophic pathways
from basal resource to consumers. Pioneer papers explained the algorithm to
compute the set of dominator trees and we refer to them for further
information and examples [@bodini_using_2009; @allesina_who_2004;
@allesina_functional_2009]. Here, we aim to show that while implementing this
analysis could be difficult at the time, this algorithm is now widely
available. In R, it is available in `igraph::dominator_tree()`. Below we show
how the number of functional and redundant links can be  easily computed from R
by replicating three examples presented in previous studies.

```{r, echo = TRUE}
compute_redundancy
```

Here is an example with the food-web displayed in appendix of
@bodini_using_2009:

```{r, echo = TRUE}
# Get the food-web example from Bodini et al. (2009), Appendix 1 
g <- get_bodini_graph_appendix()
# compute reduncancy
out <- compute_redundancy(g, root = "R")
as.data.frame(out)
```

```{r}
vertex_color <- "white"
vertex_size  <- 40
```

We used this function to reproduce the three examples (Fig. \@ref(fig:examples)),
which were displayed in the Fig. 1 and Appendix 1 of @bodini_using_2009, and
in the Fig. 1 of @allesina_functional_2009. The table \@ref(tab:redun) displays
the number and the proportion of functional and redundant links from those
examples.

```{r bod1}
bodinifig1 <- get_bodini_graph_fig1()

# Complete graph
p_bodinifig1_all <- ~{
  lg <- layout_as_tree(bodinifig1, root = "root")
  lg[, 2] <- -lg[, 2]
  plot(bodinifig1, layout = lg, vertex.color = vertex_color, vertex.size = vertex_size)
}

# Only functional links
bodinifig1_dtree <- dominator_tree(bodinifig1, root = "root")
p_bodinifig1_functional <- ~{
  layout <- layout_as_tree(bodinifig1_dtree$domtree, root = "root")
  layout[, 2] <- -layout[, 2]
  plot(bodinifig1_dtree$domtree, layout = layout, vertex.color = vertex_color, vertex.size = vertex_size)
}
```

```{r bod2, echo = FALSE, results = "hide"}
g <- get_bodini_graph_appendix()

# Complete graph
p_bodiniappen_all <- ~{
  lg <- layout_as_tree(g, root = "R")
  lg[, 2] <- -lg[, 2]
  plot(g, layout = lg, vertex.color = vertex_color, vertex.size = vertex_size)
}

# Only functional links
dtree <- dominator_tree(g, root = "R")
p_bodiniappen_functional <- ~{
  lg <- layout_as_tree(dtree$domtree, root = "R")
  lg[, 2] <- -lg[, 2]
  plot(
    dtree$domtree,
    layout = lg,
    vertex.color = vertex_color,
    vertex.size = vertex_size
  )
}

p_bodini <- plot_grid(
  ggdraw(p_bodinifig1_all), ggdraw(p_bodinifig1_functional),
  ggdraw(p_bodiniappen_all), ggdraw(p_bodiniappen_functional),
  labels = "auto"
)
ggsave_multiple(
  paste0("p_bodini", c(".png", ".pdf")),
  plot = p_bodini,
  path = here::here("figures"),
  scale = 2.6,
  units = c("mm"),
  width = 80,
  height = 80 * 1.5)
```

```{r all1, echo = FALSE, results = "hide"}
allesinafig1 <- get_allesina_graph(diag = FALSE)

p_allesinafig1_all <- ~{
  lg <- layout_as_tree(allesinafig1, root = "0")
  lg[, 2] <- -lg[, 2]
  plot(allesinafig1, layout = lg, vertex.color = vertex_color, vertex.size = vertex_size)
}

# Only functional links
allesinafig1_dtree <- dominator_tree(allesinafig1, root = "0")
p_allesinafig1_functional <- ~{
  layout <- layout_as_tree(allesinafig1_dtree$domtree, root = "0")
  layout[, 2] <- -layout[, 2]
  plot(
    allesinafig1_dtree$domtree,
    layout = layout,
    vertex.color = vertex_color,
    vertex.size = vertex_size
  )
}

p_examples <- plot_grid(
  ggdraw(p_bodinifig1_all), ggdraw(p_bodinifig1_functional),
  ggdraw(p_bodiniappen_all), ggdraw(p_bodiniappen_functional),
  ggdraw(p_allesinafig1_all), ggdraw(p_allesinafig1_functional),
  nrow = 3,
  labels = "auto"
)
ggsave_multiple(
  paste0("p_examples", c(".png", ".pdf")),
  plot = p_examples,
  path = here::here("figures"),
  scale = 4.0,
  units = c("mm"),
  width = 80,
  height = 80 * 1.2)
```

(ref:example-cap) Reproduced examples of categorising functional links in
food webs using `igraph::dominator_tree()` R function. Left columns (a, c, e)
displays the complete food-web while right column displays food-web with the
functional links only (b, d, f). (a,b): reproduction of Fig. 1 of
@bodini_using_2009, (c, d): reproduction of appendix figures of @bodini_using_2009,
(e, f): reproduction of fig. 1 of @allesina_functional_2009. Note that (e, f) is
a bit different from the original, as noted by the authors, they could have
removed further redundant links (i.e. $2 \to 5$, or $3 \to 5$).

```{r examples, fig.cap = "(ref:example-cap)", out.height = "80%"}
knitr::include_graphics(here::here("figures", "p_examples.png"))
```

```{r}
examples <- list(
  bodiniappendix = g,
  bodinifig1 = bodinifig1,
  allesinafig1 = allesinafig1
)
redundancy_examples <- map2_dfr(
  examples,
  list(bodiniappendix = "R", bodinifig1 = "root", allesinafig1 = "0"),
  ~compute_redundancy(.x, root = .y), .id = "reference")
```

```{r}
paper_replacement <- c(
  bodinifig1 = "Bodini et al. (2009) Fig. 1",
  bodiniappendix = "Bodini et al. (2009) Supp material 1",
  allesinafig1 = "Allesina et al. (2009) Fig. 1"
)

redundancy_var_replacement <- function() {
  c(
    reference = "References",
    functional_links = "Nb functional links",
    redundant_links = "Nb redundant links",
    total_links = "Nb links",
    prop_functional_links = "Prop. functional links",
    prop_redundant_links = "Prop. redundant links"
  )
}

redundancy_examples %>%
  select(reference, total_links,
    functional_links, redundant_links,
    prop_functional_links, prop_redundant_links
  ) %>%
  mutate(reference = paper_replacement[reference]) %>%
  rename_with(~redundancy_var_replacement()[.x]) %>%
  kable(
    digits = 2,
    row.names = FALSE,
    label = "redun",
    booktabs = T,
    caption = "Computation of the number (Nb) of functional, redundant links, and
    their proportion (Prop.) from previous published food-webs.",
    ) %>%
  collapse_rows(columns = c(1), valign = "top") %>%
  kable_styling(latex_options = "scale_down")
```

# References {-}