### 2025-02-07 paper/manuscript.Rmd
Improve hierarchy and clarity of the objectives and findings:
- Improve hierarchy
- Mirror discussion and intro
- Mv details of spatial analysis to appendix

---
title: "Biodiversity temporal trends are reshaping food web structure and redundancy in riverine ecosystems"
subtitle: "Alain Danet (a,b), Loubna El Madouri (a,c), Andrew P. Beckerman (b),
Maud Mouchet (a), Colin Fontaine (a) and Elisa Thébault (d)"
date: "Research report"
output:
  officedown::rdocx_document:
    number_sections: false
    base_format: "bookdown::word_document2"
always_allow_html: true
number_sections: false
bibliography: references.bib
link-citations: true
#csl: ecology-letters.csl
#https://pkgs.rstudio.com/rticles/articles/examples.html
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
csl: pnas.csl
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path = "figures/"
)
options(
  knitr.table.format = "latex",
  knitr.graphics.rel_path = FALSE
)
```

```{r, results = FALSE}
library(targets)
library(tidyverse)
library(magrittr)
library(cowplot)
library(viridis)
library(kableExtra)
library(sf)
mypath <- rprojroot::find_package_root_file
source(mypath("R/variable_shortcut.R"))
source(mypath("R/misc.R"))
source_dir(get_mypath("R")) # Define your custom code as a bunch of functions.

mytheme <- theme_cowplot() +
background_grid()

theme_set(mytheme)
```

```{r}
tar_load(full_data2)
nb_st <- unique(full_data2$station) %>% length
nb_op <- nrow(full_data2) %>% round(., -3)

rm(full_data2)
tar_load(glob_tps_trends_decade_no_drivers)
```

- Corresponding author: Alain Danet (a.h.danet@sheffield.ac.uk)

- Author affiliations:

(a) Centre d’Ecologie et des Sciences de la Conservation, UMR 7204
MNHN-CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de Paris, 43
rue Buffon, 75005 Paris, France

(b) School of Biosciences, Alfred Denny Building, University of Sheffield,
Western Bank, Sheffield S10 2TN, United Kingdom

(c) Laboratoire d’Écologie Fonctionnelle et Environnement, UMR 5245
CNRS-UT3-INPT, Avenue de l’Agrobiopôle, Toulouse, France

(d) Sorbonne Université, CNRS, IRD, INRAE, Université Paris Est Créteil, 15
Institute of Ecology and Environmental Sciences of Paris (iEES-Paris), Paris,
France

- Classification: Biological Sciences (major) - Ecology (minor)

- Keywords: food web structure, biodiversity changes, redundancy, biodiversity &
  ecosystem functioning

# Abstract (229 / 250 words max)

Biodiversity changes are expected to profoundly impact ecosystem functioning and
stability, yet empirical evidence remains limited. Bridging this gap requires
aligning biodiversity trends with food web theory and Biodiversity-Ecosystem
Functioning research, which provide strong predictions about the ecological
consequences of biodiversity loss. Most insights rely on spatial comparisons of
ecological communities, assuming they reflect temporal trends—a space-for-time
substitution. Here, we analyze biodiversity and food web time series from over
400 riverine fish communities across France (1995–2018) to evaluate how
biodiversity trends reshape food web structure. We examine relationships among
species richness, community biomass, and three key food web metrics:
connectance, weighted average trophic level, and trophic pathway redundancy. We
also test the space-for-time hypothesis using spatial gradients and a
theoretical food web model. Structural equation models and robustness analyses
reveal that declines in species richness strongly correlate with reduced
community biomass, with top trophic-level species playing a disproportionate
role in food web changes. Declining biomass leads to the loss of top trophic
levels and decreased connectance, while species richness negatively correlates
with connectance. Both biomass and species richness declines reduce trophic
pathway redundancy, suggesting biodiversity loss undermines community robustness
to future perturbations. Our findings align with food web theory and
Biodiversity-Ecosystem Functioning research, support the validity of
space-for-time approaches for basic food web metrics, and highlight the utility
of food web metrics as indicators of ecosystem function and fragility.

# Significance (118 / 120 words max)

Understanding how biodiversity loss affects ecological communities is crucial
for predicting and mitigating human impacts on ecosystems. This study provides
comprehensive evidence on how biodiversity changes are reshaping food web structure using
an extensive dataset of riverine fish communities. We demonstrate that changes
in species richness and community biomass consistently alter trophic levels,
connectance, and the redundancy of feeding relationships. Our findings show how
energetic constraints and biodiversity-ecosystem functioning mechanisms jointly
shape food web assembly and disassembly. Importantly, we show that declining
biodiversity reduces the redundancy of trophic pathways, potentially making
communities more vulnerable to future perturbations. These results suggest that
food web structure metrics could serve as valuable indicators for monitoring
ecosystem health and evaluating conservation strategies.

# Introduction (487 words)

Human pressures are driving profound biodiversity changes, including freshwater
and terrestrial biodiversity declines linked to land degradation
[@newbold_global_2015; @cardinale_is_2018; @danet_past_2024] and reductions in
marine fish biomass due to intensified fishing pressures on predatory species
[@christensen_century_2014]. A central goal of Biodiversity-Ecosystem
Functioning and food web research has been to understand the ecological
consequences of these widespread biodiversity losses for ecosystem functioning
and stability [@loreau_biodiversity_2001; @loreau_ecological_2022;
@thebault_food-web_2003; @schneider_animal_2016; @dunne_network_2002;
@sole_complexity_2001; @allesina_functional_2009; @montoya_press_2009; @sanders_trophic_2018]. However,
research on biodiversity changes remains largely disconnected from studies
examining their ecological impacts, primarily due to a lack of simultaneous
temporal data on biodiversity and food web dynamics [@danet_species_2021]. As a
result, much of our understanding of the ecological consequences of biodiversity
changes over time relies on experimental and theoretical approaches. These
approaches often depend on the critical assumption that spatial comparisons of
ecological communities can serve as a proxy for temporal trends—a space-for-time
substitution—when direct temporal data are scarce. To advance our understanding
of the ecological consequences of biodiversity changes, it is essential to
investigate how these changes are reshaping food web structure in natural
settings.

Here, we bridge the gap between biodiversity changes and food web dynamics by
leveraging a unique dataset of riverine fish communities sampled repeatedly over
a 23-year period (1995–2018) across 403 locations in France. By reconstructing
yearly food web structures [@danet_species_2021; @bonnaffe_comparison_2021], we
investigate how temporal biodiversity trends reshape food web metrics such as
connectance, weighted average trophic level, and the proportion of redundant
trophic links (hereafter “redundancy”). Linking these findings to established
theories of Biodiversity-Ecosystem Functioning and food web dynamics, we address
three key objectives: (i) documenting how temporal biodiversity trends reshape
food web structure, (ii) connecting these changes to ecosystem functioning and
food web theories, and (iii) assessing the validity of the space-for-time
substitution assumption in the context of biodiversity and food web structure
changes.

The structure of ecological networks, which describes the number, distribution,
and strength of interactions among species, is critical for understanding
species coexistence, ecosystem functioning, and stability to perturbations
[@chesson_interaction_2008; @thebault_food-web_2003; @oconnor_general_2017;
@danet_species_2021; @eschenbrenner_diversity_2023; @danet_response_2024]. Both
theory and empirical evidence demonstrate that species richness and food web
structure influence coexistence [@chesson_interaction_2008], ecosystem
productivity [@thebault_food web_2003; @wang_biodiversity_2018], temporal
stability [@danet_species_2021; @danet_response_2024;
@eschenbrenner_diversity_2023], and robustness to disturbances
[@sole_complexity_2001; @dunne_network_2002; @allesina_functional_2009;
@bodini_using_2009; @montoya_press_2009; @sanders_trophic_2018]. For instance,
keystone predators can regulate dominant competitors, preventing competitive
exclusion and maintaining species richness [@heithaus_predicting_2008;
@johnson_rarity_2007; @estes_trophic_2011]. Similarly, food webs with higher
trophic levels are associated with greater ecosystem productivity
[@wang_biodiversity_2018; @schneider_animal_2016; @wang_biodiversity_2018;
@wang_intraguild_2019; @nie_will_2023] and community stability
[@danet_species_2021; @eschenbrenner_diversity_2023; @danet_response_2024],
while the number and distribution of trophic links determine resilience to
perturbations [@dunne_network_2002; @allesina_who_2004; @bodini_using_2009;
@allesina_functional_2009; @sanders_trophic_2018]. Understanding how
biodiversity changes reshape ecological networks is therefore essential for
quantifying and mitigating anthropogenic impacts on ecosystems.

Three key components link biodiversity changes to food web structure. First, the
energetic paradigm of food web dynamics posits that the presence of top
consumers depends on sufficient resource availability at lower trophic levels,
driven by bottom-up productivity [@lindeman_trophic-dynamic_1942;
@gravel_trophic_2011]. Thus, the biomass of generalist top consumers is likely
constrained by overall community biomass. Second, connectance—the proportion of
realized trophic links—typically decreases with increasing species richness, as
consumers are limited in the number of resources they can exploit
[@oconnor_general_2017; @dunne_network_2006]. However, higher species richness
is expected to increase the redundancy of trophic pathways, enhancing ecosystem
robustness to species extinctions [@dunne_network_2002; @allesina_who_2004;
@bodini_using_2009; @allesina_functional_2009]. This redundancy underpins the
insurance effect of biodiversity, where greater diversity buffers ecosystems
against perturbations and environmental changes [@yachi_biodiversity_1999;
@loreau_ecological_2022]. Finally, biodiversity-ecosystem functioning theory
suggests that species richness promotes community biomass through niche
complementarity [@loreau_biodiversity_2001]. In food webs, the presence of
higher trophic levels reduces niche overlap among producers, further boosting
productivity and biomass [@schneider_animal_2016; @wang_biodiversity_2018;
@wang_intraguild_2019; @nie_will_2023]. Together, these mechanisms highlight how
biodiversity changes can alter trophic link distribution, robustness, and
ecosystem functioning.

To achieve these objectives, we analyzed temporal trends in species richness,
biomass, and food web structure across `r nb_op` community samplings (Fig.
\@ref(fig:matmet)a-c, Table S1). Using structural equation modeling (SEM), we
inferred the relative effects of biomass and species richness trends on food web
metrics, as well as the effect of species richness on biomass (Fig.
\@ref(fig:matmet)d). We also conducted a robustness analysis on the meta-food
web (comprising all trophic species and their interactions) to identify species
extinction scenarios that could explain observed variations in food web
structure. To test the space-for-time hypothesis, we compared temporal findings
with spatial gradients using two approaches: (i) replicating our temporal
analysis using median biodiversity and food web metrics across sites (Fig.
\@ref(fig:matmet)c & e), and (ii) simulating food web dynamics across gradients
of species richness and primary productivity using a bioenergetic model
[@brose_allometric_2006; @lajaaiti_ecologicalnetworksdynamicsjl_2025]. These
simulations allowed us to evaluate whether theoretical predictions align with
our empirical data, which are limited to fish biodiversity.

```{r}
fig_mat_cap <- paste0("Methodological Approach: (a) We analyzed ", nb_st, "
  stream fish communities in France, each monitored for at least 10 years. (b)
  For each site and sampling event, we reconstructed food webs using a metaweb
  based on ontogenetic diet information and predator-prey body size ratios. (c)
  We computed temporal trends and median values for community biomass, species
  richness, connectance, average trophic level, and redundancy at each site. (d)
  We then assessed relationships among these food web descriptors over time and
  (e) across spatial gradients.")
```

```{r matmet, fig.cap=fig_mat_cap, out.width="100%"}
knitr::include_graphics(get_mypath("figures", "mat_met.png"))
```

# Results

```{r, results=FALSE}
gbm <- get_effect_ci(resp = "log_bm_std", ci_lvl = "level:0.80", p = TRUE)
gr <- get_effect_ci(resp = "log_rich_std", ci_lvl = "level:0.80", p = TRUE)
gt <- get_effect_ci(resp = "w_trph_lvl_avg", ci_lvl = "level:0.80", p = FALSE)
gct <- get_effect_ci(resp = "connectance", ci_lvl = "level:0.80", p = FALSE)
gl <- get_effect_ci(resp = "prop_redundant_links", ci_lvl = "level:0.80", p = FALSE)
```

```{r}
tar_load(c(r2_inla_tab, r2_inla, r2_sem))
r2_inla_report <- r2_inla %>%
  select(response, r2_mvp_marg95hpdci, r2_mvp_cond95hpdci, r2_mvp_cond_mean)
r2cond <- map(
  setNames(r2_inla_report$response, r2_inla_report$response),
  ~round(r2_inla_report[r2_inla_report$response == .x, ]$r2_mvp_cond_mean, 2)
)
r2ic <- map(
  setNames(r2_inla_report$response, r2_inla_report$response),
  ~r2_inla_report[r2_inla_report$response == .x, ]$r2_mvp_cond95hpdci
)
```

```{r}
tar_load(summary_pred_station_tps10)
dfv <- function(
  x = summary_pred_station_tps10,
  r = "connectance",
  v = "min",
  d = 4
) {
  out <- x[x$response == r, ][[v]]
  format(round(out, d), nsmall = d)
}
grbmn <- dfv(r = "log_bm_std", v = "min", d = 0)
grbmx <- dfv(r = "log_bm_std", v = "max", d = 0)
grrhn <- dfv(r = "log_rich_std", v = "min", d = 0)
grrhx <- dfv(r = "log_rich_std", v = "max", d = 0)
grctn <- dfv(r = "connectance", v = "min", d = 2)
grctx <- dfv(r = "connectance", v = "max", d = 2)
grtln <- dfv(r = "w_trph_lvl_avg", v = "min", d = 2)
grtlx <- dfv(r = "w_trph_lvl_avg", v = "max", d = 2)
grln <- dfv(r = "prop_redundant_links", v = "min", d = 2)
grlx <- dfv(r = "prop_redundant_links", v = "max", d = 2)
```

## Estimating temporal trends

Our food web reconstruction delivered `r nb_op` annual estimates of species
richness, biomass, connectance, weighted average trophic level, and the
proportion of redundant trophic links (hereafter “redundancy”, see Appendix B)
from our `r nb_st` fish communities with 10+ years of population dynamics
spanning 1995-2018 over France (Fig. \@ref(fig:matmet)a-c). We implemented Bayesian hierarchical
models to jointly estimate the temporal trends in these metrics. We found a
great heterogeneity in the temporal trends at the site level (community biomass:
from `r grbmn`% to +`r grbmx`% per decade, species richness: `r grrhn` to +`r
grrhx`%, average trophic level: `r grtln` to +`r grtlx`, connectance: `r grctn`
to +`r grctx`, redundancy: `r grln` to +`r grlx`, Fig. S1) which was effectively
captured by the hierarchical model ( conditional R2 ranging from `r
r2cond$log_bm_std` for community biomass to `r r2cond$log_rich_std` for species
richness, Table S2).

```{r}
sem_cap <- "(a) Structural equation models linking the temporal trends of
biodiversity and food web structure. Dashed and solid arrows display
respectively negative and positive the standardized slope coefficients, their
widths being proportional to the absolute values of the coefficients.. The
standardized slope coefficients and their associated 95% confidence intervals
obtained by bootstrapping are displayed along the path. R2 displays the marginal
R-squared of the linear models. (b to h) Direct bivariate relationships derived
from the structural equation model. Points display the temporal trends at each
site estimated with a hierarchical Bayesian model, lines display the linear
relationship estimated with the SEM. The distribution of the site temporal
trends are displayed in Fig. S1. All the effects are reported in Table S3.
Average trophic level is weighted by the biomass of the trophic species.
Redundancy is the proportion of redundant links."

```

```{r sem, fig.cap = sem_cap, out.width = "100%"}
tar_load(p_sem_tps_file)
knitr::include_graphics(p_sem_tps_file)
```

```{r}
tar_load(c(semeff_tot, semeff_tot_ok))
semeff_tot <- semeff_tot %>%
  mutate(across(c(response, predictor), ~str_remove(.x, "mean_")))

rdtps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std")
    )
)
rdsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", model = "spatial"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", model = "spatial")
    )
)
rdm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Direct", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Direct", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
```

```{r}
ridtps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "indirect")
    )
)
ridsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "indirect")
    )
)
ridm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Indirect", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Indirect", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
rttps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "total")
    )
)
rtsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std",
      model = "spatial", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "total")
    )
)
rtm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Total", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Total", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
```

## Association between biodiversity and food web temporal trends

These temporal trend estimations at the site level were the input to our SEM,
allowing us to evaluate the direct effect of the temporal trends of species
richness and community biomass on the temporal trends of food web structure, the
direct effect of  species richness temporal trends on the temporal trends of
community biomass, and the indirect effect of species richness temporal trends
on the temporal trends in food web structure through community biomass. We
detected a strong positive effect of species richness temporal trends on the
temporal trends in community biomass ($r_\delta =$ `r rdtps$bm$r`,
Fig. \@ref(fig:sem) a and b, $r_\delta$: standardized slope coefficient, [95%
CI], Table S3). This indicates that communities with decreasing, or increasing,
species richness over time experienced a concomitant decrease, or increase, in
community biomass.

We also find a positive effect of the temporal trend in community biomass on the
temporal trends of connectance, average trophic level, and redundancy (resp.
$r_\delta =$ `r rdtps$ct$bm`, `r rdtps$tlvl$bm`, `r rdtps$red$bm`, Fig.
\@ref(fig:sem)b-d). Thus, when communities experienced a decrease, or increase,
in community biomass, they also experienced a concomitant decrease, or increase,
in connectance, average trophic level and redundancy. To assess whether top
trophic level species were driving those relationships, we additionally looked
at the temporal trends in the number and the proportion of piscivorous species.
The temporal trends in community biomass were positively associated with both
the temporal trends in the number and proportion of piscivorous species (Fig.
S2a). Then, communities with increasing biomass have an increase in average
trophic level, in the number and proportion of piscivorous fishes (or
reciprocally for decreasing biomass), suggesting that temporal trends in
community biomass are driven by high trophic level fishes.

The total effect of the temporal trends in species richness on the temporal
trend in connectance was negative ($r_\delta =$ `r rttps$ct$r`, Fig.
\@ref(fig:std)a), resulting from a negative direct effect ($r_\delta =$
`r rdtps$ct$r`, Fig. \@ref(fig:sem)f) and a positive indirect effect via changes
in the temporal trends in community biomass ($r_\delta =$ `r ridtps$ct$r`).
Thus, when communities experienced an increase, or decrease, in species
richness, they also experienced a concomitant increase, or decrease, in
connectance. In contrast, species richness temporal trends had no direct effects
on trends in average trophic level ($r_\delta =$ `r rdtps$tlvl$r`, Fig.
\@ref(fig:sem)g), but a positive indirect effect via changes in the trends of
community biomass ($r_\delta =$ `r ridtps$tlvl$r`), which
adds up to a total positive effect of species richness temporal trends on
average trophic level ($r_\delta =$ `r rttps$tlvl$r`, Fig. \@ref(fig:std)a).
Temporal trends in species richness had a strong total positive effect on the
trends in redundancy (resp. $r_\delta =$ `r rdtps$red$r`, `r ridtps$red$r`), via
direct and indirect positive effects ($r_\delta =$ `r rttps$red$r`). All
together, it means that communities that have experienced species richness
declines over the last decades have decreased in average trophic level and
redundancy but increased in connectance. Conversely, those that experienced
species richness increases have increased in average trophic level, redundancy,
but decreased in connectance.

The effects of species richness were more complex, because species richness in
this model has indirect effects on food web structure through community biomass.
Species richness had a direct negative effect on connectance ($r_\delta =$
`r rdtps$ct$r`, Fig. \@ref(fig:sem)f), a positive indirect effect through
community biomass ($r_\delta =$ `r ridtps$ct$r`), so that species richness had a
negative effect in total on connectance ($r_\delta =$ `r rttps$ct$r`, Fig.
\@ref(fig:std)a). In contrast, species richness had no direct effects on in
average trophic level ($r_\delta =$ `r rdtps$tlvl$r`, Fig. \@ref(fig:sem)g), a
positive indirect effect through community biomass ($r_\delta =$
`r ridtps$tlvl$r`), which adds up to a positive total effect of species richness
temporal trends on average trophic level ($r_\delta =$ `r rttps$tlvl$r`, Fig.
\@ref(fig:std)a). Species richness temporal trends had direct and indirect
positive effect on the temporal trends in redundancy (resp. $r_\delta =$
`r rdtps$red$r`, `r ridtps$red$r`), which added up to a strong positive effect
($r_\delta =$ `r rttps$red$r`).

Spatial analyses, either using empirical data or the bioenergetic food web
model, largely aligned with temporal trends (Fig. \@ref(fig:std)a and S3),
showing consistent positive effects of species richness on community biomass and
of community biomass on average trophic level and redundancy. Species richness
also exhibited a direct negative effect on connectance and a positive effect on
redundancy, mirroring temporal patterns. However, spatial and temporal analyses
diverged in their effects of species richness on average trophic level, with
spatial data showing a negative effect and temporal trends indicating a positive
one. Detailed statistical effects and comparisons are provided in the Appendix.

## Robustness of food web structure to species extinctions

We implemented three secondary extinction simulation scenarios on the empirical
metaweb defined as a food web containing all the trophic species, i.e. the fish
species subdivided in size classes added to their resources, found across all
sites in the empirical data. This was to access a second line of inference about
whether top species were drivers of biomass and structure patterns (see section
4.2). When species are removed in a sequential order beginning by the highest
trophic levels, connectance, average trophic level, and redundancy decreased
(purple lines, Fig. \@ref(fig:std)b). On the other hand, removing low trophic level species
first increases connectance and average trophic level, but did not affect
redundancy (yellow lines, Fig. \@ref(fig:std)b). In turn, random species extinction sequence
had little effect on food web structure (black line, Fig. \@ref(fig:std)b). These results,
in conjunction with our empirical analyses above, suggest that top consumers
play a disproportionate role in driving the changes in food web structure
reported.

```{r}
std_cap <- paste0( "a, Direct and total effects of community biomass and species
  richness trends on food web structure trends. Standardized coefficients (r)
  derived from the temporal, spatial, and bioenergetic model Structural Equation
  Models. Points and lines display the estimated effects and their confidence
  intervals at 95\\%. Values are displayed in Table S3, R-squared in Table S4, and
  the visualization of the bivariate relationships for the spatial and
  bioenergetic model SEMs are displayed in Figure S3. Average trophic level is
  weighted by the biomass of the trophic species. Redundancy is the proportion
  of redundant links. b, Effect of species extinctions on the structure of the
  empirical metaweb according to three scenarios. In the Decreasing and
  Increasing scenarios, species were ordered by their trophic level and removed
  from the metaweb starting respectively from the highest and lowest trophic
  level. Species were removed in a random sequence (50 sequences, grey dots) for
  the random scenario, the black line displaying the average predicted by a GAM
  linear model. For the extinction sequence, the average trophic level was not
  weighted by the biomass of species as this information is not available at the
  metaweb scale. Similar results were obtained with trophic species (i.e. node)
  extinction (Fig. S4).")
```

```{r std, fig.cap=std_cap, out.width = "100%"}
tar_load(p_sem_tot_ext_sp_file)
knitr::include_graphics(p_sem_tot_ext_sp_file[[1]])
```

# Discussion

Our findings reveal several key patterns: (1) temporal trends in species
richness were positively associated with trends in community biomass; (2)
temporal trends in community biomass were positively linked to trends in
connectance and average trophic level; (3) temporal trends in species richness
were negatively associated with trends in connectance but only weakly related to
trends in average trophic level; and (4) both community biomass and species
richness trends were positively associated with trends in trophic pathway
redundancy. Additionally, we found that (5) top predator species play a
significant role in driving these patterns. Importantly, these temporal
relationships were largely consistent with patterns observed across spatial
gradients, both in empirical data and in simulations from a bioenergetic food
web model. In the following discussion, we explore these findings in the context
of food web assembly theory and Biodiversity-Ecosystem Functioning theory, and
consider their broader ecological implications.

We found that community biomass scaled positively with species richness, a
pattern consistent across our temporal, spatial, and theoretical analyses. This
result aligns with previous studies in fish communities [@woods_testing_2020;
@danet_species_2021] as well as in grasslands, salt marshes, and forests
[@hector_plant_1999; @tilman_biodiversity_2014; @wu_relationship_2015;
@li_relationship_2018; but see @grace_integrative_2016]. These empirical
findings are supported by theory, which suggests that food web diversity
enhances biomass by stimulating primary production and resource uptake
[@schneider_animal_2016; @wang_biodiversity_2018; @wang_intraguild_2019;
@nie_will_2023] and by Biodiversity-Ecosystem Functioning theory, which predicts
that species loss reduces community biomass [@loreau_biodiversity_2001]. Our
results reinforce the idea that declines in species richness can lead to reduced
biomass across ecosystems, highlighting the potential for conservation actions
that promote species richness to also enhance community biomass.

Our results also reveal that food webs experiencing biomass declines undergo
significant structural changes, including the loss of top predators and reduced
connectance. These findings align with energetic and trophic biogeography
theories [@lindeman_trophic-dynamic_1942; @gravel_trophic_2011], which posit
that the presence of high-trophic-level species is constrained by the
availability of energy and resources [@lindeman_trophic-dynamic_1942,
@takimoto_ecosystem_2008; @post_ecosystem_2000; @post_long_2002]. Thus, a
decline in community biomass may lead to insufficient resources to sustain top
predators, while an increase in biomass could provide the necessary energy to
support these species [@takimoto_ecosystem_2008; @post_ecosystem_2000].
Alternatively, Biodiversity-Ecosystem Functioning theory suggests that the
presence of higher trophic levels may reduce niche overlap among producers,
enhancing productivity and community biomass [@schneider_animal_2016;
@wang_biodiversity_2018; @wang_intraguild_2019; @nie_will_2023]. The
relationship between community biomass and connectance is less clear, with a
weaker positive association compared to average trophic level. Previous
theoretical studies have reported contrasting effects of connectance on
productivity [@nie_will_2023; @wang_intraguild_2019; @thebault_food-web_2003;
@poisot_trophic_2013; @schneider_animal_2016]. For instance, while increased
connectance between herbivores and plants can either enhance or reduce
productivity [@thebault_food web_2003; @poisot_trophic_2013; @nie_will_2023],
greater connectance among carnivores generally boosts productivity by reducing
top-down control and allowing biomass to accumulate across trophic levels
[@wang_intraguild_2019; @schneider_animal_2016].

Our study demonstrates that temporal trends in community biomass and species
richness are positively associated with the proportion of redundant trophic
pathways in food webs. In other words, declines in community biomass and species
richness lead to a loss of redundant links, while increases in these factors
result in greater redundancy. Since redundant links are those that can be
removed without triggering secondary extinctions [@allesina_who_2004;
@bodini_using_2009], our findings suggest that declines in community biomass and
species richness reduce food web robustness to further perturbations
[@allesina_functional_2009]. The positive effect of community biomass on
redundancy aligns with theoretical predictions: higher redundancy can reduce
trophic complementarity within species, weaken top-down control, and promote
compensatory dynamics, thereby enhancing robustness following extinctions
[@staniczenko_structural_2010]. Similarly, the positive effect of species
richness on redundancy supports the insurance hypothesis
[@yachi_biodiversity_1999; @loreau_biodiversity_2021], where greater species
richness and functional redundancy buffer ecosystems against perturbations,
increasing stability [@sanders_trophic_2018].

Furthermore, our results highlight the disproportionate role of top
trophic-level species in maintaining redundant links within food webs. This
pattern is consistent with previous empirical studies showing that top predators
are often the most generalist and highly connected species
[@montoya_ecological_2006; @montoya_press_2009]. Consequently, disturbances
affecting these top predators are likely to have the greatest impact on overall
community stability [@montoya_press_2009]. These findings underscore the
importance of preserving top predators and maintaining species richness to
safeguard food web robustness and ecosystem stability.

Species richness was negatively associated with connectance, consistent with
previous findings and likely driven by limitations in predator generality
[@dunne_network_2006]. In contrast, the effects of species richness on average
trophic level were more nuanced and considerably weaker than those of community
biomass. Direct effects of species richness on average trophic level were either
negligible or negative, as observed in temporal, spatial, and theoretical
models. However, when accounting for the indirect effects of species
richness—mediated through changes in community biomass—temporal and theoretical
analyses revealed a positive association between species richness and average
trophic level. These contrasting effects may be explained by our extinction
simulations, which showed that the removal of top trophic-level species had a
much smaller impact on average trophic level than on connectance. This suggests
that while species richness influences food web structure, its effects are more
pronounced on connectance than on average trophic level dynamics.

We found strong agreement between our temporal and spatial analyses, as well as
with predictions from food web theory and Biodiversity-Ecosystem Functioning
research. This suggests that spatial comparisons of generic topological food web
properties and biodiversity metrics may serve as reasonable surrogates for
predicting temporal trends—supporting the validity of the space-for-time
substitution in both empirical data and bioenergetic food web models. However,
this agreement did not hold for the relationship between weighted average
trophic level (hereafter “average trophic level”) and species richness. This
discrepancy may arise because average trophic level incorporates both food web
topology and biomass distribution, which are influenced by dynamic processes
such as bottom-up and top-down control. These processes likely require
additional information about species metabolism and productivity
[@brose_allometric_2006; @barbier_pyramids_2019]. This interpretation is
supported by the fact that our spatial analysis accurately predicted the
temporal association between community biomass and average trophic level.

Our study provides compelling evidence that biodiversity trends are profoundly
reshaping food web structure. By linking these patterns to
Biodiversity-Ecosystem Functioning and food web theory, we demonstrate that
widespread biodiversity declines likely have significant and detrimental
consequences for ecosystem functioning and stability in the face of future
perturbations. These findings suggest that assessments of food web structure,
community biomass, and species richness—which are often more feasible than
direct measurements of ecosystem functions—could serve as valuable indicators of
functional decline and metrics for evaluating the efficacy of mitigation
strategies. Importantly, our study also highlights critical avenues for future
research, stemming from its limitations. For instance, integrating food web
metrics into conservation planning will require understanding how they relate to
other facets of biodiversity, such as the presence of non-native species, and
drivers of biodiversity changes, including habitat degradation
[@danet_past_2024] and climate change [@bonnaffe_warming_2024]. Ultimately, our
work aims to bridge the gap between Biodiversity-Ecosystem Functioning and food
web theory, fostering greater integration with research on biodiversity changes.

# Material & Methods

We used a monitoring of freshwater fishes conducted all over France by the
Office Français de la Biodiversité (OFB). The sampling protocols were
standardized since 1995 and a complete description of the sampling process can
be found in @danet_species_2021. The fishes are sampled using electrofishing all across the
streams in shallow streams and over the bank by boat in deeper streams. All the
fishes are counted and the body size of fishes are recorded on a sample of
individuals, enabling to estimate the distribution of individual body size by
species for each sampling [@bonnaffe_comparison_2021;
@danet_species_2021].

We selected the sites that have been monitored for at least 10 years in the
1995-2018 period, with the same sampling protocol over time. We selected one
sampling per year by site. In each site, we ensured low temporal heterogeneity
within the sampling period by selecting samplings that were performed in the
same period, leading to a mean standard deviation of sampled month in each site
of three weeks [@danet_species_2021]. Those steps lead to the selection of
`r nb_st` sites (Fig. \@ref(fig:matmet)a, Table S1).

## Biomass

We first estimated the body mass of each fish individual with the allometric
law: $B_i = 0.01 \times l_i^{3.03}$, $B_i$ being the biomass expressed in gram and $l_i$ the
body length of the individual $i$ expressed in millimeters [@froese_cube_2006;
@danet_species_2021]. Then the biomass of each species and community
were obtained by summing the body mass of individuals. Finally, the biomass was
reported to the sampled surface (in $m^2$) to account for sampling effort.

## Food web reconstructions

We reconstructed food web structure in two steps [@bonnaffe_comparison_2021;
@danet_species_2021]: (1) we first computed the metaweb, describing the
potential trophic interactions among all size classes species, (2) then we
extracted yearly species size classes and their interactions according to their
presence in a given sampling event, to establish a annual snapshot of food webs.

We inferred the metaweb according to ontogenic diet data and body size
[@gravel_inferring_2013; @poisot_trophic_2013; @brose_predator_2019;
@bonnaffe_comparison_2021; @danet_species_2021]. The method is extensively
described in @bonnaffe_comparison_2021 and @danet_species_2021. The food diet of
stream fishes changes along their life cycle or stage, therefore according to
their body size. Each fish species was divided in nine size classes, i.e.
trophic species, equally distributed in their observed body size range. Thus,
a trophic species is a group of individuals of the same species and size class
that are expected to share the same preys and predators because trophic
interactions are largely determined by predator-prey size ratio
[@brose_allometric_2006; @brose_predator_2019]. We also added seven nodes present in the ontogenic
food diet database: detritus, biofilm, phytoplankton, zooplankton, macrophages,
phytobenthos and zoobenthos.  The resulting metaweb contained 45 fish species
divided into 9 size classes plus 7 resource nodes, summing up to 412 nodes.
Consideration of size classes, i.e.  trophic species instead of species,
improves inference of trophic interactions [@jennings_weak_2001].

We determined the feeding interactions of trophic species according to their
species identity and their body size. The diet database described two or three
life stages by fish species, each one being delimited by body size. The
ontogenic fish diet database was filled according to fishbase
[@froese_fishbase_2021] and literature [@bonnaffe_comparison_2021;
@danet_species_2021].

The feeding interactions among fishes were based on the body size ratio
between predators and preys. We defined the predation window of a
piscivorous trophic species as 3% to 45% of its midpoint
[@mittelbach_ontogeny_1998; @claessen_impact_2002; @hart_handbook_2008].  A
trophic link was defined between the piscivorous trophic species and all the
trophic species whose midpoint in body size range was included in the predation
window.  The fish-resource trophic links were set between the trophic species
and the resource nodes according to the food items of their size class.
Finally, the feeding interactions between resources were defined according to the
literature [@allan_stream_2007; @hart_handbook_2008].

## Community and food web structure

At each site, yearly food webs were described by their community biomass and
species richness. Yearly food web structure was described by their connectance,
describing the degree of generalism in a network [@dunne_network_2006;
@delmas_analysing_2019]. Connectance was computed as the ratio between the
number of actual links and the number of potential links ($C =
\frac{L}{S(S-1)}$, with $L$ being the number of links and $S$ the number of
trophic species). We also computed the average trophic level, weighted by the
biomass of the trophic species. The trophic level of a trophic species was
computed recursively as 1 added to the mean trophic level of the preys it
consumes [@herendeen_ecological_2008]. We computed the number of redundant and
functional links using the dominator tree approach [@allesina_functional_2009].
A species $i$ is a dominator of species $j$ if $i$ mediates the connection $j$
to basal resources such as the extinction of $i$ remove the connection of $j$ to
basal resources. It is then to identify the set of trophic links that are
essential to connect species to basal resources, i.e. functional links, and the
redundant links. This method requires a single basal root in the food webs. To
do so, we added a basal dummy resource node to the empirical food webs connected
to biofilm and phytoplankton resource nodes. For food webs generated from
theoretical simulations (see below), we added a resource node connected to all
primary producers. We computed the set of dominator trees using the
`dominator_tree()` function from the `igraph` R package (Appendix B). We
computed the number of functional links as the number of links in the resulting
dominator tree of the food web. We checked that the procedure reproduced the
examples presented in previous studies [@allesina_functional_2009, Appendix B].
We then derived the proportion of redundant links in food webs of empirical and
theoretical simulations.

## Temporal trends of food web

We modeled the temporal trends of food web structure at each site ($Y_i$) using
linear hierarchical model. Time was considered as the number of years since the
first sampling ($t = 0$) at each site. We accounted for the spatial structure of
the data by adding random intercepts ($a$) and slopes ($b$) for the
basin ($a_n$, $b_n$) and site ($a_{i|n}$, $b_{i|n}$) identity, site effects
being nested in the basin one.

$$
\begin{aligned}
    Y_{i|n, t} &= \alpha + \beta Time_t + \epsilon_{i|n, t}
  (\#eq:gen)
\end{aligned}
$$

- $\alpha = \alpha_0 + a_n + a_{i|n}$
- $\beta = \mu + b_n + b_{i|n}$
- $a_n, a_{i|n}, b_n, b_{i|n}, \epsilon_{i|n, t} \sim \mathcal{N}(0,\,\sigma^{2})$
- $n$: basin, $i$: site $i$, $t$: time $t$

The model parameters were estimated using INLA (Integrated Nested Laplacian
Approximation) Bayesian inference with default uninformative priors
[@rue_bayesian_2017]. The prior distributions for the fixed intercept and slope were
modeled as a flat normal distribution with a mean of zero ($\mathcal{N}(0,\, 1000)$).
Prior distributions for the Gaussian error $\epsilon_{i|n, t}$ and the random
effects were modeled as a log gamma distribution with
shape and inverse scale parameters ($\mathcal{G}(s, \, \tau) = \mathcal{G}(1, \,
5.10^-5)$), then their respective standard deviation were obtained by
back-transforming the inverse scale parameter ($\sigma = 1 / \sqrt{\tau}$).
Community biomass and species richness were log-transformed.

The performance of the model was assessed visually by plotting the fitted versus
the observed values (Fig. S5). We also visually inspected the Posterior
Integral Transform (PIT) and Conditional Predictive Ordinate (CPO) to assess the
quality of the model fit. We also computed the marginal ($R^2_m$) and conditional
($R^2_c$) R-squared of the models following previous formulations
[@nakagawa_general_2013; @lahuis_explained_2014]. In order to account for the
variability associated with the priors [@gelman_r-squared_2019], we computed the
variance of the posterior distribution  of each predicted values ($Var_{fit} =
\sigma^2(\hat{Y_i})$). We computed $R^2$ values for each predicted values, then
obtaining a $R^2$ distribution [@gelman_r-squared_2019]. We reported mean
marginal and conditional $R^2$ associated the 95% credible interval computed
using the Highest Posterior Density method.

$$
R^2_m = \dfrac{Var_{fit}}{Var_{fit} + Var_{res}} = \dfrac{\sigma^2(\hat{Y_i})}{\sigma^2(\hat{Y_i}) + \sigma^2(Y_i - \hat{Y_i})}
$$

$$
R^2_c = \dfrac{Var_{fit} + (a_n)^2 + (a_{i|n})^2}{Var_{fit}  + (a_n)^2 + (a_{i|n})^2 + Var_{res}}
$$

$y_i$ and $\hat{y_i}$ being respectively the observations and the
predicted values, $Var_{fit}$ and $Var_{res}$ being respectively the variance of
predictive means and the variance of the residuals. $a_n$ and $a_{i|n}$ are
respectively the standard deviation on the random intercept associated to the
basin and the site.

We extracted the temporal trends at each site from the hierarchical model
using mean BLUP values (Best Linear Unbiaised Prediction), then multiplied
by 10 to get temporal trends by decade. For community biomass and species
richness which were log-transformed, we derived temporal trends in percentage
change by decade by back-transforming the slopes ($\beta\prime = (e^{\beta } - 1)
\times 100 \times 10$).

Overall, we found that standard deviation attributed to site effects in both
baseline and temporal trends in community structure were much higher than the
ones found for hydrographic basin (Table S5).

## Spatial

We assessed the validity of the space-for-time substitution hypothesis by
comparing the relationships estimated among food web structure in time to the
ones inferred in space. We used the same sites as in the temporal analysis,
but we summarized the temporal series of each station by the median value over time
of each facet of food web structure described above (Fig. \@ref(fig:matmet)).

## Statistical analysis

We modeled the relationships among food web structure indices, i.e. connectance,
average trophic level, the proportion of redundant trophic links, and community
biomass and species richness, with Structural Equation Models. We
modeled species richness as related to community biomass and food web structure.
In turn, community biomass was related to food web structure. We computed one
structural equation model to assess the relationships among the temporal trends
and one for the relationships over the spatial gradient, we added a random
effect of the hydrographic basin on the intercept for the latter one.

In order to control for the uncertainties of the estimated temporal trends, we
weighted each temporal trend by the inverse of the sum of the standard deviation
of the estimated trends, i.e. the estimated slopes for the response and the
explicative variables ($w_j = 1 / \sum{e_j}$, $w_j$ being the weight, and
$s_{e_j}$ the standard deviation on the temporal trends of the variables).

The coefficients of the Structural Equation Models were computed with
PiecewiseSEM R package [@lefcheck_piecewisesem_2016]. The coefficients were
standardized and the confidence intervals were computed with semEff R package
[@murphy_semeff_2021]. Confidence intervals were obtained by bootstrapping the
linear models 1000 times and computing adjusted bootstrap percentile intervals
(BCa), which are adjusted for the asymmetric distribution of resampling. The marginal
and conditional $R^2$ associated to the linear models of the SEMs were computed
with the Nagelkerke method [@lefcheck_piecewisesem_2016, Table S4].

All the statistical models displayed low multicollinearity (Variance inflation
factors < 3, Table S6). We visually inspected the residuals of the linear
models.

## Extinction simulation in the metaweb

We simulated extinction sequences in the metaweb according to three scenarios to
get expectations on how the food web structure should change according to the
addition and deletion of species and nodes. The scenario of extinction sequences
were related to the trophic level of the species/nodes. In the decreasing and
increasing scenario, we removed species/nodes in decreasing, resp. increasing
order, of their trophic level, i.e. respectively beginning by the highest and
lowest trophic level. In the random extinction sequence, we built 50 random
extinction sequences. For each simulation, we removed fish species/nodes that
lost all their preys. We then computed the connectance, the average and the
maximum trophic level after each extinction.

## Theoretical simulations

```{r}
tar_load(sim_param)
nsim <- nrow(sim_param)
nfw <- length(unique(sim_param$fw_id))
s_range <- c(min = min(sim_param$S), max = max(sim_param$S))
c_range <- c(min = min(sim_param$C), max = max(sim_param$C))
k_range <- c(min = min(sim_param$K), max = max(sim_param$K))
```

We used a bioenergetic food web model to simulate the effects of community
biomass and species richness changes on food web structure. This food web model
lifts one of the limitations of our data, that is it includes the biomass of all
the food webs from primary producers to top predators. This model allows
simulating complex communities because its parameterization depends mainly on
species bodymasses. We first generated `r nfw` food webs varying in species
richness (range: `r s_range["min"]`-`r s_range["max"]`)
and connectance (range: `r c_range["min"]`-`r c_range["max"]`) using the niche
model algorithm, which reproduces food web topology typically found in empirical
food webs [@williams_simple_2000]. We set species bodymasses ($M_i$) using a predator-prey
mass ratio of 100 ($Z$) [@brose_allometric_2006; @brose_predator_2019], such as $M_i =
Z^{(TL_i - 1)}$, $TL_i$ being the trophic level of the species $i$. The
metabolic rates of the species ($x_i$) were then derived from allometric
relationships between metabolism and bodymass empirically estimated
[@brose_allometric_2006], such as $x_i = -0.314M_i^{-0.25}$.

The dynamics of species biomasses are described by an ODE systems of two
equations, for primary producers (eq. \@ref(eq:prod)) and consumers (eq.
\@ref(eq:cons)) respectively:

\begin{equation}
    (\#eq:prod)
	\frac{dB_i}{dt} =
    r_i B_i G_i
    - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
\end{equation}

\begin{equation}
    (\#eq:growth)
    G_i = (1 - \dfrac{\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j}{K_i})
\end{equation}

The primary producer biomass increase with a growth rate ($r_i$) of 1 and a
logistic growth (eq. \@ref(eq:growth)), growth being limited by intraspecific
($\alpha_{ii}$, set to 1), interspecific ($\alpha_{ij}$, set to 0.5) competition,
and carrying capacity ($K_i$, varying in the simulations). Primary producers
loses biomass by being eaten by consumers at a rate that depend on consumer
metabolic rates ($x_j$), attack rates ($y_j$, set to 8 @delmas_analysing_2019),
a functional response ($F_{ji}$), and assimilation efficiency ($e_{ij}$, set to
0.45 for herbivores and 0.85 for carnivores, @delmas_analysing_2019).

\begin{equation}
    (\#eq:cons)
        \frac{dB_i}{dt} =
        \sum_{j \in \{\mathrm{res}\}_i} x_i y_i B_k F_{ij}
        - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
        - x_i B_i
\end{equation}

Consumers gain biomass by eating other species and lose biomass by being
consumed and by metabolic loses ($x_i$, eq. \@ref(eq:cons)). The functional
response (eq. \@ref(eq:holling)) of consumer feeding depends on the preference
for its preys ($\omega_{ij}$, set even across consumer preys) and
half-saturation rate ($B_0$, set to 0.5). The hill coefficient ($h$, set to 1.5)
determines the shape of the functional response (Holling's type II or III).

\begin{equation}
    (\#eq:holling)
    F_{ij} = \frac{\omega_{ij} B_j^h}
    {B_0^h + \sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h}
\end{equation}

We simulated the dynamics of the `r nfw` food webs across a gradient of global carrying capacity (K,
range: `r k_range["min"]`-`r k_range["max"]`). The carrying capacity were
standardized according to the number of producers and interspecific competition
($K_i = \frac{1 + \alpha_{ij} (S - 1)}{S}K$) to prevent changes in productivity
due to system size [@ives_stability_2000; @loreau_species_2008]. We ran
simulations for 5000 timesteps. At the end of simulations, we removed dead
species and measured community biomass, species richness and food web structure.
If the final community contained disconnected producers (i.e. without consumers)
or disconnected consumers (i.e. consumers without preys), we discarded them and
run the simulations again with the same parametrization. The simulations were
done using the Julia package `EcologicalNetworksDynamics.jl`
[@lajaaiti_ecologicalnetworksdynamicsjl_2024]. Then, we computed the community
and food web metrics at the last timesteps and analyzed the data with the same
SEM described above.

# References  {-}

<div id="refs"></div>

\newpage