### 2025-02-07 paper/appendix.Rmd
Minor edits: sent to Loubna & Willem for review

---
title: "Biodiversity temporal trends are reshaping food web structure and redundancy in riverine ecosystems"
subtitle: "Alain Danet (a,b), Loubna El Madouri (a,c), Andrew P. Beckerman (b),
Maud Mouchet (a), Colin Fontaine (a) and Elisa Thébault (d)"
date: Supplementary Material
output:
  bookdown::pdf_document2:
    fig_caption: true
    keep_tex: true
    toc: false
fontsize: 10pt
header-includes:
   - \newcommand{\beginsupplement}{\renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
   - \renewcommand{\arraystretch}{1}
geometry: margin=2cm
bibliography: references.bib
number_sections: false
csl: pnas.csl
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.align = "center",
  fig.fullwidth = TRUE,
  fig.show = "hold",
  fig.pos = "H",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE
)
options(
  #kableExtra.latex.load_packages = FALSE,
  knitr.table.format = "latex",
  knitr.graphics.rel_path = FALSE
)
```

```{r, echo=FALSE, message=FALSE, results="hide"}
library(targets)
library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
library(sf)
library(nlme)
library(igraph)

mypath <- rprojroot::find_package_root_file
source(mypath("R/variable_shortcut.R")) # Define your custom code as a bunch of
#functions.

source(mypath("R/misc.R"))
source_dir(mypath("R")) # Define your custom code as a bunch of functions.

mytheme <- theme_cowplot() +
  background_grid()

theme_set(mytheme)
```

```{r}
tar_load(c(semeff_tot, semeff_tot_ok))
semeff_tot <- semeff_tot %>%
  mutate(across(c(response, predictor), ~str_remove(.x, "mean_")))

rdtps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std")
    )
)
rdsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", model = "spatial"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", model = "spatial")
    )
)
rdm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Direct", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Direct", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
```

```{r}
ridtps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "indirect")
    )
)
ridsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "indirect")
    )
)
ridm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Indirect", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Indirect", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
rttps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "total")
    )
)
rtsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std",
      model = "spatial", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "total")
    )
)
rtm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Total", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Total", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
```

\beginsupplement

# Comparison of analysis in space and time

We further assessed the consistency of the reported associations among temporal
trends by implementing two additional analyses. We first tested if the
associations among the temporal trends within sites was holding across sites,
i.e. using spatial variations in food web structure rather than temporal changes
over time (space-for-time hypothesis). Second, we performed theoretical
simulations with a bioenergetic food web model to cross-reference our site and
species specific empirical data on riverine fish with a more generalized
representation of food web dynamics known to represent accurate relationships
among our focal variables [@brose_allometric_2006; @eschenbrenner_diversity_2023;
@danet_response_2024]. With those additional empirical and theoretical data, we
assessed additional SEM models with the same structure as in Section 4.2.

As in the temporal analysis, species richness had a positive effect on community
biomass both in empirical data and in theoretical simulations (resp. $r_\delta
=$ `r rdsp$bm$r` and `r rdm$bm$r`, Fig. 3a and S3). In both
empirical data and theoretical simulations, community biomass had a positive
effect on average trophic level (resp. $r_\delta =$ `r rdsp$tlvl$bm` and
`r rdm$tlvl$bm`), and redundancy (resp. $r_\delta =$ `r rdsp$red$bm` and
`r rdm$red$bm`), as in temporal analysis. We found weak evidence of a positive
effect of spatial variation in community biomass on connectance in empirical
data ($r_\delta =$ `r rdsp$ct$bm`) but strong evidence in theoretical
simulations ($r_\delta =$ `r rdm$ct$bm`, Fig. 3a).

The effects of spatial variations in species richness on food web structure
mostly aligned with those from the temporal trend analysis, except for average
trophic level. We found a direct negative effect of species richness on
connectance both using data and model simulations (resp. $r_\delta =$
`r rdsp$ct$r` and `r rdm$ct$r`, Fig. 3a), on average trophic level
(resp. $r_\delta =$ `r rdsp$tlvl$r` and `r rdm$tlvl$r`), but a direct positive
effect on redundancy (resp. $r_\delta =$ `r rdsp$red$r` and `r rdm$red$r`).
Species richness had an indirect positive effect of species richness through
community biomass, which added up to strong total negative effects on
connectance in empirical data and simulations (resp. $r_\delta =$
`r rtsp$ct$r` and `r rtm$ct$r`) and positive on redundancy (resp. $r_\delta =$
`r rtsp$red$r` and `r rtm$red$r`). We found that
species richness had a total positive effect on average trophic level in
bioenergetic model, in agreement with temporal trends (resp. $r_\delta =$ `r rtm$tlvl$r`),
but a total negative effect in spatial variations (resp. $r_\delta =$ `r rtsp$tlvl$r`).

# Disantengling functional and redundant links

Because species biomass are involved in trophic interactions, species losses can
have cascading effects and trigger secondary extinctions [@dunne_network_2002].
In the context of biodiversity crisis, assessing the robustness of ecological
networks to species losses has been then a big question in Ecology. Previous
research showed that food-web structure, such as connectance, increased
robustness [@dunne_network_2002]. The idea is that more connected food-web will
be more stable because they contain a large number of pathways convey biomass
through the food-web. However, connectance as some limitations as a robustness
metric, because each trophic link does not contribute equally to the robustness
of an ecological network [@allesina_who_2004]. Trophic links can be categorised as
functional or redundant, depending on their role in transmitting biomass through
the food-web. Functional links are the set of links that is required to connect
every species to the basal resource of the food-web, all of the other being
redundant. Only functional links affect directly robustness, as their removal
will trigger secondary extinctions. The removal of redundant links in turn do
not directly trigger secondary extinctions, but makes the food-webs more fragile
to further extinctions [@allesina_functional_2009].

The identification of functional and redundant links is based on the use of
denominator trees [@bodini_using_2009; @allesina_who_2004;
@allesina_functional_2009]. Dominator trees allows to identify the set of
trophic links that are essential to connect all the consumers to the basal resource, i.e.
the functional links, and the redundant links, i.e. the redundant trophic pathways
from basal resource to consumers. Pioneer papers explained the algorithm to
compute the set of dominator trees and we refer to them for further
information and examples [@bodini_using_2009; @allesina_who_2004;
@allesina_functional_2009]. Here, we aim to show that while implementing this
analysis could be difficult at the time, this algorithm is now widely
available. In R, it is available in `igraph::dominator_tree()`. Below we show
how the number of functional and redundant links can be  easily computed from R
by replicating three examples presented in previous studies.

```{r, echo = TRUE}
compute_redundancy
```

Here is an example with the food-web displayed in appendix of
@bodini_using_2009:

```{r, echo = TRUE}
# Get the food-web example from Bodini et al. (2009), Appendix 1 
g <- get_bodini_graph_appendix()
# compute reduncancy
out <- compute_redundancy(g, root = "R")
as.data.frame(out)
```

```{r}
vertex_color <- "white"
vertex_size  <- 40
```

We used this function to reproduce the three examples (Fig. \@ref(fig:examples)),
which were displayed in the Fig. 1 and Appendix 1 of @bodini_using_2009, and
in the Fig. 1 of @allesina_functional_2009. The table \@ref(tab:redun) displays
the number and the proportion of functional and redundant links from those
examples.

```{r bod1}
bodinifig1 <- get_bodini_graph_fig1()

# Complete graph
p_bodinifig1_all <- ~{
  lg <- layout_as_tree(bodinifig1, root = "root")
  lg[, 2] <- -lg[, 2]
  plot(bodinifig1, layout = lg, vertex.color = vertex_color, vertex.size = vertex_size)
}

# Only functional links
bodinifig1_dtree <- dominator_tree(bodinifig1, root = "root")
p_bodinifig1_functional <- ~{
  layout <- layout_as_tree(bodinifig1_dtree$domtree, root = "root")
  layout[, 2] <- -layout[, 2]
  plot(bodinifig1_dtree$domtree, layout = layout, vertex.color = vertex_color, vertex.size = vertex_size)
}
```

```{r bod2, echo = FALSE, results = "hide"}
g <- get_bodini_graph_appendix()

# Complete graph
p_bodiniappen_all <- ~{
  lg <- layout_as_tree(g, root = "R")
  lg[, 2] <- -lg[, 2]
  plot(g, layout = lg, vertex.color = vertex_color, vertex.size = vertex_size)
}

# Only functional links
dtree <- dominator_tree(g, root = "R")
p_bodiniappen_functional <- ~{
  lg <- layout_as_tree(dtree$domtree, root = "R")
  lg[, 2] <- -lg[, 2]
  plot(
    dtree$domtree,
    layout = lg,
    vertex.color = vertex_color,
    vertex.size = vertex_size
  )
}

p_bodini <- plot_grid(
  ggdraw(p_bodinifig1_all), ggdraw(p_bodinifig1_functional),
  ggdraw(p_bodiniappen_all), ggdraw(p_bodiniappen_functional),
  labels = "auto"
)
ggsave_multiple(
  paste0("p_bodini", c(".png", ".pdf")),
  plot = p_bodini,
  path = here::here("figures"),
  scale = 2.6,
  units = c("mm"),
  width = 80,
  height = 80 * 1.5)
```

```{r all1, echo = FALSE, results = "hide"}
allesinafig1 <- get_allesina_graph(diag = FALSE)

p_allesinafig1_all <- ~{
  lg <- layout_as_tree(allesinafig1, root = "0")
  lg[, 2] <- -lg[, 2]
  plot(allesinafig1, layout = lg, vertex.color = vertex_color, vertex.size = vertex_size)
}

# Only functional links
allesinafig1_dtree <- dominator_tree(allesinafig1, root = "0")
p_allesinafig1_functional <- ~{
  layout <- layout_as_tree(allesinafig1_dtree$domtree, root = "0")
  layout[, 2] <- -layout[, 2]
  plot(
    allesinafig1_dtree$domtree,
    layout = layout,
    vertex.color = vertex_color,
    vertex.size = vertex_size
  )
}

p_examples <- plot_grid(
  ggdraw(p_bodinifig1_all), ggdraw(p_bodinifig1_functional),
  ggdraw(p_bodiniappen_all), ggdraw(p_bodiniappen_functional),
  ggdraw(p_allesinafig1_all), ggdraw(p_allesinafig1_functional),
  nrow = 3,
  labels = "auto"
)
ggsave_multiple(
  paste0("p_examples", c(".png", ".pdf")),
  plot = p_examples,
  path = here::here("figures"),
  scale = 4.0,
  units = c("mm"),
  width = 80,
  height = 80 * 1.2)
```

(ref:example-cap) Reproduced examples of categorising functional links in
food webs using `igraph::dominator_tree()` R function. Left columns (a, c, e)
displays the complete food-web while right column displays food-web with the
functional links only (b, d, f). (a,b): reproduction of Fig. 1 of
@bodini_using_2009, (c, d): reproduction of appendix figures of @bodini_using_2009,
(e, f): reproduction of fig. 1 of @allesina_functional_2009. Note that (e, f) is
a bit different from the original, as noted by the authors, they could have
removed further redundant links (i.e. $2 \to 5$, or $3 \to 5$).

```{r examples, fig.cap = "(ref:example-cap)", out.height = "80%"}
knitr::include_graphics(here::here("figures", "p_examples.png"))
```

```{r}
examples <- list(
  bodiniappendix = g,
  bodinifig1 = bodinifig1,
  allesinafig1 = allesinafig1
)
redundancy_examples <- map2_dfr(
  examples,
  list(bodiniappendix = "R", bodinifig1 = "root", allesinafig1 = "0"),
  ~compute_redundancy(.x, root = .y), .id = "reference")
```

```{r}
paper_replacement <- c(
  bodinifig1 = "Bodini et al. (2009) Fig. 1",
  bodiniappendix = "Bodini et al. (2009) Supp material 1",
  allesinafig1 = "Allesina et al. (2009) Fig. 1"
)

redundancy_var_replacement <- function() {
  c(
    reference = "References",
    functional_links = "Nb functional links",
    redundant_links = "Nb redundant links",
    total_links = "Nb links",
    prop_functional_links = "Prop. functional links",
    prop_redundant_links = "Prop. redundant links"
  )
}

redundancy_examples %>%
  select(reference, total_links,
    functional_links, redundant_links,
    prop_functional_links, prop_redundant_links
  ) %>%
  mutate(reference = paper_replacement[reference]) %>%
  rename_with(~redundancy_var_replacement()[.x]) %>%
  kable(
    digits = 2,
    row.names = FALSE,
    label = "redun",
    booktabs = T,
    caption = "Computation of the number (Nb) of functional, redundant links, and
    their proportion (Prop.) from previous published food-webs.",
    ) %>%
  collapse_rows(columns = c(1), valign = "top") %>%
  kable_styling(latex_options = "scale_down")
```

# Supplementary figures {-}

```{r}
tps10_cap <- paste0("Distribution of temporal trends per decade of the food-web
  structure by site estimated from the Bayesian models using Best Linear
  Unbiaised Prediction (BLUP). Species richness and biomass temporal trends are
  reported in percentage change per decade. Dashed red line and solid black line
  respectively indicate 0 and median change.")

```

```{r tps10, fig.cap=tps10_cap}
tar_load(figStps10)
knitr::include_graphics(figStps10)
```

```{r}
cor_cap <- paste0("Correlation among food-web metrics for a) temporal trends by
  site, b) median values by site (cf Main Methods), c) bioenergetic model.", " Pisc: piscivorous, rich.:
species richness, node: node of the food-web corresponding to a trophic species,
prop.: proportion."
)
```

```{r cor, fig.cap=cor_cap}
tar_load(figcorr)
knitr::include_graphics(figcorr)
```

```{r}
sem_sp_mod_cap <- "Direct effects of biomass and species richness on food-web
structure, derived from the structural equation models for the (a) spatial and
(b) bioenergetic model. Points display the data, black lines display the
predicted relationships."

```

```{r, fig.cap=sem_sp_mod_cap, out.width = "100%"}
tar_load(p_sem_sp_sim_file)
knitr::include_graphics(p_sem_sp_sim_file)
```

```{r}
rand_sim_cap <- paste0(
  "Simulations of (b) nodes extinctions in the metaweb according to three
  scenarios: random extinction sequence (grey dots), extinction sequence by
  decreasing order of trophic level, extinction sequence by increasing order of
  trophic level. The black line represents the average prediction of a GAM
  linear model for the 50 random extinction sequences."
)
```

```{r ext, fig.cap=rand_sim_cap}
tar_load(figSext)
knitr::include_graphics(figSext[[1]])
```

```{r}
obs_fit_cap <- "Fitted versus observed values from the linear models estimating
the temporal trends of the food-web structure. The black line depicts the first
bissector. Pearson's correlation coefficients are displayed."
```

```{r, fig.cap=obs_fit_cap}
tar_load(figSobsfit)
knitr::include_graphics(figSobsfit)
```

\pagebreak
\newpage

# Supplementary tables {-}

\pagebreak

```{r}
tar_load(c(op_table, com_metric_tab))
tab_quanti <- purrr::map_dfr(
  list(com_metric_tab, op_table),
  function(x) {colnames(x)[1] <- "Variable"; x}
)
tab_quanti %>%
  kbl(
    caption = "Summary distribution of community metrics and fish monitoring variables.",
    booktabs = T
    ) %>%
  kable_styling() %>%
  pack_rows("Community metrics", 1, 8) %>%
  pack_rows("Summary of fish monitoring", 9, 13)
```

```{r}
tar_load(r2_inla_tab)
kable(r2_inla_tab,
  row.names = FALSE,
  booktabs = T,
  caption = "Marginal and Conditional R-squared for the INLA models estimating
  the temporal trends of the food-web structure. Mean [95\\% Credible Interval].",
  label = "inlarsq") %>%
  kable_styling()
```

```{r}
tar_load(semeff_tot_tab)
ti <- semeff_tot_tab %>%
  pivot_wider(names_from = "Effect type", values_from = "Effect") %>%
  mutate(Response = factor(Response, levels = unique(var_replacement()))) %>%
  arrange(Model, Response)
kable(ti %>%
    select(-Model),
  row.names = FALSE,
  booktabs = T,
  caption = "Direct and total Standardized effects derived from the structural
  equation models. [95\\% Confidence Interval]. The confidence intervals were
  obtained with bootstrap (see Methods).") %>%
  collapse_rows(columns = c(1), valign = "top") %>%
  pack_rows(index = table(ti$Model))
```

```{r}
tar_load(r2_sem_tab)
kable(r2_sem_tab %>%
  mutate(
    Type = factor(Type, levels = effect_type_var()),
    Response = factor(Response, levels = unique(var_replacement()))
    ) %>%
  arrange(Type, Response),
  row.names = FALSE,
  booktabs = T,
  caption = "Marginal and conditional R squared for the Structural Equation
  Models.",
  label = "r2sem",
  ) %>%
  kableExtra::collapse_rows(columns = 1, valign = "top")
```

```{r}
tar_load(tab_inla_rand)
tab_inla_rand %>%
  mutate(Response = factor(Response, levels = unique(var_replacement()))) %>%
  arrange(Response) %>%
  kable(
    row.names = FALSE,
    booktabs = T,
    caption = "Standard deviation associated to the random effects and the residual error of
    the models. Mean [95\\% Credible Interval].",
    label = "inlarand") %>%
  collapse_rows(columns = c(1), valign = "top") %>%
  kable_styling()
```

```{r}
tar_load(tabvif)
kable(tabvif,
  row.names = FALSE,
  booktabs = T,
  caption = "Variance Inflation Factor for the more complex linear models
  included in the Structural Equation Models.",
  label = "vif",
  )
```