### 2025-04-10 paper/manuscript.Rmd
Update manuscript with Colin and Elisa comments

---
title: "Biodiversity temporal trends are reshaping food web structure and redundancy in riverine ecosystems"
subtitle: "Alain Danet (a,b), Loubna El Madouri (a,c), Willem Bonnaffé (d), Andrew P. Beckerman (b),
Maud Mouchet (a), Colin Fontaine (a) and Elisa Thébault (e)"
date: "Research report"
output:
  bookdown::pdf_document2:
    fig_caption: true
    keep_tex: true
    toc: false
  officedown::rdocx_document:
    number_sections: false
    base_format: "bookdown::word_document2"
always_allow_html: true
number_sections: false
bibliography: references.bib
link-citations: true
#https://pkgs.rstudio.com/rticles/articles/examples.html
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
csl: pnas.csl
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path = "figures/"
)
options(
  knitr.table.format = "latex",
  knitr.graphics.rel_path = FALSE
)
```

```{r, results = FALSE}
library(targets)
library(tidyverse)
library(magrittr)
library(cowplot)
library(viridis)
library(kableExtra)
library(sf)
mypath <- rprojroot::find_package_root_file
source(mypath("R/variable_shortcut.R"))
source(mypath("R/misc.R"))
source_dir(get_mypath("R")) # Define your custom code as a bunch of functions.

mytheme <- theme_cowplot() +
background_grid()

theme_set(mytheme)
```

```{r}
tar_load(full_data2)
nb_st <- unique(full_data2$station) %>% length
nb_op <- nrow(full_data2) %>% round(., -3)

rm(full_data2)
tar_load(glob_tps_trends_decade_no_drivers)
```

- Corresponding author: Alain Danet (a.h.danet@sheffield.ac.uk)

- Author affiliations:

(a) Centre d’Ecologie et des Sciences de la Conservation, UMR 7204
MNHN-CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de Paris, 43
rue Buffon, 75005 Paris, France

(b) School of Biosciences, Alfred Denny Building, University of Sheffield,
Western Bank, Sheffield S10 2TN, United Kingdom

(c) Laboratoire d’Écologie Fonctionnelle et Environnement, UMR 5245
CNRS-UT3-INPT, Avenue de l’Agrobiopôle, Toulouse, France

(d) Department of Biology, University of Oxford, Oxford, United Kingdom

(e) Sorbonne Université, CNRS, IRD, INRAE, Université Paris Est Créteil, 15
Institute of Ecology and Environmental Sciences of Paris (iEES-Paris), Paris,
France

- Classification: Biological Sciences (major) - Ecology (minor)

- Keywords: food web, biodiversity changes, redundancy, ecosystem functioning,
  vulnerability

# Abstract {-}

Biodiversity temporal changes are expected to profoundly impact ecosystem
functioning and stability, yet large-scale empirical evidence remains limited.
Bridging this gap requires aligning biodiversity temporal trends with food web
theory and Biodiversity-Ecosystem Functioning research, which provide strong
predictions about the ecological consequences of biodiversity loss. Most
insights rely on spatial comparisons of ecological communities, assuming they
reflect temporal trends—a space-for-time substitution. Here, we analyze
biodiversity and food web time series from over 400 riverine fish communities
across France (1995–2018) to evaluate how biodiversity temporal trends reshape
food web structure and community functioning. We examine relationships among
species richness, community biomass, and three key food web metrics:
connectance, weighted average trophic level, and trophic pathway redundancy. We
also test the space-for-time hypothesis using spatial gradients and a
theoretical food web model. Structural equation models and robustness analyses
reveal that declines in species richness strongly correlate with reduced
community biomass, with top trophic-level species playing a disproportionate
role in food web changes. Declining biomass is associated with the loss of top
trophic levels and decreased connectance, while temporal changes of species
richness negatively correlates with those of connectance. Both biomass and
species richness declines reduce trophic pathway redundancy, suggesting
biodiversity loss undermines community robustness to future perturbations. Our
findings align with food web theory and Biodiversity-Ecosystem Functioning
research, support the validity of space-for-time approaches for basic food web
metrics, and highlight the relevance of food web metrics as indicators of
ecosystem function and fragility.

# Significance {-}

Understanding how biodiversity loss affects ecological communities is crucial
for predicting and mitigating human impacts on ecosystems. This study provides
comprehensive evidence on how biodiversity changes are reshaping food web
structure using an extensive dataset of riverine fish communities. We
demonstrate that temporal changes in species richness and community biomass
consistently correlate with altered trophic levels, connectance, and the
redundancy of feeding relationships, suggesting energetic constraints and
biodiversity-ecosystem functioning mechanisms jointly shape food web assembly
and disassembly. Importantly, we show that declining biodiversity reduces the
redundancy of trophic pathways, potentially making communities more vulnerable
to future perturbations. These results suggest that food web structure metrics
could serve as valuable indicators for monitoring ecosystem health and
evaluating conservation strategies.

# Introduction

Human pressures are driving profound biodiversity changes, including freshwater
and terrestrial biodiversity declines linked to land degradation
[@newbold_global_2015; @cardinale_is_2018; @danet_past_2024] and reductions in
marine fish biomass due to intensified fishing pressures on predatory species
[@christensen_century_2014]. A central goal of Biodiversity-Ecosystem
Functioning and food web research has been to understand the ecological
consequences of these widespread biodiversity losses for ecosystem functioning
and stability [@loreau_biodiversity_2001; @loreau_ecological_2022;
@thebault_food-web_2003; @schneider_animal_2016; @dunne_network_2002;
@sole_complexity_2001; @allesina_functional_2009; @montoya_press_2009;
@sanders_trophic_2018]. However, research on biodiversity changes remains
largely disconnected from studies examining their ecological impacts, primarily
due to a lack of simultaneous temporal data on biodiversity and food web
dynamics [@danet_species_2021]. As a result, much of our understanding of the
ecological consequences of biodiversity changes over time relies on experimental
and theoretical approaches [@loreau_biodiversity_2001; @loreau_ecological_2022;
@thebault_food-web_2003; @dunne_network_2002; @sole_complexity_2001;
@allesina_functional_2009]. These approaches often rely on the critical
assumption that spatial comparisons of ecological communities can serve as a
proxy for temporal trends—a space-for-time substitution—when direct temporal
data are scarce [@damgaard_a_2019]. To advance our understanding of the
ecological consequences of biodiversity changes, it is essential to investigate
how these changes are reshaping food web structure in natural settings.

Here, we bridge the gap between biodiversity changes and food web dynamics by
leveraging a unique dataset of riverine fish communities sampled repeatedly over
a 23-year period (1995–2018) across 403 locations in France. By reconstructing
yearly food web structures [@danet_species_2021; @bonnaffe_comparison_2021], we
address three key objectives: (i) documenting how temporal biodiversity trends
reshape food web structure, (ii) connecting these changes to established
theories of food webs and Biodiversity-Ecosystem functioning relationships, and
(iii) assessing the validity of the space-for-time substitution assumption in
the context of biodiversity and food web structure changes.

The structure of ecological networks, which describes the number, distribution,
and strength of interactions among species, is critical for understanding
species coexistence, ecosystem functioning, and stability to perturbations
[@chesson_interaction_2008; @thebault_food-web_2003; @oconnor_general_2017;
@danet_species_2021; @eschenbrenner_diversity_2023; @danet_response_2024]. Both
theory and empirical evidence demonstrate that species richness and food web
structure influence coexistence [@chesson_interaction_2008], ecosystem
productivity [@thebault_food-web_2003; @wang_biodiversity_2018], temporal
stability [@danet_species_2021; @danet_response_2024;
@eschenbrenner_diversity_2023], and robustness to disturbances
[@sole_complexity_2001; @dunne_network_2002; @allesina_functional_2009;
@bodini_using_2009; @montoya_press_2009; @sanders_trophic_2018]. For instance,
keystone predators can regulate dominant competitors, preventing
competitive exclusion and maintaining species richness [@heithaus_predicting_2008;
@johnson_rarity_2007; @estes_trophic_2011]. Similarly, food
webs with higher trophic levels are associated with greater ecosystem
productivity [@wang_biodiversity_2018; @schneider_animal_2016;
@wang_intraguild_2019; @nie_will_2023] and community stability
[@danet_species_2021; @eschenbrenner_diversity_2023; @danet_response_2024],
while the number and distribution of trophic links determine resilience to
perturbations [@dunne_network_2002; @allesina_who_2004; @bodini_using_2009;
@allesina_functional_2009; @sanders_trophic_2018]. Understanding how biodiversity changes
reshape ecological networks is therefore essential for quantifying and
mitigating anthropogenic impacts on ecosystems.

Three key components link biodiversity changes to food web structure. First, the
energetic paradigm of food web dynamics posits that the presence of top
consumers depends on sufficient resource availability at lower trophic levels,
driven by bottom-up productivity [@lindeman_trophic-dynamic_1942;
@gravel_trophic_2011]. Thus, the biomass of generalist top consumers is likely
constrained by overall community biomass. Second, connectance—the proportion of
realized trophic links—typically decreases with increasing species richness, as
consumers are limited in the number of resources they can exploit
[@oconnor_general_2017; @dunne_network_2006]. However, higher species richness
is expected to increase the redundancy of trophic pathways, enhancing ecosystem
robustness to species extinctions [@dunne_network_2002; @allesina_who_2004;
@bodini_using_2009; @allesina_functional_2009]. This redundancy underpins the
insurance effect of biodiversity, where greater diversity buffers ecosystems
against perturbations and environmental changes [@yachi_biodiversity_1999;
@loreau_ecological_2022; @danet_response_2024]. Finally, biodiversity-ecosystem
functioning theory suggests that species richness promotes community biomass
through niche complementarity [@loreau_biodiversity_2001]. In food webs, the
presence of higher trophic levels reduces niche overlap among producers, further
boosting productivity and biomass [@schneider_animal_2016;
@wang_biodiversity_2018; @wang_intraguild_2019; @nie_will_2023]. Together, these
mechanisms highlight how biodiversity changes can alter trophic link
distribution, robustness, and ecosystem functioning.

To investigate how temporal biodiversity changes correlate with changes in food
webs, we analyzed temporal trends in species richness, biomass, and food web
structure across `r nb_op` sampling events from `r nb_st` fish communities (Fig.
\@ref(fig:matmet)a-c, Table S2). Using structural equation modeling (SEM), we
inferred the relative effects of biomass and species richness temporal trends on
trends of food web metrics, as well as the effect of species richness trends on
biomass trends (Fig. \@ref(fig:matmet)d). Following a classical approach of
network analysis [@dunne_network_2002], we also analyzed the robustness to
species extinction of the meta-food web (comprising all trophic species and
their interactions) to identify species extinction scenarios that could explain
observed variations in food web structure. To test the space-for-time
hypothesis, we compared temporal findings with spatial gradients using two
approaches: (i) replicating our temporal analysis using median biodiversity and
food web metrics across sites to remove temporal trends (Fig. \@ref(fig:matmet)c
& e), and (ii) simulating food web dynamics across gradients of species richness
and primary productivity using a bioenergetic model [@brose_allometric_2006;
@lajaaiti_ecologicalnetworksdynamicsjl_nodate]. These simulations allowed us to
evaluate whether theoretical predictions align with our empirical data, which
are limited to fish biodiversity.

(ref:fig-met-cap) Methodological Approach: (a) We analyzed `r nb_st` stream fish
communities in France, each monitored for at least 10 years. (b) For each site
and sampling event, we reconstructed food webs using a metaweb based on
ontogenetic diet information and predator-prey body size ratios. (c) We computed
temporal trends and median values for community biomass, species richness,
connectance, average trophic level, and redundancy at each site. (d) We then
assessed relationships among these food web descriptors over time and (e) across
spatial gradients.

```{r matmet, fig.cap="(fig-met-cap)", out.width="100%"}
knitr::include_graphics(get_mypath("figures", "mat_met.png"))
```

# Results

```{r, results=FALSE}
gbm <- get_effect_ci(resp = "log_bm_std", ci_lvl = "level:0.80", p = TRUE)
gr <- get_effect_ci(resp = "log_rich_std", ci_lvl = "level:0.80", p = TRUE)
gt <- get_effect_ci(resp = "w_trph_lvl_avg", ci_lvl = "level:0.80", p = FALSE)
gct <- get_effect_ci(resp = "connectance", ci_lvl = "level:0.80", p = FALSE)
gl <- get_effect_ci(resp = "prop_redundant_links", ci_lvl = "level:0.80", p = FALSE)
```

```{r}
tar_load(c(r2_inla_tab, r2_inla, r2_sem))
r2_inla_report <- r2_inla %>%
  select(response, r2_mvp_marg95hpdci, r2_mvp_cond95hpdci, r2_mvp_cond_mean)
r2cond <- map(
  setNames(r2_inla_report$response, r2_inla_report$response),
  ~round(r2_inla_report[r2_inla_report$response == .x, ]$r2_mvp_cond_mean, 2)
)
r2ic <- map(
  setNames(r2_inla_report$response, r2_inla_report$response),
  ~r2_inla_report[r2_inla_report$response == .x, ]$r2_mvp_cond95hpdci
)
```

```{r}
tar_load(summary_pred_station_tps10)
dfv <- function(
  x = summary_pred_station_tps10,
  r = "connectance",
  v = "min",
  d = 4
) {
  out <- x[x$response == r, ][[v]]
  format(round(out, d), nsmall = d)
}
grbmn <- dfv(r = "log_bm_std", v = "min", d = 0)
grbmx <- dfv(r = "log_bm_std", v = "max", d = 0)
grrhn <- dfv(r = "log_rich_std", v = "min", d = 0)
grrhx <- dfv(r = "log_rich_std", v = "max", d = 0)
grctn <- dfv(r = "connectance", v = "min", d = 2)
grctx <- dfv(r = "connectance", v = "max", d = 2)
grtln <- dfv(r = "w_trph_lvl_avg", v = "min", d = 2)
grtlx <- dfv(r = "w_trph_lvl_avg", v = "max", d = 2)
grln <- dfv(r = "prop_redundant_links", v = "min", d = 2)
grlx <- dfv(r = "prop_redundant_links", v = "max", d = 2)
```

## Estimating temporal trends

Our food web reconstruction delivered `r nb_op` annual estimates of species richness,
biomass, connectance, weighted average trophic level, and proportion of
redundant trophic links (hereafter “redundancy”, see Appendix 1,
Fig. S1, Table S2) from the `r nb_st`
sampling sites with 10+ years of community dynamics (Fig. \@ref(fig:matmet)a-c). We used
ontogenic food diet data at the species level and individual fish body size to
reconstruct trophic species food webs, i.e. at the size class species level (See
Methods). We implemented Bayesian hierarchical models accounting for spatial
variation at both hydrographic river basin and site levels to jointly estimate
the temporal trends in these metrics (see Methods). We found a great
heterogeneity in the temporal trends at the site level (community biomass: from
`r grbmn`% to +`r grbmx`% per decade, species richness: `r grrhn` to
+`r grrhx`%, average trophic level: `r grtln` to +`r grtlx`, connectance:
`r grctn` to +`r grctx`, redundancy: `r grln` to +`r grlx`, Fig.
S2) which was effectively captured by the hierarchical model (conditional R2
ranging from `r r2cond$log_bm_std` for community biomass to
`r r2cond$log_rich_std` for species richness, Table S3).

(ref:sem-cap) (a) Structural equation models linking the temporal trends of
biodiversity and food web structure. Dashed and solid arrows display
respectively negative and positive the standardized slope coefficients, their
widths being proportional to the absolute values of the coefficients. The
standardized slope coefficients and their associated 95% confidence intervals
obtained by bootstrapping are displayed along the path. $R^2$ displays the marginal
R-squared of the linear models. (b to h) Direct bivariate relationships derived
from the structural equation model. Points display the temporal trends at each
site estimated with a hierarchical Bayesian model, lines display the linear
relationship estimated with the SEM. The distribution of the site temporal
trends are displayed in Fig. S2. All the effects are reported in Table S4.
Average trophic level is weighted by the biomass of the trophic species.
Redundancy is the proportion of redundant links.

```{r sem, fig.cap = "(ref:sem-cap)", out.width = "100%"}
tar_load(p_sem_tps_file)
knitr::include_graphics(p_sem_tps_file)
```

```{r}
tar_load(c(semeff_tot, semeff_tot_ok))
semeff_tot <- semeff_tot %>%
  mutate(across(c(response, predictor), ~str_remove(.x, "mean_")))

rdtps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std")
    )
)
rdsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", model = "spatial"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", model = "spatial")
    )
)
rdm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Direct", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Direct", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
```

```{r}
ridtps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "indirect")
    )
)
ridsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "indirect")
    )
)
ridm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Indirect", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Indirect", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
rttps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "total")
    )
)
rtsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std",
      model = "spatial", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "total")
    )
)
rtm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Total", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Total", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
```

## Association between biodiversity and food web temporal trends

These temporal trend estimations at the site level were the input to our SEM
which accounts for uncertainties in trend estimation (see Methods). It allows us
to evaluate the direct effect of the temporal trends of species richness and
community biomass on the temporal trends of food web structure, the direct
effect of species richness temporal trends on the temporal trends of community
biomass, and the indirect effect of species richness temporal trends on the
temporal trends in food web structure through community biomass. We detected a
strong positive effect of the temporal trends in species richness on the
temporal trends in community biomass ($r_\delta =$ `r rdtps$bm$r`,
Fig. \@ref(fig:sem) a and b, $r_\delta$: standardized slope coefficient, [95%
CI], Table S4). This indicates that communities with decreasing, or increasing,
species richness over time experienced a concomitant decrease, or increase, in
community biomass.

We also find positive effects of the temporal trend in community biomass on the
temporal trends of connectance, average trophic level, and redundancy (resp.
$r_\delta =$ `r rdtps$ct$bm`, `r rdtps$tlvl$bm`, `r rdtps$red$bm`, Fig.
\@ref(fig:sem)b-d). Thus, when communities experienced a decrease, or increase,
in community biomass, they also experienced a concomitant decrease, or increase,
in connectance, average trophic level and redundancy. To assess whether top
trophic level species were linked to those relationships, we additionally looked
at the temporal trends in the number and the proportion of piscivorous species.
The temporal trends in community biomass were positively associated with the
temporal trends in both the number and the proportion of piscivorous species
(Fig. S3a). It suggests that high trophic level fishes are strongly linked to
temporal trends in community biomass.

The total effect of the temporal trends in species richness on the temporal
trend in connectance was negative ($r_\delta =$ `r rttps$ct$r`, Fig.
\@ref(fig:std)a), resulting from a negative direct effect ($r_\delta =$
`r rdtps$ct$r`, Fig. \@ref(fig:sem)f) and a positive indirect effect via changes
in the temporal trends in community biomass ($r_\delta =$ `r ridtps$ct$r`). Thus, when
communities experienced an increase, or decrease, in
species richness, they also experienced a concomitant decrease, or increase, in
connectance. In contrast, the total effects of the temporal trends in species
richness on the temporal trend of both trophic level and redundancy were
positive (resp. $r_\delta =$ `r rttps$tlvl$r`, `r rttps$red$r`, respectively,
Fig. \@ref(fig:std)a). The total positive effect on the temporal trend of
trophic level resulted from a combination of negative direct and positive
indirect effect while the effect on the temporal trend of redundancy resulted
from both positive direct and indirect effects (Fig. \@ref(fig:std)a). All together, this
means that communities that have experienced decline in species richness over
the last decades have decreased in average trophic level and redundancy but
increased in connectance. Conversely, those that experienced species richness
increases have increased in average trophic level, redundancy, but decreased in
connectance.

## Spatial association between biodiversity and food web

Spatial analyses, either using empirical data or the bioenergetic food web
model, largely aligned with temporal trends (Fig. \@ref(fig:std)a and S4),
showing consistent positive effects of species richness on community biomass and
of community biomass on average trophic level and redundancy. Across space,
sites with higher fish biomass on average had food webs with higher connectance,
redundancy and average trophic level than sites with lower fish biomass. The
theoretical model predicted similar relations, with a particularly strong
positive association between total biomass and average trophic level across
modelled food webs. Species richness also exhibited a direct negative effect on
connectance and a positive effect on redundancy, mirroring temporal patterns.
Both modelled food webs and local fish communities characterized by greater
species richness on average showed higher average biomass and food web
redundancy, and lower connectance. However, spatial and temporal analyses
diverged in their effects of species richness on average trophic level, with
spatial data showing a negative effect and temporal trends indicating a positive
one. Detailed statistical effects and comparisons are provided in the Appendix.

## Robustness of food web structure to species extinctions

To investigate whether top species were drivers of biomass and structure
patterns (see section 4.2), we implemented three secondary extinction simulation
scenarios on the empirical metaweb defined as a food web containing all the
trophic species, i.e. the fish species subdivided in size classes added to their
resources, found across all sites in the empirical data (See Methods). When
trophic species were removed in a sequential order beginning by the highest
trophic levels, connectance, average trophic level, and redundancy decreased
(purple lines, Fig. \@ref(fig:std)b). On the other hand, removing low level
trophic species first increases connectance and average trophic level, but does
not affect redundancy (yellow lines, Fig. \@ref(fig:std)b). In turn, species
extinction at random had little effect on food web structure (black line, Fig.
\@ref(fig:std)b). These results, in conjunction with our empirical analyses
above, suggest that top consumers play a disproportionate role in driving the
changes in food web structure reported.

(ref:fig-std-cap) a, Direct and total effects of community biomass and species
richness trends on food web structure trends. Standardized coefficients (r)
derived from the temporal, spatial, and bioenergetic model Structural Equation
Models. Points and lines display the estimated effects and their confidence
intervals at 95%. Values are displayed in Table S4, R-squared in Table S4, and
the visualization of the bivariate relationships for the spatial and
bioenergetic model SEMs are displayed in Figure S4. Average trophic level is
weighted by the biomass of the trophic species. Redundancy is the proportion of
redundant links. b, Effect of species extinctions on the structure of the
empirical metaweb according to three scenarios. In the Decreasing and Increasing
scenarios, species were ordered by their trophic level and removed from the
metaweb starting respectively from the highest and lowest trophic level. Species
were removed in a random sequence (50 sequences, grey dots) for the random
scenario, the black line displaying the average predicted by a GAM linear model.
For the extinction sequence, the average trophic level was not weighted by the
biomass of species as this information is not available at the metaweb scale.
Similar results were obtained with trophic species (i.e. node) extinction (Fig.
S5).

```{r std, fig.cap="(ref:fig-std-cap)", out.width = "100%"}
tar_load(p_sem_tot_ext_sp_file)
knitr::include_graphics(p_sem_tot_ext_sp_file[[1]])
```

# Discussion

Our findings reveal several key patterns: (1) temporal trends in species
richness were positively associated with trends in community biomass; (2)
temporal trends in community biomass were positively linked to trends in
connectance and average trophic level; (3) temporal trends in species richness
were negatively associated with trends in connectance and only weakly related to
trends in average trophic level; and (4) both community biomass and species
richness trends were positively associated with trends in trophic pathway
redundancy. Additionally, we found evidence that (5) top predator species play a
significant role in driving these patterns. Importantly, these temporal
relationships were largely consistent with patterns observed across spatial
gradients, both in empirical data and in simulations from a bioenergetic food
web model. In the following, we discuss these findings in the context of food
web assembly theory and Biodiversity-Ecosystem Functioning theory, and consider
their broader ecological implications.

We found that community biomass scaled positively with species richness, a
pattern consistent across our temporal, spatial, and theoretical analyses. This
result aligns with previous studies on fish communities across space
[@woods_testing_2020; @danet_species_2021] as well as grasslands, salt marshes,
and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018; but see @grace_integrative_2016].
These empirical findings are supported by theory, suggesting that food web
diversity enhances biomass by stimulating primary production and resource uptake
[@schneider_animal_2016; @wang_biodiversity_2018; @wang_intraguild_2019;
@nie_will_2023], and by Biodiversity-Ecosystem Functioning theory, which
predicts that species loss reduces community biomass
[@loreau_biodiversity_2001]. Our results bring a temporal dimension to these
findings using long-term trends in empirical data, and clearly exemplify that
declines in species richness goes along with a reduction of biomass in
ecosystems. Conversely, our result also indicates that conservation actions
promoting species richness should also promote community biomass.

Our results also reveal that food webs experiencing biomass and/or species
richness declines undergo significant structural changes, including the loss of
top predators and a reduction of connectance. These findings align with
energetic and trophic biogeography theories [@lindeman_trophic-dynamic_1942;
@gravel_trophic_2011], which posit that the presence of high-trophic-level
species is constrained by the availability of energy and resources
[@post_ecosystem_2000]. Thus, a decline in community biomass may lead to
insufficient resources to sustain top predators, while an increase in biomass
could provide the necessary energy to support these species
[@takimoto_ecosystem_2008; @post_ecosystem_2000]. Further,
Biodiversity-Ecosystem Functioning theory suggests that the presence of higher
trophic levels may reduce niche overlap among producers, enhancing productivity
and community biomass [@schneider_animal_2016; @wang_biodiversity_2018;
@wang_intraguild_2019; @nie_will_2023]. The relationship between community
biomass and connectance is less clear, with a weaker positive association
compared to average trophic level. Previous theoretical studies have reported
contrasting effects of connectance on productivity [@nie_will_2023;
@wang_intraguild_2019; @thebault_food-web_2003; @poisot_trophic_2013;
@schneider_animal_2016]. For instance, while increased connectance between
herbivores and plants can either enhance or reduce productivity
[@thebault_food-web_2003; @poisot_trophic_2013; @nie_will_2023], greater
connectance among carnivores generally boosts productivity by reducing top-down
control and allowing biomass to accumulate across trophic levels
[@wang_intraguild_2019; @schneider_animal_2016]. Interestingly, our theoretical
model predicts overall a positive relation between total biomass and
connectance, which contrasts with the negative relation found recently by
another complex allometric food web model [@nie_will_2023], but this relation is
far weaker than the strong positive relation between total biomass and average
trophic level.

Our study suggests that declines in community biomass and species richness
are associated with a loss of redundant trophic links, while their increase
results in greater redundancy. Since redundant links are those that can be
removed without triggering secondary extinctions [@allesina_who_2004;
@bodini_using_2009], our findings suggest that declines in community biomass and
species richness might reduce food web robustness to further perturbations
[@allesina_functional_2009]. Such positive effect of species richness on
redundancy would then further support the insurance hypothesis
[@yachi_biodiversity_1999; @loreau_biodiversity_2021], where greater species
richness and functional redundancy buffer ecosystems against perturbations,
increasing stability [@sanders_trophic_2018]. Furthermore, our results highlight
the disproportionate role of top trophic-level species in maintaining redundant
links within food webs. This pattern is consistent with previous empirical
studies showing that top predators are often the most generalist and highly
connected species [@montoya_press_2009; @montoya_ecological_2006]. Consequently,
disturbances affecting these top predators are likely to have the greatest
impact on overall community stability [@montoya_press_2009]. These findings
underscore the importance of preserving top predators and maintaining species
richness to safeguard food web robustness and ecosystem stability.

Species richness was negatively associated with connectance, which is consistent
with previous findings and likely driven by limitations in predator generality
[@dunne_network_2006]. In contrast, the effects of species richness on average
trophic level were more nuanced and considerably weaker than those of community
biomass. Direct effects of species richness on average trophic level were either
negligible or negative, as observed in temporal, spatial, and theoretical
models. However, when accounting for the indirect effects of species richness —
mediated through changes in community biomass — our temporal and theoretical
analyses revealed a positive association between species richness and average
trophic level. This suggests that while species richness influences food web
structure, its effects are more pronounced on connectance than on average
trophic level dynamics.

We found strong agreement between our temporal and spatial analyses, as well as
with predictions from food web theory and Biodiversity-Ecosystem Functioning
research. This suggests that spatial comparisons of generic topological food web
properties and biodiversity metrics may serve as reasonable surrogates for
predicting temporal trends—supporting the validity of the space-for-time
substitution in both empirical data and bioenergetic food web models. However,
this agreement did not hold for the relationship between weighted average
trophic level (hereafter “average trophic level”) and species richness. This
discrepancy may arise because average trophic level incorporates both food web
topology and biomass distribution, which are influenced by dynamic processes
such as bottom-up and top-down control. Indeed, biomass distribution across
trophic levels depends on  species metabolism and productivity which in turn
control the strength of top-down and bottom-up processes
[@brose_allometric_2006; @barbier_pyramids_2019].

Our study provides compelling evidence that biodiversity trends are profoundly
reshaping food web structure. By linking these patterns to
Biodiversity-Ecosystem Functioning and food web theory, we demonstrate that
widespread biodiversity declines likely have significant and detrimental
consequences for ecosystem functioning and stability in the face of future
perturbations. These findings suggest that assessments of food web
structure—based on metaweb reconstruction like in the present study—, added to
more classic metrics of community biomass, and species richness could serve as
valuable indicators of functional decline and metrics for evaluating the
efficacy of mitigation strategies. Finally, our study also highlights critical
avenues for future research. For instance, integrating food web metrics into
conservation planning will require understanding how they relate to other facets
of biodiversity, such as the presence of non-native species, and drivers of
biodiversity changes, including habitat degradation [@danet_past_2024] and
climate change [@bonnaffe_interaction_2024].

# Material & Methods

We used a monitoring of freshwater fishes conducted all over France by the
Office Français de la Biodiversité (OFB). The sampling protocols were
standardized since 1995 and a complete description of the sampling process can
be found in @danet_species_2021. The fishes are sampled using electrofishing all across the
streams in shallow streams and over the bank by boat in deeper streams. All the
fishes are counted and the body size of fishes are recorded on a sample of
individuals, enabling to estimate the distribution of individual body size by
species for each sampling [@bonnaffe_comparison_2021; @danet_species_2021].

We selected the sites that have been monitored for at least 10 years in the
1995-2018 period, with the same sampling protocol over time. We selected one
sampling per year by site. In each site, we ensured low temporal heterogeneity
within the sampling period by selecting samplings that were performed in the
same period, leading to a mean standard deviation of sampled month in each site
of three weeks [@danet_species_2021]. Those steps lead to the selection of `r nb_st` sites (Fig.
\@ref(fig:matmet)a, Table S2).

## Biomass

We first estimated the body mass of each fish individual with the allometric
law: $B_i = 0.01 \times l_i^{3.03}$, $B_i$ being the biomass expressed in gram
and li the body length of the individual $i$ expressed in millimeters
[@froese_cube_2006; @danet_species_2021]. Then the biomass of each species and
community were obtained by summing the body mass of individuals. This resulted
in a biomass score for the whole community of each site. Finally, the biomass
was reported to the sampled surface (in $m^2$) to account for sampling effort.

## Food web reconstructions

We reconstructed food web structure in two steps [@bonnaffe_comparison_2021;
@danet_species_2021]: (1) we first computed the metaweb, describing the
potential trophic interactions among all size classes species, (2) then we
extracted yearly species size classes and their interactions according to their
presence in a given sampling event, to establish a annual snapshot of food webs.

We built the metaweb according to ontogenic diet data and body size
[@gravel_inferring_2013; @poisot_trophic_2013; @brose_predator_2019;
@bonnaffe_comparison_2021; @danet_species_2021; @bonnaffe_interaction_2024]. The
method is extensively described in @bonnaffe_comparison_2021 and
@danet_species_2021. The food diet of stream fishes changes along their life
cycle or stage, therefore according to their body size. Each fish species was
divided in nine size classes, i.e. trophic species, equally distributed in their
observed body size range. Thus, a trophic species is a group of individuals of
the same species and size class that are expected to share the same prey and
predators because trophic interactions are largely determined by predator-prey
size ratio [@brose_allometric_2006; @brose_predator_2019]. We also added seven
nodes present in the ontogenic food diet database: detritus, biofilm,
phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos. The
resulting metaweb contained 45 fish species divided into 9 size classes plus 7
resource nodes, summing up to 412 nodes. Consideration of size classes, i.e.
trophic species instead of species, improves inference of trophic interactions
[@jennings_weak_2001].

We determined the feeding interactions of trophic species according to their
species identity and their body size. The diet database described two or three
life stages by fish species, each one being delimited by body size. The
ontogenic fish diet database was filled according to fishbase
[@froese_fishbase_2021] and literature [@bonnaffe_comparison_2021;
@danet_species_2021].

The feeding interactions among fishes were based on the body size ratio between
predators and preys. We defined the predation window of a piscivorous trophic
species as 3% to 45% of its midpoint [@mittelbach_ontogeny_1998;
@claessen_impact_2002; @hart_handbook_2008]. A trophic link was defined between
the piscivorous trophic species and all the trophic species whose midpoint in
body size range was included in the predation window. The fish-resource trophic
links were set between the trophic species and the resource nodes according to
the food items of their size class. Finally, the feeding interactions between
resources were defined according to the literature [@allan_stream_2007;
@hart_handbook_2008].

## Community and food web structure

At each site, yearly food webs were described by their community biomass and
species richness. Yearly food web structure was described by their connectance,
describing the degree of generalism in a network [@dunne_network_2006;
@delmas_analysing_2019]. Connectance was computed as the ratio between the
number of actual links and the number of potential links $C = \frac{L}{S(S-1)}$,
with $L$ being the number of links and $S$ the number of trophic species). We
also computed the average trophic level, weighted by the biomass of the trophic
species. The trophic level of a trophic species was computed recursively as 1
added to the mean trophic level of the preys it consumes
[@herendeen_ecological_2008]. We computed the number of redundant and functional
links using the dominator tree approach [@allesina_functional_2009]. A species
$i$ is a dominator of species $j$ if species $i$ mediates the connection of
species $j$ to basal resources such that the extinction of i removes the
connection of $j$ to basal resources. It is then used to identify the set of
trophic links that are essential to connect species to basal resources, i.e.
functional links, and the redundant links. This method requires a single basal
root in the food webs. To do so, we added a basal dummy resource node to the
empirical food webs connected to biofilm and phytoplankton resource nodes. For
food webs generated from theoretical simulations (see below), we added a
resource node connected to all primary producers. We computed dominator trees
using the `dominator_tree()` function from the `igraph` R package (See Appendix
for details). We computed the number of functional links as the number of links
in the resulting dominator tree of the food web. We checked that the procedure
reproduced the examples presented in previous studies
[@allesina_functional_2009, Appendix]. We then derived the proportion of
redundant links in food webs of empirical and theoretical simulations.

## Temporal trends of food web

We modeled the temporal trends of food web structure at each site ($Y_i$) using
a linear hierarchical model which accounts for the hierarchical spatial structure
of the data, i.e. the sites nested in hydrographic basins. Time was considered as the number of years since the
first sampling ($t = 0$) at each site. $\alpha$ is the intercept and $\beta$ the
slope estimating the temporal trends. The intercept and the slope parameters
have a fixed ($\alpha_0$, $\mu$) part and a random part. We accounted for the
spatial structure of the data by adding random intercepts ($a$) and slopes ($b$)
for the basin ($a_n$, $b_n$) and site ($a_{i|n}$, $b_{i|n}$) identity, site
effects being nested in the basin one.

$$
\begin{aligned}
    Y_{i|n, t} &= \alpha + \beta Time_t + \epsilon_{i|n, t}
  (\#eq:gen)
\end{aligned}
$$

- $\alpha = \alpha_0 + a_n + a_{i|n}$
- $\beta = \mu + b_n + b_{i|n}$
- $a_n, a_{i|n}, b_n, b_{i|n}, \epsilon_{i|n, t} \sim \mathcal{N}(0,\,\sigma^{2})$
- $n$: basin, $i$: site $i$, $t$: time $t$

The model parameters were estimated using INLA (Integrated Nested Laplacian
Approximation) Bayesian inference with default uninformative priors
[@rue_bayesian_2017]. The prior distributions for the fixed intercept
($\alpha_0$) and slope ($\mu$) were modeled as a flat normal distribution with a
mean of zero ($\mathcal{N}(0,\, 1000)$). Prior distributions for the Gaussian
error $\epsilon_{i|n, t}$ and the random effects were modeled as a log gamma
distribution with shape and inverse scale parameters ($\mathcal{G}(s, \, \tau) =
\mathcal{G}(1, \, 5.10^-5)$), then their respective standard deviation were
obtained by back-transforming the inverse scale parameter ($\sigma = 1 /
\sqrt{\tau}$). Community biomass and species richness were log-transformed.

The performance of the model was assessed visually by plotting the fitted versus
the observed values (Fig. S6). We also computed the marginal ($R^2_m$) and conditional
($R^2_c$) R-squared of the models following previous formulations
[@nakagawa_general_2013; @lahuis_explained_2014]. In order to account for the
variability associated with the priors [@gelman_r-squared_2019], we computed the
variance of the posterior distribution  of each predicted values ($Var_{fit} =
\sigma^2(\hat{Y_i})$). We computed $R^2$ values for each predicted values, then
obtaining a $R^2$ distribution [@gelman_r-squared_2019]. We reported mean
marginal and conditional $R^2$ associated the 95% credible interval computed
using the Highest Posterior Density method.

$$
R^2_m = \dfrac{Var_{fit}}{Var_{fit} + Var_{res}} = \dfrac{\sigma^2(\hat{Y_i})}{\sigma^2(\hat{Y_i}) + \sigma^2(Y_i - \hat{Y_i})}
$$

$$
R^2_c = \dfrac{Var_{fit} + (a_n)^2 + (a_{i|n})^2}{Var_{fit}  + (a_n)^2 + (a_{i|n})^2 + Var_{res}}
$$

$y_i$ and $\hat{y_i}$ being respectively the observations and the
predicted values, $Var_{fit}$ and $Var_{res}$ being respectively the variance of
predictive means and the variance of the residuals. $a_n$ and $a_{i|n}$ are
respectively the standard deviation on the random intercept associated to the
basin and the site.

We extracted the annual temporal trends at each site from the hierarchical model
using mean Best Linear Unbiased Prediction (BLUP) values, then multiplied
by 10 to get temporal trends by decade. For community biomass and species
richness which were log-transformed, we derived temporal trends in percentage
change by decade by back-transforming the slopes ($\beta\prime = (e^{\beta } - 1)
\times 100 \times 10$).

Overall, we found that standard deviation attributed to site effects in both
baseline (i.e. the intercept) and temporal trends in community structure were much higher than the
ones found for hydrographic basin (Table S6).

## Spatial

We assessed the validity of the space-for-time substitution hypothesis by
comparing the relationships estimated among food web structure in time to the
ones inferred in space. We used the same sites as in the temporal analysis, but
we summarized the temporal series of each station by the median value over time
of each facet of food web structure described above (Fig. \@ref(fig:matmet)).

## Statistical analysis

We modeled the relationships among food web structure indices, i.e. connectance,
average trophic level, the proportion of redundant trophic links, and community
biomass and species richness, with Structural Equation Models. We modeled
species richness as related to community biomass and food web structure. In
turn, community biomass was related to food web structure. We computed one
structural equation model to assess the relationships among the temporal trends
and one for the relationships over the spatial gradient, we added a random
effect of the hydrographic basin on the intercept for the latter one.

In order to control for the uncertainties of the estimated temporal trends, we
weighted each temporal trend by the inverse of the sum of the standard deviation
of the estimated trends, i.e. the estimated slopes for the response and the
explicative variables ($w_j = 1 / \sum{s_j}$, $w_j$ being the weight, and $s_j$
the standard deviation on the temporal trends of the variables).

The coefficients of the Structural Equation Models were computed with
PiecewiseSEM R package [@lefcheck_piecewisesem_2016]. The coefficients were standardized and the
confidence intervals were computed with semEff R package [@murphy_semeff_2021]. Confidence
intervals were obtained by bootstrapping the linear models 1000 times and
computing adjusted bootstrap percentile intervals (BCa), which are adjusted for
the asymmetric distribution of resampling. The marginal and conditional R2
associated with the linear models of the SEMs were computed with the Nagelkerke
method [@lefcheck_piecewisesem_2016, Table S5].

All the statistical models displayed low multicollinearity (Variance Inflation
Factors < 3, Table S7). We visually inspected the residuals of the linear
models.

## Extinction simulation in the metaweb

We simulated extinction sequences in the metaweb according to three scenarios to
get expectations on how the food web structure should change according to the
addition and deletion of species (results presented in Fig. 3b) and trophic
species (results in Appendix), i.e. species subdivided in size classes. The
scenario of extinction sequences were related to the trophic level of the
species/trophic species. In the decreasing and increasing scenario, we removed
species/nodes in decreasing, resp. increasing order, of their trophic level,
i.e. respectively beginning by the highest and lowest trophic level. In the
random extinction sequence, we built 50 random extinction sequences. For each
simulation, we removed fish species/nodes that lost all their prey. We then
computed the connectance, the average and the maximum trophic level after each
extinction.

## Theoretical simulations

```{r}
tar_load(sim_param)
nsim <- nrow(sim_param)
nfw <- length(unique(sim_param$fw_id))
s_range <- c(min = min(sim_param$S), max = max(sim_param$S))
c_range <- c(min = min(sim_param$C), max = max(sim_param$C))
k_range <- c(min = min(sim_param$K), max = max(sim_param$K))
```

We used a bioenergetic food web model to simulate the relations between
community biomass, species richness and food web structure. This food web model
lifts one of the limitations of our data, that is it includes the biomass of all
species in the food webs from primary producers to top predators. This model
allows simulating complex communities because its parameterization depends
mainly on species body masses. We first generated `r nfw` food webs varying in species
richness (range: `r s_range["min"]`-`r s_range["max"]`)
and connectance (range: `r c_range["min"]`-`r c_range["max"]`) using the niche
model algorithm, which reproduces food web topology typically found in empirical
food webs [@williams_simple_2000]. We set species body masses ($M_i$) using a predator-prey
mass ratio of 100 ($Z$) [@brose_allometric_2006; @brose_predator_2019], such as $M_i =
Z^{(TL_i - 1)}$, $TL_i$ being the trophic level of the species $i$. The
metabolic rates of the species ($x_i$) were then derived from allometric
relationships between metabolism and body mass empirically estimated
[@brose_allometric_2006], such as $x_i = 0.88 \times M_i^{-0.25}$.

The dynamics of species biomasses are described by an ODE systems of two
equations, for primary producers (eq. \@ref(eq:prod)) and consumers (eq.
\@ref(eq:cons)) respectively:

\begin{equation}
    (\#eq:prod)
	\frac{dB_i}{dt} =
    r_i B_i G_i
    - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
\end{equation}

\begin{equation}
    (\#eq:growth)
    G_i = (1 - \dfrac{\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j}{K_i})
\end{equation}

The primary producer biomass increases with a growth rate ($r_i$) of 1 and a
logistic growth ($G_i$, eq. \@ref(eq:growth)), growth being limited by intraspecific
($\alpha_{ii}$, set to 1), interspecific ($\alpha_{ij}$, set to 0.5) competition,
and carrying capacity ($K_i$, varying in the simulations). Primary producers
loose biomass by being eaten by consumers at a rate that depend on consumer
metabolic rates ($x_j$), attack rates ($y_j$, set to 4
@lajaaiti_ecologicalnetworksdynamicsjl_nodate), a functional response
($F_{ji}$), and assimilation efficiency accounting for the fact that
consumers assimilate better the primary producer biomass than the consumer ones
($e_{ij}$, set to 0.45 for herbivores and 0.85 for carnivores,
@lajaaiti_ecologicalnetworksdynamicsjl_nodate).

\begin{equation}
    (\#eq:cons)
        \frac{dB_i}{dt} =
        \sum_{j \in \{\mathrm{res}\}_i} x_i y_i B_k F_{ij}
        - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
        - x_i B_i
\end{equation}

Consumers gain biomass by eating other species and lose biomass by being
consumed and by metabolic loses ($x_i$, eq. \@ref(eq:cons)). The functional
response (eq. \@ref(eq:holling)) of consumer feeding depends on the preference
for its preys ($\omega_{ij}$, set even across consumer preys) and
half-saturation rate ($B_0$, set to 0.5). The hill coefficient ($h$, set to 1.5)
determines the shape of the functional response (Holling's type II or III).

\begin{equation}
    (\#eq:holling)
    F_{ij} = \frac{\omega_{ij} B_j^h}
    {B_0^h + \sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h}
\end{equation}

We simulated the dynamics of the `r nfw` food webs across a gradient of global
carrying capacity ($K$, range: `r k_range["min"]`-`r k_range["max"]`) to
represent variation in space of productivity. The carrying capacity were
standardized according to the number of producers and interspecific competition
($K_i = \frac{1 + \alpha_{ij} (S - 1)}{S}K$) to prevent changes in productivity
due to system size and complementarity among producers [@ives_stability_2000;
@loreau_species_2008]. We ran simulations for 5000 timesteps. At the end of
simulations, we removed dead species (biomass < 10^-6) and measured community
biomass, species richness and food web structure. If the final community
contained disconnected producers (i.e. without consumers) or disconnected
consumers (i.e. consumers without preys), we discarded them and run the
simulations again with the same parametrization. The simulations were done using
the Julia package `EcologicalNetworksDynamics.jl`
[@lajaaiti_ecologicalnetworksdynamicsjl_nodate]. Then, we computed the community
and food web metrics at the last timesteps and analyzed the data with the same
SEM described above.

# References  {-}

<div id="refs"></div>

\newpage