### 2025-04-16 README.Rmd

---
output:
  md_document:
    variant: markdown_github
---

# Assembly/Disassembly of food-web according to changes in species richness and community biomass