### 2025-04-16 paper/cover_letter.Rmd

---
author:
- On behalf of all the authors, Dr. Alain Danet
opening: Dear Editor in Chief and Associate Editors,
closing: Sincerely yours,
fontsize: 10pt
return-address:
- Dr. Alain Danet
- School of Biosciences
- The University of Sheffield
- Western Bank, S10 2TN, Sheffield, UK
output:
  linl::linl:
    keep_tex: true

letterhead: ./misc/letterhead.png
signature: /home/alain/Documents/Administration/signature.pdf
letterhead-y-offset: 0pt
letterhead-scale: .7
colorlinks: false
---

```{r, results = "hide", include = FALSE}
Sys.setlocale("LC_TIME", "C")
```

Please find enclosed our manuscript entitled *“Biodiversity temporal trends are
reshaping food web structure and redundancy in riverine ecosystems”*, which we
submit for consideration in *PNAS*.

In the midst of the current biodiversity crisis, understanding the ecological
consequences of biodiversity changes is critical for designing and implementing
effective conservation policies. This is especially urgent in freshwater
systems, which are among the most threatened ecosystems globally. While
biodiversity temporal trends have been widely documented, their impacts on
ecosystem functioning and stability remain poorly understood. Conversely, food
web and Biodiversity-Ecosystem Functioning research have provided robust
theoretical and experimental predictions on how biodiversity loss affects
ecosystem productivity and stability. For instance, food web structural
properties such as connectance and average trophic level are predicted to
strongly influence ecosystem productivity, while the presence of redundant
trophic pathways is linked to robustness against species extinctions. However,
aligning these predictions with empirical data from natural settings remains
challenging, as it requires simultaneous temporal data on biodiversity and food
web structure, which are rarely available.

We address this challenge here and reveal several effects of biodiversity
temporal trends on food web structure and function in riverine fish communities.
Using a unique dataset of over 400 fish community time series from multiple
hydrographic basins across France, we reconstructed food web structures for each
sampling event using individual body size, ontogenetic diet data, and predation
size windows. We then related changes in biodiversity (community biomass and
species richness) to changes in food web structure (connectance, average trophic
level, and the proportion of redundant trophic pathways). Our results reveal
that declines in species richness are associated with losses in community
biomass, while declines in community biomass are linked to reduced connectance
and average trophic level. We also found that top trophic-level species play a
disproportionate role in driving these trends. Importantly, declines in both
community biomass and species richness are associated with a reduction in the
proportion of redundant trophic pathways, suggesting weakened ecosystem
robustness to future perturbations.

Our study provides a large-scale perspective on how biodiversity trends reshape
food web structure in freshwater ecosystems, offering critical insights into the
ecological consequences of biodiversity loss. Our findings align with
predictions from food web and Biodiversity-Ecosystem Functioning research and
are supported by spatial analyses and bioenergetic food web modeling, which
validate the use of space-for-time substitution in this context. These results
underscore that widely reported biodiversity changes across taxa and ecosystems
likely have profound impacts on ecosystem functioning and resilience.

As *PNAS* has been at the forefront of research on biodiversity changes and food
web structure, we are confident that our study will interest the journal’s broad
readership and stimulate new perspectives on the ecological consequences of
biodiversity loss. The methods we developed to reconstruct food webs and
quantify redundancy using graph theory metrics will be of particular interest to
researchers across multiple disciplines. Additionally, this study provides
valuable insights for freshwater scientists and conservation practitioners,
offering the first comprehensive analysis of how biodiversity changes reshape
food web structure and integrating the fields of biodiversity change, food web
ecology, and Biodiversity-Ecosystem Functioning.

The enclosed manuscript has been approved by all authors and has not been
published or submitted elsewhere. We have had no prior discussions with a *PNAS*
editor regarding this work.

We suggest the following experts as potential editors and reviewers for their relevant expertise:

- Editorial Board members:
  - Power, Mary
  - Hastings, Alan
  - Marquet, Pablo A.
- NAS members:
  - Stephen R. Carpenter
  - Peter B. Reich
  - Sandra Lavorel
  - Marten Scheffer
  - Gene E. Likens
  - Lisa A. Levin
- Qualified referees (name & email link, institution, expertise):
  - [Shaopeng Wang](shaopeng.wang@pku.edu.cn), Peking University (expertise in Biodiversity-Ecosystem Functioning in food webs)
  - [Mary O’Connor](oconnor@zoology.ubc.ca), The University of British Columbia (expertise in the effects of biodiversity changes on ecosystem functioning)
  - [Jennifer Dunne](jdunne@santafe.edu), Santa Fe Institute (expertise in the effects of biodiversity loss on food web structure)
  - [Antonio Bodini](antonio.bodini@unipr.it), Universita di Parma (expertise in food web alteration and graph theory)
  - [Mike Heithaus](CASEdean@fiu.edu), Florida International University (expertise in the ecological consequences of top predator removal)

Thank you for considering our submission. We look forward to the possibility of
contributing to *PNAS*.