### 2025-04-16 paper/method_figure.Rmd

# For Loubna

```{r}
library('targets')
library(tidyverse)
library('magrittr')
library(cowplot)

mypath <- rprojroot::find_package_root_file

source(mypath("R", "misc.R"))
source(mypath("R", "variable_shortcut.R"))
source_dir(mypath("R"))
theme_set(theme_cowplot())
```

```{r}
get_bm_dyn_year <- function (com = NULL, station = NULL) {

  st <- station
  test <- com %>%
    filter(station == st)

  test %<>%
    mutate(sp_tibble = map(sp_vector, ~tibble(species = names(.x), biomass_sp = .x))) %>%
    unnest(sp_tibble) %>%
    mutate(bm_std_sp = biomass_sp / surface) %>%
    dplyr::select(station, year, species, bm_std_sp)

  test %<>% 
    pivot_wider(names_from = species, values_from = "bm_std_sp")%>%
    mutate_at(vars(matches("[A-Z]{3}", ignore.case = FALSE)), ~ifelse(is.na(.), 0, .)) %>%
    pivot_longer(cols = matches("[A-Z]{3}", ignore.case = FALSE), names_to = "species", values_to = "bm_std_sp")

  return(test)
}
get_network_summary <- function (com = NULL, station = NULL, stab = TRUE, com_avg = TRUE) {

  st <- station
  tmp <- com %>%
    filter(station == st) %>%
    mutate_if(is.double, ~round(., 2))

  label_stab <- c(
    "Stability" = tmp$bm_std_stab,
    "Synchrony" = tmp$sync, 
    "CVsp" = tmp$cv_sp,
    "Median total biomass" = tmp$prod_std #"' * (g/m^2) * '"
    )
  label_com <- c(
    "Median species richness" = tmp$richness_med,
    "Median connectance" = tmp$ct,
    "Median avg trophic level" = tmp$t_lvl
  )
  label <- NULL

  if (stab) {
    label <- c(label, label_stab)

  }

  if (com_avg) {

    label <- c(label, label_com)

  }

  return(label)

}

plot_dyn_web  <- function (
  com = NULL, sem_df = NULL, station = NULL,
  stacked = FALSE, .log = FALSE, color_species = NULL, label = NULL,
  label_parsed = FALSE, label_size = 4.5, y_label = NULL, my_ylim = NULL) {

  # get bm dynamic 
  com <- get_bm_dyn_year(com = com, station = station)

  # get total biomass 
    total_bm <- com %>%
	  group_by(year) %>%
	  summarise(bm_std_sp = sum(bm_std_sp)) %>%
	  mutate(species = "Total")

  p <- com %>%
    ggplot(aes(x = year, y = bm_std_sp, color = species))

  if (stacked) {
    p <- p + 
      geom_area(aes(fill = species))
  } else {
    p <- p + 
      geom_line() +
      geom_line(data = total_bm, color = "black")
  }

  if ( !is.null(my_ylim) ) 
  {
    p <- p +
      ylim(my_ylim)
  }

  # Make it professional:
  p <- p +
    labs(y = expression(Biomass~density~(g/m^2)), x = "Year")

  if (!is.null(sem_df)) {

    label <- get_network_summary(com = sem_df, station = station) 
  }

  if (!is.null(label)) {

    if (is.null(y_label)) {
      y_label <- max(total_bm$bm_std_sp) + max(total_bm$bm_std_sp) * 5/100
    }

    p <- p +
    annotate("text", x = median(total_bm$year),
      y = y_label,
      label = label, parse = label_parsed, size = label_size)
  }

  if (.log) {
    p <- p + scale_y_log10()
  }

  if (!is.null(color_species)) {
    p <- p + 
      scale_fill_manual(values = color_species) +
      scale_color_manual(values = color_species)
  }

  return(p)

}

get_label_from_named_vector <- function ( label = NULL, collapse = "\n") {
  paste0(names(label), " = ", label, collapse = collapse)
} 
```

```{r}
get_adj_net_station <- function (net = NULL, station = NULL) {

  st <- station

  net <- net %>%
    filter(station %in% st) %>%
    dplyr::select(station, year, network)

  if ("data.frame" %in% class(net$network[[1]])) {

   net <- net %>%
    unnest(cols = c(network)) %>%
    dplyr::select(year, from, to) %>%
    arrange(year) %>%
    group_by(year) %>%
    nest(network = c(from, to)) %>%
    mutate(
      igraph_obj = map(network, igraph::graph_from_data_frame,
	directed = TRUE),
      adj_mat = map(igraph_obj, igraph::as_adjacency_matrix,
	sparse = FALSE),
      troph = map(adj_mat, ~NetIndices::TrophInd(.x)),
      obs_troph_level_vector = map(troph, function (x) {
	TL <- x$TL
	names(TL) <- rownames(x) 
	return(TL)
	})
    )
  } else {

    net <- net %>%
    arrange(year) %>%
    group_by(year) %>%
    rename(adj_mat = network) %>%
    mutate(
      troph = map(adj_mat, ~NetIndices::TrophInd(.x)),
      obs_troph_level_vector = map(troph, function (x) {
	TL <- x$TL
	names(TL) <- rownames(x) 
	return(TL)
	})
    )
  }

  return(net)

}

get_net_biomass <- function (net = NULL, station = NULL) {

  st <- station

  # Get bm
  net <- net %>%
    filter(station %in% st) %>%
    dplyr::select(year, composition) %>%
    unnest(cols = c(composition))  %>%
    dplyr::select(year, sp_class, bm_std)

   # Make list
    net %<>%
      arrange(year) %>%
      group_by(year) %>%
      nest(bm = c(sp_class, bm_std)) %>%
      mutate(bm = map(bm, deframe))
  net
}

set_color_nodes <- function (
  node_list = NULL,
  color_species_resources = NULL
  ) {

  node_list_sp <- stringr::str_replace_all(node_list, "_\\d", "")

  colour <- color_species_resources[node_list_sp]
  names(colour) <- node_list
  return(colour)
}

set_node_resources_abun <- function (
  node_list = NULL,
  node_abun = NULL, resource_abun = NULL) {

  abun <- vector("numeric", length(node_list))  
  names(abun) <- node_list

  if (is.null(resource_abun)) {
    resource_abun <- mean(node_abun)
  }

  abun[names(node_abun)] <- node_abun 
  abun[abun == 0] <- resource_abun 
  return(abun)

}

get_network_temporal_plot <- function (
  net_list = NULL,
  station = NULL,
  selected_years = NULL,
  node_color = NULL,
  scale_node_size = 0.01,
  par_options = par(mar = c(0,0,4,0)),
  add_title = FALSE,
  rel_abun = FALSE,
  log_abun = FALSE
  ) {

  net <- get_adj_net_station(net = net_list, station = station)
  bm <- get_net_biomass(net = net_list, station = station)

  if (!is.null(selected_years)) {
    net %<>% filter(year %in% selected_years) 
    bm %<>% filter(year %in% selected_years) 
  }

  if (is.null(node_color)) {
    node_color <- set_color_species(
      node_list = unique(names(unlist(net$obs_troph_level_vector))),
      species_list = NULL, resource_list = NULL, col_resource = NULL)
  }

  list_p <- list()
  index <- seq_along(net$year)

  for (i in index) {
    if(!is.null(par_options)) {
      par_options
    }
    plot.new()
    PlotWeb(
      TL = net$obs_troph_level_vector[[index[i]]],
      webTL = net$adj_mat[[index[i]]],
      colnode = set_color_nodes(
	node_list = names(net$obs_troph_level_vector[[index[i]]]),
	color_species_resources = node_color),
      abund = set_node_resources_abun(
	node_list = names(net$obs_troph_level_vector[[index[i]]]),
	node_abun = bm$bm[[index[i]]],
	resource_abun = 1
	),
      collink="grey70",
      scale_abun = scale_node_size,
      rel_abun = rel_abun,
      log_abun = log_abun
    )
    if (add_title) {
     title(main = paste0(net$year[[index[i]]]))
    }

    p <- recordPlot()

    list_p[[i]] <- p 
  }

  gg <- map(list_p, ggdraw)
  gg
}
```

```{r}
tar_load(metaweb_analysis)
my_species_color <- set_color_species(
  node_list = colnames(metaweb_analysis$metaweb),
  resource_list = metaweb_analysis$resource,
  col_resource = NULL
)
```

```{r}
tar_load(c(com_data, op_data, com_analysis_data))
op_data %<>%
  mutate(year = as.integer(year))

ti <- com_analysis_data %>%
  group_by(opcod) %>%
  summarise(sp_vector = list(setNames(biomass, species))) %>%
  left_join(dplyr::select(op_data, station, opcod, year, surface))

com_data %<>%
    left_join(dplyr::select(op_data, station, opcod, year))

tar_load(gaussian_pred_station_tps10)
tr10 <- gaussian_pred_station_tps10 %>%
  select(response, station, mean) %>%
  pivot_wider(names_from = "response", values_from = "mean")

first_st <- tr10 %>%
  arrange(log_bm_std)

tes <- map(first_st$station[12:24], ~plot_dyn_web(
    com = ti,
    sem_df = NULL,
    station = .x,
#    .log = TRUE,
    stacked = TRUE
    ))
plot_grid(plotlist = tes)
```
```{r}
last_st <- tr10 %>%
  arrange(desc(log_bm_std))

tes <- map(last_st$station[13:24], ~plot_dyn_web(
    com = ti,
    sem_df = NULL,
    station = .x,
#    .log = TRUE,
    stacked = TRUE
    ))
plot_grid(plotlist = tes)
```

```{r}
st <- first_st$station[20]
st[2] <- last_st$station[16]
names(st) <- c("decreasing", "increasing")
#11432 my choice
```

```{r}
tar_load(c(net_analysis_data, net_data))
load(here::here("data", "classes", "network_analysis.rda"))
network_analysis %<>%
  left_join(dplyr::select(op_data, opcod, station, year)) %>%
  filter(!is.na(opcod))

p <- get_network_temporal_plot(
  net_list = network_analysis,
  station = st[1],
  selected_years = c(1995, 2005),
  node_color = NULL, scale_node_size = 0.02,
  par_options = par(mar = c(0,0,0,0)),
  log_abun = FALSE 
)
g <- plot_grid(plotlist = map(p, ggdraw), ncol = 2)

save_plot(
  get_mypath("figures/decreasing_biomass_network.pdf"),
  g,
  nrow = 1,
  ncol = 2,
  base_height = 4,
  base_width = 4)

p2 <- get_network_temporal_plot(
  net_list = network_analysis,
  station = st[2],
  selected_years = NULL,
  node_color = NULL, scale_node_size = 0.06,
  par_options = par(mar = c(0,0,0,0)),
  log_abun = TRUE
)
g2 <- plot_grid(plotlist = map(p2, ggdraw), nrow = 4)

save_plot(
  get_mypath("figures/increasing_biomass_network.png"), g2,
  nrow = 4,
  ncol = 4,
  base_height = 4,
  base_width = 4)
```

```{r}
library(lubridate)

tes <- plot_dyn_web(
    com = mutate(ti, year = ymd(year, truncated = 2L)),
    sem_df = NULL,
    station = st[1],
#    .log = TRUE,
    stacked = TRUE
    )

tps_bm_dec  <- tes +
  theme(legend.position = "none")

save_plot(
  get_mypath("figures/decreasing_biomass.pdf"), tps_bm_dec,
  nrow = 1,
  ncol = 1,
  base_height = 4,
  base_width = 7)

tes2 <- plot_dyn_web(
    com = mutate(com_data, year = ymd(year, truncated = 2L)),
    sem_df = NULL,
    station = st[2],
#    .log = TRUE,
    stacked = TRUE
    )

tps_bm_inc  <- tes2 +
  theme(legend.position = "none")

save_plot(
  get_mypath("figures/increasing_biomass.pdf"), tps_bm_inc,
  nrow = 1,
  ncol = 1,
  base_height = 4,
  base_width = 7)
```

# Plot temporal dynamics

```{r}
loadd(full_data2, rigal_classification)
```
```{r}
rig_st_dec <- rigal_classification %>%
  filter(variable == "ct_ff") %>%
  unnest(classif) %>%
  filter(station == st[1])
```

```{r}
tar_load(full_data2)
tps_ct_dec <- full_data2 %>%
  mutate(year = ymd(year, truncated = 2L)) %>%
  filter(station == st[1]) %>%
  ggplot(aes(x = year, y = ct_ff)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x,
    size = 1, se = FALSE, color = "black") +
  labs(x = "Year", y = "Connectance")

plot_temporal_variable_lm <- function(
  .df = full_data2,
  var = "ct_ff",
  st = 15818,
  col_trend = "black",
  add_median = TRUE,
  col_median = "blue",
  size_line = 1
  ) {
  tex <- .df %>%
    filter(station %in% st) %>%
    mutate(
      year = ymd(year, truncated = 2L)
    )

    if(!is.null(names(st))) {
      tex %<>%
        mutate(station = str_replace_all(station, setNames(names(st), st)))
    }

    p <- tex %>%
      ggplot(aes_string(x = "year", y = var)) +
      geom_point() +
      geom_smooth(
        aes(group = station), method = "lm", formula = y ~ x,
        size = size_line, se = FALSE, color = col_trend) +
      labs(x = "Year", y = str_replace_all(var, var_replacement()))

    if (add_median) {

      p <- p + 
        geom_hline(yintercept = median(tex[[var]]), col = col_median, size =
          size_line)
    }
    return(p)
}

tps_dec_inc <- map(c("ct_ff", "w_trph_lvl_avg", "log_rich_std", "piel_bm"),
  ~plot_temporal_variable_lm(st = st, var = .x) +
  facet_grid(cols = vars(station))
)

plot_grid(plotlist = tps_dec_inc)
```

## plot map

```{r}
loadd(map_st)
save_plot(map_st,
  base_height = 11,
  base_asp = 1.618,
  file = get_mypath("figures", "map_st.png")
)
```

```{r}
test_fig_mat_met <- plot_grid(
  map_st,
  plot_grid( plot_grid(tps_bm_dec, tps_bm_inc)),
  plot_grid(plotlist = tps_dec_inc),
  nrow = 3
)
save_plot(
  test_fig_mat_met,
  base_height = 12,
  base_asp = 1.618,
  file = get_mypath("figures", "matmet.png")
)
```

# New figure

```{r, eval = FALSE}
# get a station that represents the results
st_var_shape <- rigal_classification %>%
  unnest(classif) %>%
  filter(variable %in% c(get_com_str_var(), "log_bm_std")) %>%
  select(station, variable, direction)  %>%
  pivot_wider(names_from = variable, values_from = direction)
dec_representative <- st_var_shape %>%
  filter(
    log_bm_std == "decrease",
    ct_ff == "decrease",
    w_trph_lvl_avg == "decrease")
```

```{r}
color_tes <- map(st[1],
  ~set_color_species(
  node_list = ti %>%
  filter(station == .x) %$%
  .$sp_vector %>%
  reduce(., c) %>%
  names(.) %>%
  unique(),
  resource_list = metaweb_analysis$resource,
  col_resource = NULL
  )
)
names(color_tes) <- as.character(st[1])

tes <- map(st[1], ~plot_dyn_web(
    com = ti,
    sem_df = NULL,
    station = .x,
#    .log = TRUE,
    stacked = TRUE,
    color_species = color_tes[[as.character(.x)]]
    ))

names(tes) <- as.character(st["decreasing"])
plot_grid(plotlist = tes)
```
```{r}
st_decreasing <- st["decreasing"]

p_dec_log_bm <- tes[[as.character(st_decreasing)]] +
  labs(color = "Species", fill = "Species") +
  theme(legend.position = c(.2, .9), legend.direction = "horizontal")

save_plot(
  get_mypath("figures/p_dec_log_bm.pdf"),
  p_dec_log_bm,
  nrow = 1,
  ncol = 1,
  base_height = 4,
  base_width = 7
  )
```

```{r}
tar_load(c(net_analysis_data, net_data))
load(here::here("data", "classes", "network_analysis.rda"))
network_analysis %<>%
  left_join(dplyr::select(op_data, opcod, station, year)) %>%
  filter(!is.na(opcod))

p <- get_network_temporal_plot(
  net_list = network_analysis,
  station = st[1],
  selected_years = c(1995, 2005),
  node_color = color_tes[[st[1]]], scale_node_size = 0.05,
  par_options = par(mar = c(0, 0, 0, 0)),
  log_abun = TRUE 
)
g <- plot_grid(plotlist = map(p, ggdraw), ncol = 2)

save_plot(
  get_mypath("figures/decreasing_biomass_network.pdf"),
  g,
  nrow = 1,
  ncol = 2,
  base_height = 4,
  base_width = 4)

p2 <- get_network_temporal_plot(
  net_list = network_analysis,
  station = st[2],
  selected_years = NULL,
  node_color = NULL, scale_node_size = 0.06,
  par_options = par(mar = c(0,0,0,0)),
  log_abun = TRUE
)
g2 <- plot_grid(plotlist = map(p2, ggdraw), nrow = 4)

save_plot(
  get_mypath("figures/increasing_biomass_network.png"), g2,
  nrow = 4,
  ncol = 4,
  base_height = 4,
  base_width = 4)
```

```{r}
v_colors =  viridisLite::viridis(n = 2, begin = 0.5, end = 1)
names(v_colors) <- c("tps", "sp")

tps_dec_inc <- map(c("log_bm_std", "log_rich_std", "ct_ff", "w_trph_lvl_avg"), # 
  ~plot_temporal_variable_lm(st = st_decreasing, var = .x, col_median =
    v_colors["sp"], col_trend = v_colors["tps"]) +
    theme(axis.title = element_text(size = 12))
)
# Increase point size
tps_dec_inc  <- map(tps_dec_inc, function(x) {
  x$layers[[1]]$aes_params$size <- 4
  return(x)
})

# Delete x axis
#tps_dec_inc[-length(tps_dec_inc)]  <- map(tps_dec_inc[-length(tps_dec_inc)], function(x) {
#  x + theme(
#    axis.text.x = element_blank(),
#    axis.ticks.x = element_blank(),
#    axis.title.x = element_blank()
#  )
#})

save_plot(
  filename = get_mypath("figures/net_str_dec_bm.pdf"),
  plot_grid(plotlist = tps_dec_inc, nrow = 1, align = "h", axis = "l"),
  ncol = length(tps_dec_inc), nrow = 1, 
  base_height = 2.6,
  base_asp = 1.2 
)
```

```{r temporal foodwebs}
p <- get_network_temporal_plot(
  net_list = network_analysis,
  station = st_decreasing,
  selected_years = c(1995, 2005, 2015),
  node_color = color_tes[[as.character(st_decreasing)]],
  scale_node_size = 0.027,
  par_options = par(mar = c(0,0,0,0)),
  log_abun = TRUE
)
g <- plot_grid(plotlist = map(p, ggdraw), ncol = 3)

save_plot(
  get_mypath("figures/decreasing_biomass_network.pdf"),
  g,
  ncol = 3, nrow = 1,
  base_width = 7, base_height = 7)
```

## Plot PCA

```{r}
loadd(p_pca)
save_plot(get_mypath("figures/rotated_pca.pdf"),
  p_pca,
  ncol = 1, nrow = 1,
  base_width = 4, base_height = 4
)
```

## plot map

```{r}
loadd(map_st)
save_plot(map_st,
  base_height = 11,
  base_asp = 1.618,
  file = get_mypath("figures", "map_st.png")
)
```

```{r}
loadd(station_analysis, st_mono_trends_stable_rich_bm)
```
```{r}
fr_map <- ne_countries(scale = "medium", returnclass = "sf") %>%
  filter(geounit == "France") %>%
  st_transform(crs = 2154)
st <- station_analysis %>%
  filter(id %in% st_mono_trends_stable_rich_bm) %>%
  st_transform(crs = 2154)

p <- ggplot() +
    geom_sf(data = st_geometry(fr_map), lwd = 0) +
    geom_sf(data = st_geometry(st)) +
    coord_sf(
      xlim = st_bbox(st)[c(1, 3)],
      ylim = st_bbox(st)[c(2, 4)]
    ) +
  theme_bw() +
  theme(
    panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.ticks = element_blank(), axis.text = element_blank())
p
```
```{r}
save_plot(
  p,
  base_height = 12,
  base_asp = 1.618,
  file = get_mypath("figures", "simple_map_matmet.pdf")
)
```

```{r}
test_fig_mat_met <- plot_grid(
  map_st,
  plot_grid( plot_grid(tps_bm_dec, tps_bm_inc)),
  plot_grid(plotlist = tps_dec_inc),
  nrow = 3
)
save_plot(
  test_fig_mat_met,
  base_height = 12,
  base_asp = 1.618,
  file = get_mypath("figures", "matmet.png")
)
```

```{r}
test_mat_met <- plot_grid(
  plot_grid(map_st, p_dec_log_bm, ncol = 2),
  plot_grid(plot_grid(plotlist = tps_dec_inc), g, ncol = 2),
  nrow = 2
)
save_plot(
  get_mypath("figures/test_mat_met.png"),
  test_mat_met,
  base_asp = .5,
  base_width = 9
)
```